---
title: "Air Temperature Daily Mins and Max"
author: "Elizabeth Van Metre"
date: '2023-01-26'
output: html_document
---

#How to Use This Template
1. Change "data_path" variable input to file path for air temp data on line #17
2. Check line #48 for note on an invalid sensor - this column will be excluded
-> you need to check if this matches up with new data
3. For calculating degree days, we use a baseline temperature (in 째C) to compare daily averages
-> change DD_baseline on line #18 if using a baseline other than 0째C
4. Run the code!

```{r}
data_path = "~/Desktop/Arctic/Season_AIRTMP_6_19_22_9_20_22_2022_09_21_16_21_48_UTC_1.csv"
DD_baseline = 0
```


##Loading packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(splitstackshape) # basically like text to columns
library(naniar)
library(knitr)
```


#Upload air temp data and clean
```{r}
airtemps =read.csv(data_path)

names(airtemps) <- sub("Temperature..RXW.THC.", "" ,names(airtemps))
names(airtemps) <- sub("Temperature..S.THC.", "" ,names(airtemps))
names(airtemps) <- sub("S.", "" ,names(airtemps))
names(airtemps) <- sub("....C..RX3000_BRW1", "" ,names(airtemps))
names(airtemps) <- sub("....C..RX3000_BRW4", "" ,names(airtemps))
names(airtemps) <- sub("....C..RX3000_BRW5", "" ,names(airtemps))
names(airtemps) <- sub("....C..RX3000_BRW6", "" ,names(airtemps))
airtemps$Date <- as.POSIXct(airtemps$Date,format="%m/%d/%y %H:%M",tz="UTC")

airtemps = airtemps[-4] #NOTE: column 4 in original data is an invalid Sensor - check for new data
```

##Remove bad readings (Temp < 100 degrees C)
```{r}
for(r in 1:nrow(airtemps)) {
  for(c in 3:ncol(airtemps)) {
    if(airtemps[r,c] < -100 | is.na(airtemps[r,c])) {
      airtemps[r,c] = NA
    }
  }
}
```


#Sensor Reference with Names
Commented out - uncomment if you need the reference
Note: does not change for new data - originally for 6/19-9/20 
```{r}

 c = 3
 sensorID = NULL
 columnNum = NULL
 while(c <= ncol(airtemps)){
   sensorID = append(sensorID, names(airtemps)[c])
   columnNum = append(columnNum, c)
   c = c + 1
 }
location = c("TNHA", "TNHA", "TNHA", "TNHA", "TNHA", "SSMH", "SSMH", "SSMH", "BUECI",
              "BUECI", "BUECI", "BUECI", "BUECI", "BUECI", "BEO")
ID = c("SA", "SD", "SC", "SB", "B02", "B02", "SB01", "SA01", "SE01", "SB01", "B02", "SD02", "SC01", "SA01", "B02")
sensors = data.frame(columnNum, sensorID, location, ID)

kable(sensors)

```

##How might we be able to remove zeroes?

Turn Dates into Day Format
```{r}
airtemps$Date = as.Date(airtemps$Date)
```

Pull Each Sensor
```{r}

maximum = setNames(aggregate(airtemps[,3], by=list(airtemps$Date),max, na.rm=T), c("Date", paste(sensors[1, 3],           sensors[1,4])))
minimum = setNames(aggregate(airtemps[,3], by=list(airtemps$Date),min, na.rm=T), c("Date", paste(sensors[1, 3],           sensors[1,4])))

maxs = maximum
mins = minimum


```

While-Loop: Join Mins and Maxes to 1 Table by Date
```{r}
i = 4

while(i <= ncol(airtemps)) {
  
maximum = setNames(aggregate(airtemps[,i], by=list(airtemps$Date),max, na.rm=T), c("Date", paste(sensors[i-1, 3],           sensors[i-1,4])))
minimum = setNames(aggregate(airtemps[,i], by=list(airtemps$Date),min, na.rm=T), c("Date", paste(sensors[i-1, 3],           sensors[i-1,4])))

maxs = maxs %>% full_join(maximum, by="Date")
mins = mins %>% full_join(minimum, by="Date")

i = i + 1

}
```

#Average Temperature by Site

Convert to Long DataFrame
```{r}

#Turn wide data frames into long
mins_long = gather(mins, sensor, min_air_temp, 2:16)
maxs_long = gather(maxs, sensor, max_air_temp, 2:16)

#Split sensor name and station
mins_long = cSplit(mins_long, 'sensor', sep=" ", direction = "wide")
maxs_long = cSplit(maxs_long, 'sensor', sep=" ", direction = "wide")
#rename resulting columns
colnames(mins_long)[3:4] <- c("station", "sensor")
colnames(maxs_long)[3:4] <- c("station", "sensor")
```

```{r}

#Turn wide data frames into long
mins_long = gather(mins, sensor, min_air_temp, 2:16)
maxs_long = gather(maxs, sensor, max_air_temp, 2:16)

#Combine data frames into one


min_max <- full_join(mins_long, maxs_long)
min_max = min_max %>% filter(max_air_temp < 100)

#texttocolumns
library(splitstackshape)
min_max <- cSplit(min_max, "sensor", sep = " ", direction = "wide")
min_max$sensor <- paste(min_max$sensor_1, min_max$sensor_2, sep="-") 
min_max$sensor_2 <- NULL
colnames(min_max)[4] <- "station"
colnames(min_max)[5] <- "sensor"


```


Get Average (Don't Run More Than Once!)
```{r}

avg_temp = (min_max[,2] + min_max[,3])/2
min_max = data.frame(min_max, avg_temp)
colnames(min_max)[6] <- "avg_temp"

min_max <- data.frame(min_max)


```

Get Daily Air Temperature Average by Station
```{r}

TNHA = min_max %>% filter(station == "TNHA")
SSMH = min_max %>% filter(station == "SSMH")
#BUECI = min_max %>% filter(station == "BUECI")
BEO = min_max %>% filter(station == "BEO")


TNHA_avg = aggregate(avg_temp~Date, TNHA, mean, na.rm=T)
SSMH_avg = aggregate(avg_temp~Date, SSMH, mean, na.rm=T)
#BUECI_avg = aggregate(avg_temp~Date, BUECI, mean, na.rm=T)
BEO_avg = aggregate(avg_temp~Date, BEO, mean, na.rm=T)
  
```

#Degree Days Calculation
Degree Days = Avg Daily Temp - Baseline Temp
Our Baseline: 0째C (or otherwise defined at beginning)
```{r}
TNHA_DD = TNHA_avg[,2] - DD_baseline
SSMH_DD = SSMH_avg[,2] - DD_baseline
#BUECI_DD = BUECI_avg[,2] - DD_baseline
BEO_DD = BEO_avg[,2] - DD_baseline

TNHA_avg = data.frame(TNHA_avg, TNHA_DD)
SSMH_avg = data.frame(SSMH_avg, SSMH_DD)
#BUECI_avg = data.frame(BUECI_avg, BUECI_DD)
BEO_avg = data.frame(BEO_avg, BEO_DD)
```

##Degree Day Totals
```{r}
TNHA_DD_tot = sum(TNHA_avg[,3])
SSMH_DD_tot = sum(SSMH_avg[,3])
#BUECI_DD_tot = sum(BUECI_avg[,3])
BEO_DD_tot = sum(BEO_avg[,3])
```

##Output Degree Day Totals
```{r}
dr = paste("Date Range:",TNHA_avg[1,1], "to", TNHA_avg[nrow(TNHA_avg),1])
dr
```

TNHA:
```{r}
TNHA_DD_tot
```

SSMH:
```{r}
SSMH_DD_tot
```

BUECI:
```{r}
#BUECI_DD_tot
```

BEO:
```{r}
BEO_DD_tot
```
MAX, MIN, AVG AIR TEMPS AT SELECT STATION OVER SUMMER MONTHS
```{r}
min_max %>% filter(!avg_temp < -5, !avg_temp > 50, station == "SSMH", Date>="2022-07-04" & Date<="2022-07-10") %>%
  ggplot() + 
  geom_point(aes(x = Date, y = min_air_temp, color = "min")) +
  geom_point(aes(x = Date, y = max_air_temp, color = "max")) +
  geom_line(aes(x = Date, y = avg_temp, color = "average")) +
  guides(color = guide_legend(override.aes = list(size=0.5)))+
  theme_bw() +
  labs(y="Temperature (째C)",x="End of Week",color="Variable", title = "SSMH") +
  scale_x_date(date_breaks = "1 day") +
  ylim(-10,35) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic"))
```
