---
title: "ANNUAL_DIFFERENCES_PLOTTING"
author: "Mirella Shaban"
date: "2024-01-24"
output: html_document
---

THIS SCRIPT MUST BE RUN *AFTER* THE ANNUAL_DIFFERENCES SCRIPT (WHICH LOADS AND CLEANS THE DATA)

THIS IS SOLELY A STATISTICS AND PLOTTING SCRIPT for DIFFERENCES B/W YEARS

####################################################################################################################################
                                                PLOTTING 2022 VS 2023 DATA (summer and fall)
####################################################################################################################################
*ground temp:*
summer 22: summer22grndtmp
summer 23: sum23grndtmp

combined summer data (by station daily): SUMMER_22_23_GRNDTMP
combined summer data (by SITE) annually: Sum22_Sum23_Annual_Temps

fall 22: fall22grndtemp
fall 23:
combined fall data:

*air temp:* (code to clean airtemp data is in the Differences_in_averages.RMD)
summer 22: summer22airtemp
summer 23: summer23airtemp
combined summer data:

fall 22: fall22airtemp
fall 23:
combined fall data:

####################################################################################################################################

####################################################################################################################################

*GROUND TEMP*

STATS: average daily difference between SITES (all stations combined) for Xcm depth (define in x <- )
```{r}
wide <- sum23grndtmp %>%
    pivot_wider(names_from = station, values_from = value)

wide$TBDIFF <- wide$TNHA - wide$BEO
wide$SBDIFF <- wide$SSMH - wide$BEO
wide$TBSBDIFF <- wide$SSMH - wide$TNHA

x<- wide %>% filter(depth == "10cm") 
mean(x$TBDIFF,na.rm = TRUE)
mean(x$SBDIFF, na.rm = TRUE)
mean(x$TBSBDIFF, na.rm = TRUE)


xxx <- sum23grndtmp %>% filter(station == "TNHA" & value == is.na(FALSE) & depth == "10cm" )
```

AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station$fullname <- c("TNHA-SA")
averaged_by_station2 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date+depth, mean)
averaged_by_station2$fullname <- c("TNHA-SC")
averaged_by_station3 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "SSMH-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station3$fullname <- c("SSMH-SA")
averaged_by_station4 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "SSMH-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station4$fullname <- c("SSMH-SB")
averaged_by_station5 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06") %>% aggregate(value~date+depth, mean)
averaged_by_station5$fullname <- c("BEO")
TNHA_NvsS <- rbind(averaged_by_station, averaged_by_station2, averaged_by_station3,averaged_by_station4,averaged_by_station5)
```


STATS: average daily difference between SPECIFIC STATIONS for Xcm depth (define in x <- )
```{r}
wide <- TNHA_NvsS %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO`

wide$SSMH_NS_DIFF <- wide$`SSMH-SA` - wide$`SSMH-SB`
wide$SSMH_N_BEO_DIFF <- wide$`SSMH-SB` - wide$`BEO`
wide$SSMH_S_BEO_DIFF <- wide$`SSMH-SA` - wide$`BEO`

wide$TNHA_SSMH_N_COMP <- wide$`TNHA-SC`-wide$`SSMH-SB`
wide$TNHA_SSMH_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SA`

#write.csv(wide, "~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_daily_average_difference_N_vs_S.csv")

x<- wide %>% filter(depth == "30cm") 

mean(x$TNHA_NS_DIFF,na.rm = TRUE)
mean(x$SSMH_NS_DIFF, na.rm = TRUE)


mean(x$TNHA_S_BEO_DIFF, na.rm = TRUE)
mean(x$SSMH_S_BEO_DIFF, na.rm = TRUE)

max(x$TNHA_NS_DIFF, na.rm = TRUE)
max(x$SSMH_NS_DIFF, na.rm = TRUE)

x<- wide %>% filter(depth == "90cm") 
#below, p1 = 6/19-8/01 , p2= 8/01-
p1 = with(x, mean(TNHA_SSMH_N_COMP[date >= "2022-06-19" & date <= "2022-07-08"]))
p1

```

# might want to delete the stats above, work only with 22 vs 23?



Plot 2022 vs 2023 for a specified station

```{r fig1, fig.height = 18, fig.width = 16}
SUMMER_22_23_GRNDTMP %>% filter(depth == "30cm" & !fullname == "FAKEVAR" & !fullname == "FAKEVAR1" & !fullname == "FAKEVAR2" & !fullname == "FAKEVAR3" & !fullname == "FAUX1" & !fullname == "FAUX2" & !fullname == "FAUX3" & !fullname == "FAUX4" & !fullname == "FAUX5" & station == "SSMH") %>%
  
  ggplot(aes(x = day, y = value, colour = factor(year))) +
  geom_line() +
  geom_point() +
  theme_bw()+
  theme(legend.title = element_text(size = 25),legend.text = element_text(size = 20)) +
  labs(title = "SUMMER COMPARISONS (Daily average ground temperatures), 30CM", colour = "Year", x = "Date", y = "Temperature (°C)") +
    scale_x_date(date_labels = "%b-%d", date_breaks =  "1 weeks") +
    scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20),
        strip.text= element_text(size=35, face = "bold"),
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 30),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  facet_wrap(~fullname,ncol=1,strip.position="top",)

```

```{r fig2, fig.height = 18, fig.width = 16}
FALL_22_23_GRNDTMP %>% filter(depth == "30cm" & !fullname == "FAKEVAR" & !fullname == "FAKEVAR1" & !fullname == "FAKEVAR2" & !fullname == "FAKEVAR3" & !fullname == "FAUX1" & !fullname == "FAUX2" & !fullname == "FAUX3" & !fullname == "FAUX4" & !fullname == "FAUX5" & station == "SSMH") %>%
  # make datetime: force unique year
  #mutate(datetime = lubridate::make_date(2020, month, day)) %>%  #left off at this point, need to be able to fix this 
  
  ggplot(aes(x = day, y = value, colour = factor(year))) +
  geom_line() +
  geom_point() +
  theme_bw()+
  theme(legend.title = element_text(size = 25),legend.text = element_text(size = 20)) +
  labs(title = "FALL COMPARISONS (Daily average ground temperatures), 30CM", colour = "Year", x = "Date", y = "Temperature (°C)") +
  scale_x_date(date_labels = "%b-%d", date_breaks =  "1 weeks") +
    scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20),
        strip.text= element_text(size=35, face = "bold"),
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 30),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  facet_wrap(~fullname,ncol=1,strip.position="top",)

```

*CONSTRAIN DATE RANGES FOR SPECIFIC STATIONS TO COMPARE 22 VS 23 TEMPERATURES DURING THOSE DOY'S*
summer
TNHA-BASE: NA
TNHA-SA: 7/9-7/15 *
TNHA-SB: NA
TNHA-SC: 7/9-7/15 *
TNHA-SD: 7/9-7/15 *
SSMH-BASE: NA
SSMH-SA: 6/28-7/04 *
SSMH-SB: 8/15-9/01 *
*SSMH-SD: 8/15-9/01 *NOT TRUE*
BEO: 7/4-7/15 (BEO-B05 VS BEO-B05) --> FOR SUM 23, CAN AGGREGATE B05 & B06, BOTH HAVE DATA IN DATE RANGE *

fall
TNHA-BASE: 9/15-9/22 *
TNHA-SA: NA *
TNHA-SB: 9/20-10/04 *
TNHA-SC: 9/15-10/21 *
TNHA-SD: NA
SSMH-BASE: NA
SSMH-SA:NA
*below, do together*
SSMH-SB:9/1-10/16 *
SSMH-SD: 9/1-10/16 *
BEO: 9/16-10/16 (BEO-B05 VS BEO-B07) *

*above, need to adjust the BEO fall figure so both 05 and 07 are on the same figure...*


Bounds the date ranges
```{r}
SUMMER_22_23_GRNDTMP
FALL_22_23_GRNDTMP

#need to check the DEPTH and the DATES

Summ_22_23_daterange = SUMMER_22_23_GRNDTMP %>% filter(day >= "2024-08-15" & day <= "2024-09-01" & (fullname == "SSMH-SB") & depth == "30cm")

Fall_22_23_daterange = FALL_22_23_GRNDTMP %>% filter(day >= "2024-09-16" & day <= "2024-10-16" & (fullname == "BEO-B05" | fullname == "BEO-B07") & depth == "30cm")

```

plots the bound range
```{r fig2, fig.height = 18, fig.width = 16}
#check the FIGURE SIZE (one plot in facet, change to 8. more than 1 plot, change to 18)

Fall_22_23_daterange %>% filter(depth == "30cm" & !fullname == "FAKEVAR" & !fullname == "FAKEVAR1" & !fullname == "FAKEVAR2" & !fullname == "FAKEVAR3" & !fullname == "FAUX1" & !fullname == "FAUX2" & !fullname == "FAUX3" & !fullname == "FAUX4" & !fullname == "FAUX5") %>%
  # make datetime: force unique year
  #mutate(datetime = lubridate::make_date(2020, month, day)) %>%  #left off at this point, need to be able to fix this 
  
  ggplot(aes(x = day, y = value, colour = factor(year))) +
  geom_line() +
  geom_point() +
  theme_bw()+
  theme(legend.title = element_text(size = 25),legend.text = element_text(size = 20)) +
  labs(title = "FALLL (09/16-10/16) COMPARISONS (Daily average ground temp), 30CM", colour = "Year", x = "Date", y = "Temperature (°C)") +
  scale_x_date(date_labels = "%b-%d", date_breaks =  "1 weeks") +
    scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20),
        strip.text= element_text(size=35, face = "bold"),
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 30),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  facet_wrap(~fullname,ncol=1,strip.position="top",)

```









                     #######
                           ### VWC!!! ####
                                        #######
```{r fig3, fig.height = 18, fig.width = 16}
SUMMER_22_23_VWC_DAILY %>% filter(depth == "15:30cm" & !fullname == "FAKEVAR" & !fullname == "FAKEVAR1" & !fullname == "FAKEVAR2" & !fullname == "FAKEVAR3" & !fullname == "FAUX1" & !fullname == "FAUX2" & !fullname == "FAUX3" & !fullname == "FAUX4" & !fullname == "FAUX5" & station == "SSMH") %>%
  
  ggplot(aes(x = day, y = value, colour = factor(year))) +
  geom_line() +
  geom_point() +
  theme_bw()+
  theme(legend.title = element_text(size = 25),legend.text = element_text(size = 20)) +
  labs(title = "SUMMER COMPARISONS (Daily average VWC), 15:30CM", colour = "Year", x = "Date", y = "VWC ("~ m ^ 3 ~ "/" ~ m ^ 3 ~")") +
  scale_x_date(date_labels = "%b-%d", date_breaks =  "1 weeks") +
  scale_y_continuous(breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6), limits=c(0, 0.6)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20),
        strip.text= element_text(size=35, face = "bold"),
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 30),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  facet_wrap(~fullname,ncol=1,strip.position="top",)
```

```{r fig4, fig.height = 18, fig.width = 16}
FALL_22_23_VWC_DAILY %>% filter(depth == "15:30cm" & !fullname == "FAKEVAR" & !fullname == "FAKEVAR1" & !fullname == "FAKEVAR2" & !fullname == "FAKEVAR3" & !fullname == "FAUX1" & !fullname == "FAUX2" & !fullname == "FAUX3" & !fullname == "FAUX4" & !fullname == "FAUX5" & station == "BEO" |sensor == "BEO-NEW") %>%
  
  ggplot(aes(x = day, y = value, colour = factor(year))) +
  geom_line() +
  geom_point() +
  theme_bw()+
  theme(legend.title = element_text(size = 25),legend.text = element_text(size = 20)) +
  labs(title = "FALL COMPARISONS (VWC), 15:30CM", colour = "Year", x = "Date", y = "VWC ("~ m ^ 3 ~ "/" ~ m ^ 3 ~")") +
  scale_x_date(date_labels = "%b-%d", date_breaks =  "1 weeks") +
  scale_y_continuous(breaks=c(0,0.1,0.2,0.3,0.4,0.5), limits=c(0, 0.5)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20),
        strip.text= element_text(size=35, face = "bold"),
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 30),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  facet_wrap(~fullname,ncol=1,strip.position="top",)
```

