---
title: "Averaged_Figures"
author: "Mirella Shaban"
date: "2023-04-12, updated 2023-04-24"
output: html_document
---

 Loading packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape) 
library(naniar)
library(RColorBrewer)
library(clock)
```
Ground Temperatures for all sites data prepping
```{r}
# importing ground temp data for 6-19 through 9-20
grndtmp619_920=read.csv("C:/Users/qwe2qh/OneDrive - University of Virginia/UVA-OfficeComputer/Utqiagvik/Meteorological_Seasons_Data/Summer_grndtmp_2023_04_06_12_21_30_UTC_1.csv")
names(grndtmp619_920) <- sub("....C..RX3000_BRW1", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW4", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW5", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW6", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub(".RXW.GP6.", "", names(grndtmp619_920))
grndtmp619_920$Date <-as.POSIXct(grndtmp619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```


Ground Moisture for all sites data prepping
```{r}
mst619_920=read.csv("C:/Users/qwe2qh/OneDrive - University of Virginia/UVA-OfficeComputer/Utqiagvik/Meteorological_Seasons_Data/Summer_VWC_2023_04_06_12_33_54_UTC_1.csv")
names(mst619_920) <- sub("...m.3.m.3..RX3000_BRW1", "", names(mst619_920))
names(mst619_920) <- sub("...m.3.m.3..RX3000_BRW4", "", names(mst619_920))
names(mst619_920) <- sub("...m.3.m.3..RX3000_BRW5", "", names(mst619_920))
names(mst619_920) <- sub("...m.3.m.3..RX3000_BRW6", "", names(mst619_920))
names(mst619_920) <- sub("..RXW.GP6.", "", names(mst619_920))
mst619_920$Date <-as.POSIXct(mst619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

Solar radiation for all sites data prepping
```{r}
solarrad619_920 = read.csv("C:/Users/qwe2qh/OneDrive - University of Virginia/UVA-OfficeComputer/Utqiagvik/Meteorological_Seasons_Data/Summer_solar_2023_04_06_13_16_22_UTC_1.csv")
names(solarrad619_920) <- sub("olar.Radiation..RXW.LIB", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("olar.Radiation..S.LIB", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("S.", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW1", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW4", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW5", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW6", "" ,names(solarrad619_920))
solarrad619_920$Date <- as.POSIXct(solarrad619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

Air temp for all sites data prepping
```{r}
airtemp619_920 =read.csv("C:/Users/qwe2qh/OneDrive - University of Virginia/UVA-OfficeComputer/Utqiagvik/Meteorological_Seasons_Data/Summer_airtemp_2023_04_06_11_59_55_UTC_1.csv")
names(airtemp619_920) <- sub("Temperature..RXW.THC.", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("Temperature..S.THC.", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("S.", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW1", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW4", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW5", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW6", "" ,names(airtemp619_920))
airtemp619_920$Date <- as.POSIXct(airtemp619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

Remove Sneaky Air Temperature from ground temp data (*)
```{r}
grndtmp619_920 = grndtmp619_920[,-14]
```

Removing ground temp zeroes occuring at multiple (3) consecutive depths for loop
```{r}
c = 3 
zeroes = 0
while(c <= ncol(grndtmp619_920)-10){
  sensorGroup = grndtmp619_920 %>% select(2,c, c+1, c+2, c+3, c+4, c+5, c+6, c+7, c+8, c+9, c+10)
  for(r in 1:nrow(sensorGroup)) {
    zeroes = 0
    col = 2
    while(col <= 12){
      if(sensorGroup[r,col] == 0 | is.na(sensorGroup[r,col])) { 
        if(!is.na(sensorGroup[r,col])){
          zeroes = zeroes + 1 
        }
      }
      col = col + 1
    }
    if(zeroes >= 3){ 
      for(columnn in 2:ncol(sensorGroup)){
      if(sensorGroup[r,columnn] == 0 | is.na(sensorGroup[r,columnn])) { 
        grndtmp619_920[r,columnn+c] = NA
      }
    }
    }
  }
  c = c + 11 
}
```

LONG DATAFRAMES
```{r}
dataframelong = gather(mst619_920, variable, response, Water.Content21198259.21206939.1:Water.Content21401803.21398591.6) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

df = cSplit(dataframelong, 'variable', sep=".", direction = "wide") 
colnames(df)[5:7] <- c("station", "sensor", "depth")
df$variable_1 <- NULL 
df$depth <- as.factor(as.character(df$depth)) 
df$station <- as.factor(df$station)

df$sensor <- as.factor(as.character(df$sensor))

df$station <- revalue(df$station, c("Content21198259" = "TNHA", "Content21401800" = "BUECI", "Content21401801" = "SSMH", "Content21401803" = "BEO")) 
df$depth <- revalue(df$depth, c("1" = "0:15cm", "2" = "15:30cm", "3" = "30:45cm", "4" = "45:60cm", "5" = "60:75cm", "6" = "75:90cm"))

df$sensor <- revalue(df$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05")) 
```

```{r}
grndtmp = gather(grndtmp619_920, variable, response, Temperature.21198259.21206939.7:Temperature..RXW.THC.21401800.21398671.1) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
dfgrndtemp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(dfgrndtemp)[3] <- ("value")
colnames(dfgrndtemp)[5:7] <- c("station", "sensor", "depth")

dfgrndtemp$variable_5 <- NULL
dfgrndtemp$variable_6 <- NULL
dfgrndtemp$variable_7 <- NULL
dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth))
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))
dfgrndtemp$station <- as.factor(dfgrndtemp$station)
dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))

dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))

```


```{r fig9, fig.heigt = 14, fig.width = 18}
solarrad = gather(solarrad619_920, variable, response, '21198259.21176526.1':'21401800.21390411.1') %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

solarrad= cSplit(solarrad, 'variable', sep=".", direction = "wide") # texttocols, 
solarrad$variable_3 <- NULL
colnames(solarrad)[3:5] <- c("value", "station", "sensor")
solarrad$value <- as.numeric(as.character(solarrad$value)) 
solarrad$sensor <- as.factor(as.character(solarrad$sensor))
solarrad$station <- as.factor(solarrad$station)

solarrad$station <- revalue(solarrad$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO"))
solarrad$sensor <- revalue(solarrad$sensor, c("21390411" = "BUECI-BASE", "21398618" = "BUECI-SA", "21398624" = "BUECI-SB", "21362313" = "BUECI-SC", "21362316" = "BUECI-SD", "21362320" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", 
                                              "21390413" ="SSMH-BASE", "21398622" = "SSMH-SA", "21362319" = "SSMH-SB", "21166008" = "SSMH-SC", "21393049" = "SSMH-SD" ,
                                              "21176526" = "TNHA-BASE", "21398620" = "TNHA-SA", "21398616"="TNHA-SB","21362315" = "TNHA-SD", "21362317" = "TNHA-SC", 
                                              "21390415" = "BEO-Base"))

```

```{r}
airtemp = gather(airtemp619_920, variable, response, "21198259.21398666.1":"21401800.21398671.1") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

airtemp= cSplit(airtemp, 'variable', sep=".", direction = "wide") 
airtemp$variable_3 <- NULL
colnames(airtemp)[3:5] <- c("value", "station", "sensor")
airtemp$value <- as.numeric(as.character(airtemp$value)) 
airtemp$sensor <- as.factor(as.character(airtemp$sensor))
airtemp$station <- as.factor(airtemp$station)

airtemp$station <- revalue(airtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) 


airtemp$sensor <- revalue(airtemp$sensor, c("21390851" = "BUECI-BASE", "21398671" = "BUECI-SA", "21362254" = "BUECI-SB", "21398668" = "BUECI-SC", "21398660" = "BUECI-SD", "21187245" = "BUECI-SE",
                                            "21390849" ="SSMH-BASE","21393049" = "SSMH-SD", "21398670" = "SSMH-SA", "21398665" = "SSMH-SB",
                                            "21218018" = "TNHA-BASE", "21380919" = "TNHA-SA" , "21398676"="TNHA-SB","21398666" = "TNHA-SD", "21398674" = "TNHA-SC", "21397541" = "BEO-BASE"))
#TNHA-A and TNHA-B are GUESSES - need metadata on ground
```

MADMINMED AIR TEMPERATURE : LONG DATAFRAME
```{r}
newairtemp = gather(Weekly_Airtemp_MaxMinMed, variable, response, Max, Min, p) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))


newairtemp1= cSplit(newairtemp, 'variable', sep=".", direction = "wide") 

newairtemp1$sensor <- as.factor(as.character(newairtemp1$sensor))



#Plotting max, min, med values for air temp per week 
```



------
Daily average air temperature per sensor at station "X"
```{r}
tnha_airmean <- airtemp %>% subset(station == c("TNHA")) %>% select(value,sensor,Date) %>% filter(!value == "0.0", !value == "-888.8", !sensor == "21398664") 

tnha_airmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

tnha_airmean$time <- format(as.Date(
  tnha_airmean$Date),format = "%H:%M:%S")
tnha_airmean$day <- as.Date(tnha_airmean$Date)


Daily_average_airtmp_per_sensor <- tnha_airmean %>% 
  group_by(sensor, day) %>% filter(!sensor == '21398664') %>%
  aggregate(value ~ sensor + day, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season

```


Weekly Max,Min,Median averages for air temp
```{r fig20, fig.heigt = 4, fig.width = 8}
maxminmedair<- tnha_airmean %>% filter(!sensor == '21398664') %>%
  group_by(day,sensor) %>%
  summarise(Max = max(value, na.rm = T),Min = min(value, na.rm = T), p = median(value,na.rm = T))
#above is daily max, min, med per sensor

maxminmedair$Week_Day <- as.numeric(format(maxminmedair$day, format='%w'))
maxminmedair$End_of_Week <- maxminmedair$day + (6 - maxminmedair$Week_Day)

Weekly_Airtemp_MaxMinMed <- maxminmedair %>% group_by(sensor, time = as.Date(End_of_Week)) %>% summarise(across(c(Max, Min, p), mean))

```



AVERAGING OF SOLAR RAD DATA, AVERAGE DAILY VALUES (all sensors combined)
```{r}
#TNHA AVG DAILY SOLAR ALL SENSORS COMBINED
avgdailysolar <- solarrad %>% filter(!station == "BEO", !station == "SSMH", !station == "BUECI") 
  aggregate(list(Value = avgdailysolar$value), 
          list(dailyavg = cut(avgdailysolar$Date, "1 day")), 
          mean)

#TNHA AVG DAILY SOLAR, SELECT SENSOR  
avgdailysolar <- solarrad %>% filter(!station == "BEO", !station == "SSMH", !station == "BUECI", !sensor == "TNHA-SD", !sensor == "TNHA-BASE", !sensor == "TNHA-SA", !sensor == "TNHA-SB") 
aggregate(list(Value = avgdailysolar$value), 
          list(dailyavg = cut(avgdailysolar$Date, "1 day")), 
          mean)


```
AVG DAILY SOLAR FIGURE FOR SELECT SENSOR (BASED ON WHATS IN CHUNK ABOVE)
```{r fig10, fig.height = 6, fig.width = 16}

avgdailysolar %>%
  filter(!value == 0) %>%
  aggregate(value ~ Date + sensor, mean) %>%
  ggplot() +
  geom_line(aes(x=Date, y=value, color = sensor), size = 0.9) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(y="W /"~m^2~"",color="Station") +
  scale_x_datetime(date_breaks = "1 weeks") +
  scale_y_continuous(limits=c(0, 1000)) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E"))

```
MAX MIN MEANS FOR SOLAR RAD:  variable & specified station
```{r}
library(tibbletime)
solarrad2 <- solarrad %>% subset(station == c("SSMH")) %>% select(value,sensor,Date) %>% filter(!value == "999")

solarrad2 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)


solarrad2 <- solarrad %>%
  group_by(hour = date_group(Date,"hour"), day = date_group(Date, "day"),sensor) %>%
  summarize(mean_solar = mean(value),value)


solarrad2$hour_only <- format(solarrad2$hour, "%H")
solarrad2$day_only <- format(solarrad2$hour, "%m/%d/%y")


solarrad2$hour <- NULL



solarrad2_1 <- ddply(solarrad2, 'sensor', summarize, value =mean(value), Date=head(solarrad$Date,1), Value=head(solarrad$value,1))


solarrad23 <- solarrad2 %>%
  group_by(sensor) %>%
  mutate(
    Maxval = max(value, na.rm = T),
    Minval = min(value, na.rm = T)
  ) %>%
  arrange(sensor)
```
# Averaging data by hour per day for sensor over 6/19-6/24 for TNHA
```{r}
# Filter for the specified sensor and date range
TNHA_BASE_solar <- solarrad %>%
  filter(sensor == "TNHA-BASE" & 
         Date >= "2022-06-19 00:00:00" & 
         Date <= "2022-06-24 23:55:00")

```

# averging data for TNHA (all sensors) by hour per day 
```{r}
TNHA_avgsolar <- solarrad %>%
  group_by(hour = date_group(Date,"hour"), day = date_group(Date, "day"),sensor) %>%
  summarize(mean_solar = mean(value)) #this is the average solar radiation for every hour of every day in the date range

# adding columns that have just the hour and just the day,month, and year
TNHA_avgsolar$hour_only <- format(TNHA_avgsolar$hour, "%H")
TNHA_avgsolar$hour <- NULL
head(TNHA_avgsolar,24)
```

MAX MIN MEAN VALUES GROUND TEMP 
```{r}
df2_tnha <- dfgrndtemp %>% filter(!station == "BUECI", !station == "BEO", !station == "SSMH", depth == "10cm")

tnhagrnd_maxmin <- df2_tnha %>% # at 10cm depths for TNHA what are each sensors maximum and minimum values
  group_by(sensor) %>%
  mutate(
    Maxval = max(value, na.rm = T),
    Minval = min(value, na.rm = T),
    mean_val = mean(value, na.rm = T)
  ) %>%
  arrange(sensor)

tnhagrnd_maxmin %>% distinct(sensor,Maxval,Minval)
```


#focus on ground temperature averages then air temp

AIR temp averages (TNHA - summer)
```{r fig111, fig.height = 8, fig.width = 18}
Daily_average_airtmp_per_sensor %>% 
  ggplot() +
  geom_line(aes(x=day, y=value, color = sensor), size = 0.9) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(y="Temperature (°C)",x="Day",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" ="#E7298A", "TNHA-SD"="#66A61E"))


```

```{r}
airtemp_hourly <- airtemp %>%
  group_by(sensor, hour = floor_date(Date, "hour"), day = floor_date(Date, "day")) %>%
  summarize(avg_air_temp = mean(value))


grndtmp_hourly <- dfgrndtemp  %>%
  group_by(sensor, depth, hour = floor_date(Date, "hour"), day = floor_date(Date, "day")) %>%
  summarize(avg_ground_temp = mean(value))
```
Daily average grnd temperature per sensor at station "X"
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "10cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season

```




```{r fig13, fig.height = 8, fig.width = 18}
Daily_average_grnd_per_sensor %>% filter(!sensor =="SSMH-BASE", !sensor =="SSMH-SA", !sensor =="SSMH-SB", !sensor =="SSMH-SC", !sensor =="SSMH-SD", !sensor =="BUECI-BASE", !sensor=="BUECI-SA", !sensor =="BUECI-SB", !sensor =="BUECI-SC", !sensor =="BUECI-SD", !sensor == "BUECI-SE", !sensor =="BUECI-SF.01",!sensor =="BUECI-SF.02") %>%
  ggplot() +
  geom_line(aes(x=day, y=value, color = sensor), size = 0.9) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(y="Temperature (°C)",x="Day",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-2.5, 15)) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
   theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 30),plot.title = element_text(hjust = 0.5, size = 30), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 20),
        axis.title = element_text(size = 20)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" ="#E7298A", "TNHA-SD"="#66A61E","BEO-B05"="#6BAED6", "BEO-B06"= "#084594"))

```
```{r fig13, fig.height = 8, fig.width = 18}
Daily_average_grnd_per_sensor %>% filter(!sensor =="SSMH-BASE", !sensor =="SSMH-SA", !sensor =="SSMH-SB", !sensor =="SSMH-SC", !sensor =="SSMH-SD", !sensor =="BUECI-BASE", !sensor=="BUECI-SA", !sensor =="BUECI-SB", !sensor =="BUECI-SC", !sensor =="BUECI-SD", !sensor == "BUECI-SE", !sensor =="BUECI-SF.01",!sensor =="BUECI-SF.02", !sensor == "TNHA-BASE", !sensor == "TNHA-SD", !sensor == "TNHA-SB", !sensor == "BEO-B05") %>%
  ggplot() +
  geom_line(aes(x=day, y=value, color = sensor), size = 0.9) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(y="Temperature (°C)",x="Day",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-2.5, 15)) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
   theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 30),plot.title = element_text(hjust = 0.5, size = 30), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 20),
        axis.title = element_text(size = 20)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" ="#E7298A", "TNHA-SD"="#66A61E","BEO-B05"="#6BAED6", "BEO-B06"= "#084594"))

```


```{r fig13, fig.height = 8, fig.width = 18}
Daily_average_grnd_per_sensor %>% filter( !sensor =="BUECI-BASE", !sensor=="BUECI-SA", !sensor =="BUECI-SB", !sensor =="BUECI-SC", !sensor =="BUECI-SD", !sensor == "BUECI-SE", !sensor =="BUECI-SF.01",!sensor =="BUECI-SF.02", !sensor == "TNHA-BASE", !sensor == "TNHA-SA", !sensor == "TNHA-SC", !sensor == "TNHA-SD", !sensor == "TNHA-SB", !sensor == "BEO-B05") %>%
  ggplot() +
  geom_line(aes(x=day, y=value, color = sensor), size = 0.9) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(y="Temperature (°C)",x="Day",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-2.5, 15)) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
   theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 30),plot.title = element_text(hjust = 0.5, size = 30), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 20),
        axis.title = element_text(size = 20)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" ="#E7298A", "TNHA-SD"="#66A61E","BEO-B05"="#6BAED6", "BEO-B06"= "#084594","SSMH-BASE"="#807DBA", "SSMH-SA"="#A6761D", "SSMH-SB"= "#4292C6", "SSMH-SC"= "#DD3497", "SSMH-SD"="#7A0177"))

```

---------------------


Fall Data


Ground Moisture for all sites data prepping
```{r}
moisture=read.csv("C:/Users/qwe2qh/OneDrive - University of Virginia/UVA-OfficeComputer/Utqiagvik/Meteorological_Seasons_Data/Fall_VWC_2023_04_06_12_35_08_UTC_1.csv")
names(moisture) <- sub("...m.3.m.3..RX3000_BRW1", "", names(moisture))
names(moisture) <- sub("...m.3.m.3..RX3000_BRW4", "", names(moisture))
names(moisture) <- sub("...m.3.m.3..RX3000_BRW5", "", names(moisture))
names(moisture) <- sub("...m.3.m.3..RX3000_BRW6", "", names(moisture))
names(moisture) <- sub("..RXW.GP6.", "", names(moisture))
moisture$Date <-as.POSIXct(moisture$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

Solar radiation for all sites data prepping
```{r}
solar = read.csv("C:/Users/qwe2qh/OneDrive - University of Virginia/UVA-OfficeComputer/Utqiagvik/Meteorological_Seasons_Data/Fall_solar_2023_04_06_13_09_45_UTC_1.csv")
names(solar) <- sub("olar.Radiation..RXW.LIB", "" ,names(solar))
names(solar) <- sub("olar.Radiation..S.LIB", "" ,names(solar))
names(solar) <- sub("S.", "" ,names(solar))
names(solar) <- sub("...W.m.2..RX3000_BRW1", "" ,names(solar))
names(solar) <- sub("...W.m.2..RX3000_BRW4", "" ,names(solar))
names(solar) <- sub("...W.m.2..RX3000_BRW5", "" ,names(solar))
names(solar) <- sub("...W.m.2..RX3000_BRW6", "" ,names(solar))
solar$Date <- as.POSIXct(solar$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

Air temp for all sites data prepping
```{r}
airtemp =read.csv("C:/Users/qwe2qh/OneDrive - University of Virginia/UVA-OfficeComputer/Utqiagvik/Meteorological_Seasons_Data/Fall_airtemp_2023_04_06_12_01_15_UTC_1.csv")
names(airtemp) <- sub("Temperature..RXW.THC.", "" ,names(airtemp))
names(airtemp) <- sub("Temperature..S.THC.", "" ,names(airtemp))
names(airtemp) <- sub("S.", "" ,names(airtemp))
names(airtemp) <- sub("....C..RX3000_BRW1", "" ,names(airtemp))
names(airtemp) <- sub("....C..RX3000_BRW4", "" ,names(airtemp))
names(airtemp) <- sub("....C..RX3000_BRW5", "" ,names(airtemp))
names(airtemp) <- sub("....C..RX3000_BRW6", "" ,names(airtemp))
airtemp$Date <- as.POSIXct(airtemp$Date,format="%m/%d/%y %H:%M",tz="UTC")
```


Ground Temperatures for all sites data prepping
```{r}
# importing ground temp data for 6-19 through 9-20
grndtmp921_1119=read.csv("C:/Users/qwe2qh/OneDrive - University of Virginia/UVA-OfficeComputer/Utqiagvik/Meteorological_Seasons_Data/Fall_grndtmp_2023_04_06_12_19_52_UTC_1.csv")
names(grndtmp921_1119) <- sub("....C..RX3000_BRW1", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub("....C..RX3000_BRW4", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub("....C..RX3000_BRW5", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub("....C..RX3000_BRW6", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub(".RXW.GP6.", "", names(grndtmp921_1119))
grndtmp921_1119$Date <-as.POSIXct(grndtmp921_1119$Date,format="%m/%d/%y %H:%M",tz="UTC")
```


## Return Sensor ID of Each Sensor by Placement Number
NOTE: the sensor IDs do not match up for each serial number - this is an attempt
to identify inconsistencies.
**Not necessary to run, just to understand how many sensors there are!
```{r}
c = 3
sensorID = NULL
columnNum = NULL
while(c <= ncol(grndtmp921_1119)){
  sensorID = append(sensorID, names(grndtmp921_1119)[c])
  columnNum = append(columnNum, c)
  c = c + 11
}


## ADJUST LOCATION AND ID ##
sensors = data.frame(columnNum, sensorID)

sensors = sensors %>% separate(sensorID, c("Datatype", "Station", "Sensor", "Depth"), "\\.")

sensors$Station <- revalue(sensors$Station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO"))



#Find Sensor Values

sensors$Sensor <- revalue(sensors$Sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))

```

Removing ground temp zeroes occuring at multiple (3) consecutive depths for loop
```{r}
c = 3 
zeroes = 0
while(c <= ncol(grndtmp921_1119)-10){
  sensorGroup = grndtmp921_1119 %>% select(2,c, c+1, c+2, c+3, c+4, c+5, c+6, c+7, c+8, c+9, c+10)
  for(r in 1:nrow(sensorGroup)) {
    zeroes = 0
    col = 2
    while(col <= 12){
      if(sensorGroup[r,col] == 0 | is.na(sensorGroup[r,col])) { 
        if(!is.na(sensorGroup[r,col])){
          zeroes = zeroes + 1 
        }
      }
      col = col + 1
    }
    if(zeroes >= 3){ 
      for(columnn in 2:ncol(sensorGroup)){
      if(sensorGroup[r,columnn] == 0 | is.na(sensorGroup[r,columnn])) { 
        grndtmp921_1119[r,columnn+c] = NA
      }
    }
    }
  }
  c = c + 11 
}
## Depending on the loop in the chunk above, this may need to be adjusted too ##
```


Wide to long df conversion : Clean up ground moisture data 
```{r}
#pivot longer vs gather
dataframelong = gather(moisture, variable, response, Water.Content21198259.21206939.1:Water.Content21401800.21398576.6) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

df = cSplit(dataframelong, 'variable', sep=".", direction = "wide") 
colnames(df)[5:7] <- c("station", "sensor", "depth")
df$variable_1 <- NULL 
df$depth <- as.factor(as.character(df$depth)) 
df$station <- as.factor(df$station)

df$sensor <- as.factor(as.character(df$sensor))

df$station <- revalue(df$station, c("Content21198259" = "TNHA", "Content21401800" = "BUECI", "Content21401801" = "SSMH", "Content21401803" = "BEO")) 
df$depth <- revalue(df$depth, c("1" = "0:15cm", "2" = "15:30cm", "3" = "30:45cm", "4" = "45:60cm", "5" = "60:75cm", "6" = "75:90cm"))

df$sensor <- revalue(df$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398587" = "TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05")) 
```

GROUND TEMPERATURE CSV ADJUSTMENTS, LONG DATAFRAME
```{r}
grndtmp = gather(grndtmp921_1119, variable, response, Temperature.21198259.21206939.7:Temperature.21401800.21398576.17) %>%
  mutate(variable = fct_inorder(variable))
dfgrndtemp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(dfgrndtemp)[3] <- ("value")
colnames(dfgrndtemp)[5:7] <- c("station", "sensor", "depth")

dfgrndtemp$variable_1 <- NULL
dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth))
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))
dfgrndtemp$station <- as.factor(dfgrndtemp$station)
dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398587" = "TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))

dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))

```

SOLAR RADIATION : LONG DATAFRAME
```{r fig9, fig.heigt = 14, fig.width = 18}
solarrad = gather(solar, variable, response, '21198259.21176526.1':'21401800.21398623.1') %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

solarrad= cSplit(solarrad, 'variable', sep=".", direction = "wide") # texttocols, 
solarrad$variable_3 <- NULL
colnames(solarrad)[3:5] <- c("value", "station", "sensor")
solarrad$value <- as.numeric(as.character(solarrad$value)) 
solarrad$sensor <- as.factor(as.character(solarrad$sensor))
solarrad$station <- as.factor(solarrad$station)

solarrad$station <- revalue(solarrad$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO"))
solarrad$sensor <- revalue(solarrad$sensor, c("21390411" = "BUECI-BASE", "21398618" = "BUECI-SA", "21398624" = "BUECI-SB", "21362313" = "BUECI-SC", "21362316" = "BUECI-SD", "21362320" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", 
                                              "21390413" ="SSMH-BASE", "21398622" = "SSMH-SA", "21362319" = "SSMH-SB", "21166008" = "SSMH-SC", "21393049" = "SSMH-SD" ,
                                              "21176526" = "TNHA-BASE", "21398620" = "TNHA-SA", "21398616"="TNHA-SB","21362315" = "TNHA-SD", "21362317" = "TNHA-SC", 
                                              "21390415" = "BEO-Base"))

```

Air temperature data long dataframe
```{r}
airtemp = gather(airtemp, variable, response, "21198259.21398666.1":"21401801.21398670.1") %>%
  mutate(variable = fct_inorder(variable))

airtemp= cSplit(airtemp, 'variable', sep=".", direction = "wide") 
airtemp$variable_3 <- NULL
colnames(airtemp)[3:5] <- c("value", "station", "sensor")
airtemp$value <- as.numeric(as.character(airtemp$value)) 
airtemp$sensor <- as.factor(as.character(airtemp$sensor))
airtemp$station <- as.factor(airtemp$station)

airtemp$station <- revalue(airtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) 


airtemp$sensor <- revalue(airtemp$sensor, c("21390851" = "BUECI-BASE", "21398671" = "BUECI-SA", "21362254" = "BUECI-SB", "21398668" = "BUECI-SC", "21398660" = "BUECI-SD", "21187245" = "BUECI-SE",
                                            "21390849" ="SSMH-BASE","21393049" = "SSMH-SD", "21398670" = "SSMH-SA", "21398665" = "SSMH-SB",
                                            "21218018" = "TNHA-BASE", "21380919" = "TNHA-SA" , "21398676"="TNHA-SB","21398666" = "TNHA-SD", "21398674" = "TNHA-SC", "21397541" = "BEO-BASE"))
#TNHA-A and TNHA-B are GUESSES - need metadata on ground
airtemp <- airtemp %>% filter(!value == -888.88)
```
Daily average grnd temperature per sensor at station "X"
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "10cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season

#Below -- exports as a csv to folder of choice 

#write.csv(Daily_average_grnd_per_sensor, "~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer.csv")


```

```{r fig14, fig.height = 8, fig.width = 18}
Daily_average_grnd_per_sensor %>% filter(!sensor =="SSMH-BASE", !sensor =="SSMH-SA", !sensor =="SSMH-SB", !sensor =="SSMH-SC", !sensor =="SSMH-SD", !sensor =="BUECI-BASE", !sensor=="BUECI-SA", !sensor =="BUECI-SB", !sensor =="BUECI-SC", !sensor =="BUECI-SD", !sensor == "BUECI-SE", !sensor =="BUECI-SF.01",!sensor =="BUECI-SF.02") %>%
  ggplot() +
  geom_line(aes(x=day, y=value, color = sensor), size = 0.9) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(title = "Daily average ground temperatures at TNHA and BEO (10cm depth)",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-15,-12.5,-10,-7.5,-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-15, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 35),
        axis.title.x = element_text(size = 30),
        axis.title.y = element_text(size = 30)) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" ="#E7298A", "TNHA-SD"="#66A61E","BEO-B05"="#6BAED6", "BEO-B06"= "#084594"))
```


```{r}
averaged_by_site_TNHA <- Daily_average_grnd_per_sensor %>% filter(sensor == "TNHA-BASE" | sensor == "TNHA-SA" | sensor == "TNHA-SB"| sensor == "TNHA-SC" | sensor == "TNHA-SD") %>% aggregate(value~day, mean)
averaged_by_site_TNHA$station <- c("TNHA")

averaged_by_site_BEO <- Daily_average_grnd_per_sensor %>% filter(sensor == "BEO-B06") %>% aggregate(value~day, mean)
averaged_by_site_BEO$station <- c("BEO") 

averaged_by_site_SSMH <- Daily_average_grnd_per_sensor %>% filter(sensor == "SSMH-BASE" | sensor == "SSMH-SA" | sensor == "SSMH-SB"| sensor == "SSMH-SC" | sensor == "SSMH-SD") %>% aggregate(value~day, mean)
averaged_by_site_SSMH$station <- c("SSMH")

averaged_BEO_TNHA_SSMH <- rbind(averaged_by_site_BEO, averaged_by_site_TNHA, averaged_by_site_SSMH)
```

```{r fig14, fig.height = 8, fig.width = 18}
averaged_BEO_TNHA_SSMH %>%
ggplot() +
  geom_line(aes(x=day, y = value, colour = station)) +
  theme_bw() +
  guides(color = guide_legend(override.aes = list(size=10)))+
  labs(title = "Daily average ground temperature (10cm depth )",y="Temperature (°C)",x="Week",color="Station") +
   scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-2.5, 15)) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
   theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 30),plot.title = element_text(hjust = 0.5, size = 30), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 20),
        axis.title = element_text(size = 20)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic"))
```

