---
title: "Gap_Fill_test"
author: "Kiefer Fallin"
date: "3/16/2023"
output: html_document
---

# loading packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape)
library(naniar)
library(RColorBrewer)
library(clock)    #this is a new package that I downloaded to use the date_group() function (basically group_by() but for POSIXct vectors)
library(timechange)  #my lubridate package would not work unless I loaded this one as well
library(lubridate)
library(chillR)
library(MARSS)
library(KarsTS)
library(imputeTS)
library(zoo)
```

# importing dataset
```{r}
grndtmp619_920=read.csv("/Users/kieferfallin/Desktop/Research/Season_GRNDTMP_6_19_22_9_20_22_2022_09_21_16_19_26_UTC_1.csv")
names(grndtmp619_920) <- sub("....C..RX3000_BRW1", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW4", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW5", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW6", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub(".RXW.GP6.", "", names(grndtmp619_920))
grndtmp619_920$Date <-as.POSIXct(grndtmp619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
grndtmp619_920 = grndtmp619_920[,-14]


airtemp619_920=read.csv("/Users/kieferfallin/Desktop/Research/Season_AIRTMP_6_19_22_9_20_22_2022_09_21_16_21_48_UTC_1.csv")

names(airtemp619_920) <- sub("Temperature..RXW.THC.", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("Temperature..S.THC.", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("S.", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW1", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW4", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW5", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW6", "" ,names(airtemp619_920))
airtemp619_920$Date <- as.POSIXct(airtemp619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

```{r}
c = 3 
zeroes = 0
while(c <= ncol(grndtmp619_920)-10){
  sensorGroup = grndtmp619_920 %>% select(2,c, c+1, c+2, c+3, c+4, c+5, c+6, c+7, c+8, c+9, c+10)
  for(r in 1:nrow(sensorGroup)) {
    zeroes = 0
    col = 2
    while(col <= 12){
      if(sensorGroup[r,col] == 0 | is.na(sensorGroup[r,col])) { 
        if(!is.na(sensorGroup[r,col])){
          zeroes = zeroes + 1 
        }
      }
      col = col + 1
    }
    if(zeroes >= 3){ 
      for(columnn in 2:ncol(sensorGroup)){
      if(sensorGroup[r,columnn] == 0 | is.na(sensorGroup[r,columnn])) { 
        grndtmp619_920[r,columnn+c] = NA
      }
    }
    }
  }
  c = c + 11 
}
```

# Looking where NAs are in TNHA Base air temp record

```{r}
TNHA_Base_airtemp = airtemp619_920%>%
  select(Date, "21198259.21218018.1")
```

```{r}
num_nas <- sum(is.na(TNHA_Base_airtemp$"21198259.21218018.1"))

cat("Number of NAs in my_column:", num_nas)
```

```{r}
which(is.na(TNHA_Base_airtemp$"21198259.21218018.1")) #I wanted to make this code work so that the positions of NAs were listed in reference to the Date vector; however, I was unable to figure how to do so.
#we have a short, continuous gap on 2022-06-30 17:35:00; also have a patchy gap starting on 7146 2022-09-17 02:10:00
#the output is the rows where we can find the NAs
```


# Long dataframe conversion
```{r}
grndtmp = gather(grndtmp619_920, variable, response, Temperature.21198259.21206939.7:Temperature.21401803.21398591.17) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
dfgrndtemp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(dfgrndtemp)[3] <- ("value")
colnames(dfgrndtemp)[5:7] <- c("station", "sensor", "depth")
dfgrndtemp$V212 <- NULL
dfgrndtemp$V213 <- NULL
dfgrndtemp$variable_1 <- NULL
dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth))
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))
dfgrndtemp$station <- as.factor(dfgrndtemp$station)
dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUCEI-SA", "21398583" = "BUECI-SB", "21393042" = "BUCEI-SC", "21398584" = "BUCEI-SD", "21398579" = "BUCEI-SE", "21398578" = "BUCEI-SF.01", "21398598" = "BUCEI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))
dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))
dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))


airtemp = gather(airtemp619_920, variable, response, "21198259.21380919.1":"21401803.21397541.1") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
airtemp= cSplit(airtemp, 'variable', sep=".", direction = "wide") 
airtemp$variable_3 <- NULL
colnames(airtemp)[3:5] <- c("value", "station", "sensor")
airtemp$value <- as.numeric(as.character(airtemp$value)) 
airtemp$sensor <- as.factor(as.character(airtemp$sensor))
airtemp$station <- as.factor(airtemp$station)
airtemp$station <- revalue(airtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) 
airtemp$sensor <- revalue(airtemp$sensor, c("21390851" = "BUECI-BASE", "21398671" = "BUCEI-SA", "21362254" = "BUECI-SB", "21398668" = "BUCEI-SC", "21398660" = "BUCEI-SD", "21187245" = "BUCEI-SE",
                                            "21390849" ="SSMH-BASE","21393049" = "SSMH-SD", "21398670" = "SSMH-SA", "21398665" = "SSMH-SB",
                                            "21218018" = "TNHA-BASE", "21380919" = "TNHA-SA" , "21398676"="TNHA-SB","21398666" = "TNHA-SD", "21398674" = "TNHA-SC", "21397541" = "BEO-BASE"))
#TNHA-A and TNHA-B are GUESSES - need metadata on ground

```


# Experimenting with chillR package

# Trying to get code to fit; this needs to be run prior to chillR interpolate_gaps_hourly() function
```{r}
airtemp_hourly <- airtemp %>%
  group_by(sensor, hour = floor_date(Date, "hour"), day = floor_date(Date, "day")) %>%
  summarize(avg_VWC = mean(value))
#Since this function is based on hourly data, we cannot have the more refined interval resolution that the sensors take measurements at. So, we need the average hourly data. 
```

```{r}
airtemp <- airtemp %>%
  mutate(Year = year(Date),
         Month = month(Date),
         Day = day(Date),
         Hour = hour(Date))

airtemp_hourly_avg <- airtemp %>%
  group_by(Year, Month, Day, Hour,sensor) %>%
  summarize(Temp = mean(value))


#This creates all of the columns we need to run interpolate_gaps_hourly. It's important to note that the dataset in the example for the chillR resource you gave me has Tmin and Tmax columns as well. I am struggling with how to reckon with this aspect.


```

```{r}
interp = interpolate_gaps_hourly(airtemp_hourly_avg) #running this with original airtemp dataset does not work; instead, you have to put into a specific format with the appropriate columns -> the next couple chunks are spent putting the dataset into the proper format. I am a little confused by the yielded datasets from this function. Does this provide us with anything useful? Do we need to manipulate the dataset further to make the functions work with chillR? 

# Looking where we had gaps in original code, we have a value here now
interp_TNHA_BASE=interp[["weather"]]%>%
  filter(sensor=="TNHA-BASE")%>%
  select(-Tmin_source, -Tmax_source) #getting rid of these since our dataset did not have this
```

# Comparing interpolated data at NAs and elsewhere with original hourly averages

```{r}
TNHA_air_temp_hourly = airtemp_hourly_avg%>%
  filter(sensor=="TNHA-BASE")
# Comparing TNHA_air_temp_hourly with interp_TNHA_BASE, we immediately see that the interpolation changes every single value for temperature, not just the NA ones
```


# Trying a different package: imputeTS
```{r}
temp = TNHA_Base_airtemp%>%
  select("21198259.21218018.1")
ggplot_na_distribution(temp) #This is a useful function for visualizing gaps in the data, although I am not sure how to get the X axis to display according to the "Date" vector
```

```{r}
ts_data <- zoo::zoo(TNHA_Base_airtemp$"21198259.21218018.1", TNHA_Base_airtemp$Date)
ts_data_imputed <- na_interpolation(ts_data) #interpolating data
```

```{r}
# Extract the interpolated values along with their corresponding dates
imputed_values <- coredata(ts_data_imputed)
dates <- index(ts_data_imputed)

# Create a new data frame with all the data
TNHA_base_interpolated <- data.frame(Date = dates, "21198259.21218018.1" = imputed_values)

# View the entire data frame with interpolated values
View(TNHA_base_interpolated)

#Comparing this data with the original airtemp619_920 data, we see that there are very realistic values created for our gaps from before (see above for locations according to row number). This is, of course, a sensor with relatively small gaps, we would have to explore other methods to interpolate for longer gaps. 
```


```{r}
ggplot_na_distribution(ts_data_imputed) #I did the same function as before to visualize gaps, but this time with the newly interpolated data using the na_interpolation() function. Here, we see that the graphs are identical, except for the fact that the red bars representing where missing data is are gone. This would seem to mean they were interpolated; however, I am not sure how to access the zoo series created during interpolation to look at the values it came up with.
```

