---
title: "Check for Gaps in Data"
author: "Elizabeth Van Metre"
date: '2023-01-26'
output: html_document
---


##Change this file path to read a different file
```{r}
data_path = "~/Desktop/Arctic/Season_GRNDTMP_6_19_22_9_20_22_2022_09_21_16_19_26_UTC_1.csv"
```


##Loading packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(splitstackshape) # basically like text to columns
library(naniar)
library(knitr)
library(writexl)
```

#Load and clean data
```{r}
df =read.csv(data_path)

names(df) <- sub("Temperature..RXW.THC.", "" ,names(df))
names(df) <- sub("Temperature..S.THC.", "" ,names(df))
names(df) <- sub("Temperature..RXW.GP6.", "" ,names(df))
names(df) <- sub("S.", "" ,names(df))
names(df) <- sub("....C..RX3000_BRW1", "" ,names(df))
names(df) <- sub("....C..RX3000_BRW4", "" ,names(df))
names(df) <- sub("....C..RX3000_BRW5", "" ,names(df))
names(df) <- sub("....C..RX3000_BRW6", "" ,names(df))
df$Date <- as.POSIXct(df$Date,format="%m/%d/%y %H:%M",tz="UTC")

#THESE ARE BLANK COLUMNS - rewrite if using a different data set
df = df[,-c(26:36)]
df = df[,-c(92:201)]

```

Store sensor names for final product
```{r}
sensor_names = colnames(df)[3:ncol(df)]
```

Rename sensors as 1, 2, 3... for easier sorting
```{r}
colnames(df)[3:ncol(df)] = seq(1, ncol(df)-2, 1)
```

Create long data frame of all sensor data
```{r}
df_long = gather(df, sensor, data, 3:ncol(df))
```

#Create list of gaps
```{r}
gap_list = data.frame(sensor_name=NA, date_start=NA, date_end=NA, gap_date=NA)[numeric(0), ]

for(sen in 1:ncol(df)) {
  sensor = df_long %>% filter(sensor == sen)
  
  r = 1
  
  if(!is.na(sensor[1,4])){
    date_start = sensor[1,2]
    date_end = NULL
    gap_date = NULL
  }
  else{
    date_start = NA
    date_end = NA
    gap_date = sensor[1,2]
  }
  
  while(r <= nrow(sensor)){
    while((!is.na(sensor[r,4]) | (is.na(sensor[r,4]) & !is.na(sensor[r+1, 4]))) & r <= nrow(sensor)){
      r = r + 1
    }
    date_end = append(date_end, sensor[r-1,2])
    gap_date = append(gap_date, sensor[r, 2])
    while(is.na(sensor[r,4]) & r <= nrow(sensor)){
      r = r + 1
    }
    date_start = append(date_start, sensor[r, 2])
     
  }
  if(length(date_start) > length(date_end)){
    date_end = append(date_end, sensor[r-1,2])
  }
  # if(length(date_start) < length(date_end)){
  #   date_start = append(date_start, sensor[r-1,2])
  # }
  if(length(gap_date) < length(date_start)){
    gap_date = append(gap_date, NA)
  }
  
  sensor_name = rep(sen, length(date_start))
  current_gap = data.frame(sensor_name, date_start, date_end, gap_date)
  gap_list = rbind(gap_list, current_gap)
}

```

Rewrite sensor IDs
```{r}
for(i in 1:length(sensor_names)){
  gap_list$sensor_name[which(gap_list$sensor_name == i)] <- sensor_names[i]
}
```
Create copy of sensor IDs
```{r}
gap_list = data.frame(gap_list, gap_list[,1])
```
Split IDs to station, sensor, and depth
```{r}
gap_list= cSplit(gap_list, 5, sep=".", direction = "wide") 
```
Organize and Reassign Names for Readability
```{r}
colnames(gap_list)[5:7] <- c("station", "sensor", "depth")
gap_list$station <- as.character(gap_list$station)
gap_list$sensor <- as.character(gap_list$sensor)

gap_list$station <- revalue(gap_list$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO"))
```
Reorganize column order
```{r}
gap_list = gap_list[,c(1,5,6,7,2,3,4)]
```

##Export Data Frame to Excel File
```{r}
write_xlsx(gap_list,"~/Desktop/Arctic/gap-list.xlsx")
```

