---
title: "Differences_in_averages"
author: "Mirella Shaban"
date: "2023-10-23"
output: html_document

############################################################
THIS SCRIPT GIVES YOU THE SEASONAL AVERAGES 
1) BY SITE & 2) BY ASPECT AT THE SAME SITE
*RUN LOADANDCLEAN_DIFFERENCES_IN_AVERAGES.RMD BEFORE THIS SCRIPT TO LOAD IN THE CLEANED DATA FOR EACH SEASON *
############################################################
---
*Names of seasonal data files (cleaned)*
SUMMERGRND22_ALLDEPTHS
FALL22_ALLDEPTHS
SPRING23_ALLDEPTHS
SUMMERGRND23_ALLDEPTHS
FALL23_ALLDEPTHS *use the BASE STATION as the SOUTH SENSOR here Because SA HAS NO DATA*


DF for HOURLY / HOUR comparisons between stations: CHANGE NAME OF DF YOU WANT --> FROM RAW DATA
```{r}
grndtmp_hourly <- summer22grndtmp  %>%
  group_by(sensor, depth, hour = floor_date(Date, "hour"), day = floor_date(Date, "day")) %>%
  summarize(avg_ground_temp = mean(value))


NS_Hourly <- grndtmp_hourly %>% filter(sensor == "TNHA-SC" | sensor == "TNHA-SA" | sensor == "SSMH-SA" | sensor == "SSMH-SB")
NS_Hourly <- NS_Hourly %>% filter(depth == "10cm")

Hourly_Comparisons <- NS_Hourly %>%
    pivot_wider(names_from = sensor, values_from = avg_ground_temp)

Hourly_Comparisons$TNHA_NS <- Hourly_Comparisons$`TNHA-SA` - Hourly_Comparisons$`TNHA-SC`
Hourly_Comparisons$SSMH_NS <- Hourly_Comparisons$`SSMH-SA` - Hourly_Comparisons$`SSMH-SB`


#below, gives row with max value of interest
Hourly_Comparisons[which.max(Hourly_Comparisons$SSMH_NS), ]

```



STATS: average daily difference between SITES (all stations combined) for Xcm depth (define in x <- )
```{r}
wide <- SUMMERGRND22_ALLDEPTHS %>%
    pivot_wider(names_from = station, values_from = value)

wide$TBDIFF <- wide$TNHA - wide$BEO
wide$SBDIFF <- wide$SSMH - wide$BEO
wide$TBSBDIFF <- wide$SSMH - wide$TNHA

x<- wide %>% filter(depth == "10cm") 
mean(x$TBDIFF,na.rm = TRUE)
mean(x$SBDIFF, na.rm = TRUE)
mean(x$TBSBDIFF, na.rm = TRUE)
```


####
Need to add "average by station" to the load and clean rmd file, or switch all sites averaging to this script

#####

AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-BASE") %>% aggregate(value~date+depth, mean)
averaged_by_station$fullname <- c("TNHA-BASE")

averaged_by_station1 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station1$fullname <- c("TNHA-SA")

#averaged_by_station2 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-SB") %>% aggregate(value~date+depth, mean)
#averaged_by_station2$fullname <- c("TNHA-SB")

averaged_by_station3 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date+depth, mean)
averaged_by_station3$fullname <- c("TNHA-SC")


averaged_by_station4 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-SD") %>% aggregate(value~date+depth, mean)
averaged_by_station4$fullname <- c("TNHA-SD")

averaged_by_station5 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "SSMH-BASE") %>% aggregate(value~date+depth, mean)
averaged_by_station5$fullname <- c("SSMH-BASE")

averaged_by_station6 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "SSMH-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station6$fullname <- c("SSMH-SA")

averaged_by_station7 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "SSMH-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station7$fullname <- c("SSMH-SB")


averaged_by_station8 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "SSMH-SD") %>% aggregate(value~date+depth, mean)
averaged_by_station8$fullname <- c("SSMH-SD")


averaged_by_station9 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06" | fullname == "BEO-B07" | fullname == "BEO-B08") %>% aggregate(value~date+depth, mean)
averaged_by_station9$fullname <- c("BEO")




sum22_STATIONS_AVG <- rbind(averaged_by_station, averaged_by_station1, averaged_by_station3,averaged_by_station4,averaged_by_station5,averaged_by_station6,averaged_by_station7,averaged_by_station8,averaged_by_station9)

#sum22_STATIONS_AVG$date <- as.Date(sum22_STATIONS_AVG$date, format= "%Y-%m-%d",tz="UTC")
```
STATS: Compute seasonal average of daily value over for one location/station
```{r}
wide <- sum22_STATIONS_AVG %>%
    pivot_wider(names_from = fullname, values_from = value)

x<- wide %>% filter(depth == "90cm") 

round(mean(x$`BEO`, na.rm=TRUE),2)
```



STATS: average daily difference between SPECIFIC STATIONS for Xcm depth (define in x <- )
```{r}
wide <- sum22_STATIONS_AVG %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO`

wide$SSMH_NS_DIFF <- wide$`SSMH-SA` - wide$`SSMH-SB`
wide$SSMH_N_BEO_DIFF <- wide$`SSMH-SB` - wide$`BEO`
wide$SSMH_S_BEO_DIFF <- wide$`SSMH-SA` - wide$`BEO`

wide$TNHA_SSMH_N_COMP <- wide$`TNHA-SC`-wide$`SSMH-SB`
wide$TNHA_SSMH_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SA`

#write.csv(wide, "~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_daily_average_difference_N_vs_S.csv")

x<- wide %>% filter(depth == "90cm") 

mean(x$TNHA_NS_DIFF,na.rm = TRUE)
mean(x$SSMH_NS_DIFF, na.rm = TRUE)


mean(x$TNHA_S_BEO_DIFF, na.rm = TRUE)
mean(x$SSMH_S_BEO_DIFF, na.rm = TRUE)

mean(x$TNHA_SSMH_N_COMP, na.rm = TRUE) #TNHA N - SSMH - S
mean(x$TNHA_N_BEO_DIFF, na.rm = TRUE) #TNHA-N minus BEO

max(x$TNHA_NS_DIFF, na.rm = TRUE)
max(x$SSMH_NS_DIFF, na.rm = TRUE)


#below, p1 = 6/19-8/01 , p2= 8/01-
#p1 = with(x, mean(TNHA_SSMH_N_COMP[date >= "2022-06-19" & date <= "2022-07-08"])) #this excludes NA values from avging


#x #this ignores NA values, any date with data will be factored in to the average --> USE THIS FOR THE BAR PLOTS, ABOVE FOR TABLES
#p1

```

FALL22
AVERAGE BY SITE
-----------
```{r}
averaged_by_site_TNHA <- FALL22_ALLDEPTHS %>% filter(fullname == "TNHA-BASE" | fullname == "TNHA-SA" | fullname == "TNHA-SB"| fullname == "TNHA-SC" | fullname == "TNHA-SD") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_TNHA$station <- c("TNHA")

averaged_by_site_BEO <- FALL22_ALLDEPTHS %>% filter(fullname == "BEO-B06" | fullname == "BEO-B05") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_BEO$station <- c("BEO") 

averaged_by_site_SSMH <- FALL22_ALLDEPTHS %>% filter(fullname == "SSMH-BASE" | fullname == "SSMH-SA" | fullname == "SSMH-SB"| fullname == "SSMH-SC" | fullname == "SSMH-SD") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_SSMH$station <- c("SSMH")

SITES_AVG_FALL22_GRND <- rbind(averaged_by_site_BEO, averaged_by_site_TNHA, averaged_by_site_SSMH)
```


STATS: average daily difference between SITES (all stations combined) for Xcm depth (define in x <- )
```{r}
wide <- SITES_AVG_FALL22_GRND %>%
    pivot_wider(names_from = station, values_from = value)

wide$TBDIFF <- wide$TNHA - wide$BEO
wide$SBDIFF <- wide$SSMH - wide$BEO
wide$TBSBDIFF <- wide$SSMH - wide$TNHA

x<- wide %>% filter(depth == "30cm") 
mean(x$TBDIFF,na.rm = TRUE)
mean(x$SBDIFF, na.rm = TRUE)
mean(x$TBSBDIFF, na.rm = TRUE)
```

AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- FALL22_ALLDEPTHS %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station$fullname <- c("TNHA-SA")
averaged_by_station2 <- FALL22_ALLDEPTHS %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date+depth, mean)
averaged_by_station2$fullname <- c("TNHA-SC")
averaged_by_station3 <- FALL22_ALLDEPTHS %>% filter(fullname == "SSMH-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station3$fullname <- c("SSMH-SA")
averaged_by_station4 <- FALL22_ALLDEPTHS %>% filter(fullname == "SSMH-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station4$fullname <- c("SSMH-SB")
averaged_by_station5 <- FALL22_ALLDEPTHS %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06") %>% aggregate(value~date+depth, mean)
averaged_by_station5$fullname <- c("BEO")
FALL22_STATIONS_AVG<- rbind(averaged_by_station, averaged_by_station2, averaged_by_station3,averaged_by_station4,averaged_by_station5)
```


STATS: average daily difference between SPECIFIC STATIONS for Xcm depth (define in x <- )
```{r}
wide <- FALL22_STATIONS_AVG %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO`
wide$SSMH_NS_DIFF <- wide$`SSMH-SA` - wide$`SSMH-SB`
wide$SSMH_N_BEO_DIFF <- wide$`SSMH-SB` - wide$`BEO`
wide$SSMH_S_BEO_DIFF <- wide$`SSMH-SA` - wide$`BEO`
wide$TNHA_SSMH_N_COMP <- wide$`TNHA-SC`-wide$`SSMH-SB`
wide$TNHA_SSMH_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SA`

#write.csv(wide, "~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_daily_average_difference_N_vs_S.csv")

x<- wide %>% filter(depth == "10cm") 

mean(x$TNHA_NS_DIFF,na.rm = TRUE)
mean(x$SSMH_NS_DIFF, na.rm = TRUE)
mean(x$TNHA_S_BEO_DIFF, na.rm = TRUE)
mean(x$SSMH_S_BEO_DIFF, na.rm = TRUE)

max(x$TNHA_NS_DIFF, na.rm = TRUE)
max(x$SSMH_NS_DIFF, na.rm = TRUE)

mean(x$TNHA_SSMH_N_COMP, na.rm = TRUE)



p2 = with(x, mean(TNHA_SSMH_N_COMP[date >= "2022-09-01" & date <= "2022-12-01"]))
#p1

```

--------

********

SPRING 23
--------
```{r}
averaged_by_site_TNHA <- SPRING23_ALLDEPTHS %>% filter(fullname == "TNHA-BASE" | fullname == "TNHA-SA" | fullname == "TNHA-SB"| fullname == "TNHA-SC" | fullname == "TNHA-SD") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_TNHA$station <- c("TNHA")

averaged_by_site_BEO <- SPRING23_ALLDEPTHS %>% filter(fullname == "BEO-B06" | fullname == "BEO-B05") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_BEO$station <- c("BEO") 

averaged_by_site_SSMH <- SPRING23_ALLDEPTHS %>% filter(fullname == "SSMH-BASE" | fullname == "SSMH-SA" | fullname == "SSMH-SB"| fullname == "SSMH-SC" | fullname == "SSMH-SD") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_SSMH$station <- c("SSMH")

SITES_AVG_SPRING23_GRND <- rbind(averaged_by_site_BEO, averaged_by_site_TNHA, averaged_by_site_SSMH)
```


STATS: average daily difference between SITES (all stations combined) for Xcm depth (define in x <- )
```{r}
wide <- SITES_AVG_SPRING23_GRND %>%
    pivot_wider(names_from = station, values_from = value)

wide$TBDIFF <- wide$TNHA - wide$BEO
wide$SBDIFF <- wide$SSMH - wide$BEO
wide$TBSBDIFF <- wide$SSMH - wide$TNHA

x<- wide %>% filter(depth == "10cm") 
mean(x$TBDIFF,na.rm = TRUE)
mean(x$SBDIFF, na.rm = TRUE)
mean(x$TBSBDIFF, na.rm = TRUE)
```

AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- SPRING23_ALLDEPTHS %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station$fullname <- c("TNHA-SA")
averaged_by_station2 <- SPRING23_ALLDEPTHS %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date+depth, mean)
averaged_by_station2$fullname <- c("TNHA-SC")
#averaged_by_station3 <- SPRING23_ALLDEPTHS %>% filter(fullname == "SSMH-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station3$fullname <- c("SSMH-SA")
#above, no rows to agg? check DF

averaged_by_station4 <- SPRING23_ALLDEPTHS %>% filter(fullname == "SSMH-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station4$fullname <- c("SSMH-SB")
averaged_by_station5 <- SPRING23_ALLDEPTHS %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06") %>% aggregate(value~date+depth, mean)
averaged_by_station5$fullname <- c("BEO")
SPRING23_STATIONS_AVG<- rbind(averaged_by_station, averaged_by_station2, averaged_by_station4,averaged_by_station5)
```


STATS: average daily difference between SPECIFIC STATIONS for Xcm depth (define in x <- )
```{r}
wide <- SPRING23_STATIONS_AVG %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO`

wide$SSMH_NS_DIFF <- wide$`SSMH-SA` - wide$`SSMH-SB`
wide$SSMH_N_BEO_DIFF <- wide$`SSMH-SB` - wide$`BEO`
wide$SSMH_S_BEO_DIFF <- wide$`SSMH-SA` - wide$`BEO`

wide$TNHA_SSMH_N_COMP <- wide$`TNHA-SC`-wide$`SSMH-SB`
wide$TNHA_SSMH_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SA`

wide$TNHA_SSMH_N_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SB`

#write.csv(wide, "~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_daily_average_difference_N_vs_S.csv")

x<- wide %>% filter(depth == "90cm") 

mean(x$TNHA_NS_DIFF,na.rm = TRUE)
mean(x$SSMH_NS_DIFF, na.rm = TRUE)
mean(x$TNHA_S_BEO_DIFF, na.rm = TRUE)
mean(x$SSMH_S_BEO_DIFF, na.rm = TRUE)
mean(x$TNHA_N_BEO_DIFF, na.rm = TRUE)
mean(x$TNHA_SSMH_N_COMP, na.rm = TRUE)
mean(x$TNHA_SSMH_N_S_COMP, na.rm = TRUE)

max(x$TNHA_NS_DIFF, na.rm = TRUE)
max(x$SSMH_NS_DIFF, na.rm = TRUE)


#below, p1 = 6/19-8/01 , p2= 8/01-
p3 = with(x, mean(TNHA_SSMH_N_COMP[date >= "2023-03-01" & date <= "2023-06-01"]))

#CHANGE DATA RANGE ABOVE

```

------

SUMMER 23

**
--------
```{r}
averaged_by_site_TNHA <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-BASE" | fullname == "TNHA-SA" | fullname == "TNHA-SB"| fullname == "TNHA-SC" | fullname == "TNHA-SD") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_TNHA$station <- c("TNHA")

averaged_by_site_BEO <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "BEO-B06" | fullname == "BEO-B05") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_BEO$station <- c("BEO") 

averaged_by_site_SSMH <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-BASE" | fullname == "SSMH-SA" | fullname == "SSMH-SB"| fullname == "SSMH-SC" | fullname == "SSMH-SD" | fullname == "SSMH-SE" | fullname == "SSMH-SF" | fullname == "SSMH-SG" | fullname == "SSMH-SH" | fullname == "SSMH-SI") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_SSMH$station <- c("SSMH")

SITES_AVG_SUMM23_GRND <- rbind(averaged_by_site_BEO, averaged_by_site_TNHA, averaged_by_site_SSMH)
```


STATS: average daily difference between SITES (all stations combined) for Xcm depth (define in x <- )
```{r}
wide <- SITES_AVG_SUMM23_GRND %>%
    pivot_wider(names_from = station, values_from = value)

wide$TBDIFF <- wide$TNHA - wide$BEO
wide$SBDIFF <- wide$SSMH - wide$BEO
wide$TBSBDIFF <- wide$SSMH - wide$TNHA

x<- wide %>% filter(depth == "10cm") 
mean(x$TBDIFF,na.rm = TRUE)
mean(x$SBDIFF, na.rm = TRUE)
mean(x$TBSBDIFF, na.rm = TRUE)
```

AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station$fullname <- c("TNHA-SA")
averaged_by_station2 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date+depth, mean)
averaged_by_station2$fullname <- c("TNHA-SC")

averaged_by_station3 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station3$fullname <- c("SSMH-SA")
averaged_by_station4 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station4$fullname <- c("SSMH-SB")
averaged_by_station5 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06") %>% aggregate(value~date+depth, mean)
averaged_by_station5$fullname <- c("BEO")

#averaged_by_station5_NEW <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "BEO-B07" | fullname == "BEO-B08") %>% aggregate(value~date+depth, mean)
#averaged_by_station5_NEW$fullname <- c("BEO-NEW")

averaged_by_station6 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-BASE") %>% aggregate(value~date+depth, mean)
averaged_by_station6$fullname <- c("TNHA-BASE")

averaged_by_station7 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-SD") %>% aggregate(value~date+depth, mean)
averaged_by_station7$fullname <- c("TNHA-SD")

averaged_by_station8 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station8$fullname <- c("TNHA-SB")


#averaged_by_station8 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-BASE") %>% aggregate(value~date+depth, mean)
#averaged_by_station8$fullname <- c("SSMH-BASE")
#above, no rows to aggregate (no data)

averaged_by_station9 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SD") %>% aggregate(value~date+depth, mean)
averaged_by_station9$fullname <- c("SSMH-SD")


SUM23_STATIONS_AVG<- rbind(averaged_by_station, averaged_by_station2, averaged_by_station3,averaged_by_station4,averaged_by_station5,averaged_by_station6,averaged_by_station7,averaged_by_station9,averaged_by_station8)
```


STATS: average daily difference between SPECIFIC STATIONS for Xcm depth (define in x <- )
```{r}
wide <- SUM23_STATIONS_AVG %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO`

wide$SSMH_NS_DIFF <- wide$`SSMH-SA` - wide$`SSMH-SB`
wide$SSMH_N_BEO_DIFF <- wide$`SSMH-SB` - wide$`BEO`
wide$SSMH_S_BEO_DIFF <- wide$`SSMH-SA` - wide$`BEO`

wide$TNHA_SSMH_N_COMP <- wide$`TNHA-SC`-wide$`SSMH-SB`
wide$TNHA_SSMH_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SA`

wide$TNHA_SSMH_N_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SB`

#write.csv(wide, "~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_daily_average_difference_N_vs_S.csv")

x<- wide %>% filter(depth == "10cm") 

mean(x$TNHA_NS_DIFF,na.rm = TRUE)
mean(x$SSMH_NS_DIFF, na.rm = TRUE)
mean(x$TNHA_S_BEO_DIFF, na.rm = TRUE)
mean(x$SSMH_S_BEO_DIFF, na.rm = TRUE)

mean(x$TNHA_SSMH_N_COMP, na.rm = TRUE)
mean(x$TNHA_SSMH_N_S_COMP, na.rm = TRUE)
mean(x$SSMH_N_BEO_DIFF, na.rm = TRUE)

max(x$TNHA_NS_DIFF, na.rm = TRUE)
max(x$SSMH_NS_DIFF, na.rm = TRUE)

#below, p1 = 6/19-8/01 , p2= 8/01-
p3 = with(x, mean(TNHA_SSMH_N_COMP[date >= "2023-06-01" & date <= "2023-09-01"]))

#CHANGE DATA RANGE ABOVE

```


FALL 23
**

----- 
STATS: average daily difference between SITES (all stations combined) for Xcm depth (define in x <- )
```{r}
wide <- FALL23_ALLDEPTHS %>%
    pivot_wider(names_from = station, values_from = value)

wide$TBDIFF <- wide$TNHA - wide$BEO
wide$SBDIFF <- wide$SSMH - wide$BEO
wide$TBSBDIFF <- wide$SSMH - wide$TNHA

x<- wide %>% filter(depth == "10cm") 
mean(x$TBDIFF,na.rm = TRUE)
mean(x$SBDIFF, na.rm = TRUE)
mean(x$TBSBDIFF, na.rm = TRUE)
```

AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- FALL23_ALLDEPTHS %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station$fullname <- c("TNHA-SA")
averaged_by_station2 <- FALL23_ALLDEPTHS %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date+depth, mean)
averaged_by_station2$fullname <- c("TNHA-SC")

averaged_by_station3 <- FALL23_ALLDEPTHS %>% filter(fullname == "SSMH-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station3$fullname <- c("SSMH-SA")
averaged_by_station4 <- FALL23_ALLDEPTHS %>% filter(fullname == "SSMH-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station4$fullname <- c("SSMH-SB")
# averaged_by_station5 <- FALL23_ALLDEPTHS %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06") %>% aggregate(value~date+depth, mean)
# averaged_by_station5$fullname <- c("BEO")

averaged_by_station6 <- FALL23_ALLDEPTHS %>% filter(fullname == "TNHA-BASE") %>% aggregate(value~date+depth, mean)
averaged_by_station6$fullname <- c("TNHA-BASE")

# averaged_by_station7 <- FALL23_ALLDEPTHS %>% filter(fullname == "TNHA-SD") %>% aggregate(value~date+depth, mean)
# averaged_by_station7$fullname <- c("TNHA-SD")

averaged_by_station8 <- FALL23_ALLDEPTHS %>% filter(fullname == "TNHA-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station8$fullname <- c("TNHA-SB")


#averaged_by_station8 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-BASE") %>% aggregate(value~date+depth, mean)
#averaged_by_station8$fullname <- c("SSMH-BASE")
#above, no rows to aggregate (no data)

averaged_by_station10 <- FALL23_ALLDEPTHS %>% filter(fullname == "SSMH-SE") %>% aggregate(value~date+depth, mean)
averaged_by_station10$fullname <- c("SSMH-SE")

averaged_by_station11 <- FALL23_ALLDEPTHS %>% filter(fullname == "SSMH-SF") %>% aggregate(value~date+depth, mean)
averaged_by_station11$fullname <- c("SSMH-SF")

averaged_by_station12 <- FALL23_ALLDEPTHS %>% filter(fullname == "SSMH-SG") %>% aggregate(value~date+depth, mean)
averaged_by_station12$fullname <- c("SSMH-SG")

averaged_by_station13 <- FALL23_ALLDEPTHS %>% filter(fullname == "SSMH-SH") %>% aggregate(value~date+depth, mean)
averaged_by_station13$fullname <- c("SSMH-SH")

averaged_by_station14 <- FALL23_ALLDEPTHS %>% filter(fullname == "SSMH-SI") %>% aggregate(value~date+depth, mean)
averaged_by_station14$fullname <- c("SSMH-SI")


FALL23_STATIONS_AVG<- rbind(averaged_by_station, averaged_by_station2, averaged_by_station3,averaged_by_station4,averaged_by_station5,averaged_by_station6,averaged_by_station7,averaged_by_station9,averaged_by_station8,averaged_by_station11,averaged_by_station13,averaged_by_station14)
```


STATS: average daily difference between SPECIFIC STATIONS for Xcm depth (define in x <- )
```{r}
wide <- FALL23_STATIONS_AVG %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO`

wide$SSMH_NS_DIFF <- wide$`SSMH-SA` - wide$`SSMH-SB`
wide$SSMH_N_BEO_DIFF <- wide$`SSMH-SB` - wide$`BEO`
wide$SSMH_S_BEO_DIFF <- wide$`SSMH-SA` - wide$`BEO`

wide$TNHA_SSMH_N_COMP <- wide$`TNHA-SC`-wide$`SSMH-SB`
wide$TNHA_SSMH_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SA`
wide$TNHA_SSMH_N_S_COMP <- wide$`TNHA-SC`-wide$`SSMH-SA`
wide$TNHA_N_S_DIFF <- wide$`SSMH-SB`-wide$`TNHA-SC`

#write.csv(wide, "~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_daily_average_difference_N_vs_S.csv")

x<- wide %>% filter(depth == "90cm") 

mean(x$TNHA_NS_DIFF,na.rm = TRUE)
mean(x$SSMH_NS_DIFF, na.rm = TRUE)
mean(x$TNHA_S_BEO_DIFF, na.rm = TRUE)
mean(x$SSMH_S_BEO_DIFF, na.rm = TRUE)

mean(x$SSMH_N_BEO_DIFF, na.rm = TRUE)
mean(x$TNHA_SSMH_N_S_COMP, na.rm = TRUE)
mean(x$TNHA_N_S_DIFF, na.rm = TRUE)

max(x$TNHA_NS_DIFF, na.rm = TRUE)
max(x$SSMH_NS_DIFF, na.rm = TRUE)

#below, p1 = 6/19-8/01 , p2= 8/01-
p3 = with(x, mean(TNHA_SSMH_N_COMP[date >= "2023-06-01" & date <= "2023-09-01"]))

#CHANGE DATA RANGE ABOVE

```
x

AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station$fullname <- c("TNHA-SA")
averaged_by_station2 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date+depth, mean)
averaged_by_station2$fullname <- c("TNHA-SC")

averaged_by_station3 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station3$fullname <- c("SSMH-SA")
averaged_by_station4 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station4$fullname <- c("SSMH-SB")
averaged_by_station5 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06") %>% aggregate(value~date+depth, mean)
 averaged_by_station5$fullname <- c("BEO")

averaged_by_station6 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-BASE") %>% aggregate(value~date+depth, mean)
averaged_by_station6$fullname <- c("TNHA-BASE")

averaged_by_station7 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-SD") %>% aggregate(value~date+depth, mean)
averaged_by_station7$fullname <- c("TNHA-SD")

averaged_by_station8 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station8$fullname <- c("TNHA-SB")


averaged_by_station8 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-BASE") %>% aggregate(value~date+depth, mean)
averaged_by_station8$fullname <- c("SSMH-BASE")
#above, no rows to aggregate (no data)

averaged_by_station10 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SE") %>% aggregate(value~date+depth, mean)
averaged_by_station10$fullname <- c("SSMH-SE")

averaged_by_station11 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SF") %>% aggregate(value~date+depth, mean)
averaged_by_station11$fullname <- c("SSMH-SF")

averaged_by_station12 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SG") %>% aggregate(value~date+depth, mean)
averaged_by_station12$fullname <- c("SSMH-SG")

averaged_by_station13 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SH") %>% aggregate(value~date+depth, mean)
averaged_by_station13$fullname <- c("SSMH-SH")

averaged_by_station14 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "SSMH-SI") %>% aggregate(value~date+depth, mean)
averaged_by_station14$fullname <- c("SSMH-SI")


SUMMERGRND23_STATIONS_AVG<- rbind(averaged_by_station, averaged_by_station2, averaged_by_station3,averaged_by_station4,averaged_by_station5,averaged_by_station6,averaged_by_station7,averaged_by_station9,averaged_by_station8,averaged_by_station11,averaged_by_station13,averaged_by_station14)
```


STATS: average daily difference between SPECIFIC STATIONS for Xcm depth (define in x <- )
```{r}
wide <- SUMMERGRND23_STATIONS_AVG %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO`

wide$SSMH_NS_DIFF <- wide$`SSMH-SA` - wide$`SSMH-SB`
wide$SSMH_N_BEO_DIFF <- wide$`SSMH-SB` - wide$`BEO`
wide$SSMH_S_BEO_DIFF <- wide$`SSMH-SA` - wide$`BEO`

wide$TNHA_SSMH_N_COMP <- wide$`TNHA-SC`-wide$`SSMH-SB`
wide$TNHA_SSMH_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SA`
wide$TNHA_SSMH_N_S_COMP <- wide$`TNHA-SC`-wide$`SSMH-SA`
wide$TNHA_N_S_DIFF <- wide$`SSMH-SB`-wide$`TNHA-SC`


#write.csv(wide, "~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_daily_average_difference_N_vs_S.csv")

x<- wide %>% filter(depth == "30cm") 

mean(x$TNHA_NS_DIFF,na.rm = TRUE)
mean(x$SSMH_NS_DIFF, na.rm = TRUE)
mean(x$TNHA_S_BEO_DIFF, na.rm = TRUE)
mean(x$SSMH_S_BEO_DIFF, na.rm = TRUE)

mean(x$SSMH_N_BEO_DIFF, na.rm = TRUE)
mean(x$TNHA_SSMH_N_S_COMP, na.rm = TRUE)
mean(x$TNHA_N_S_DIFF, na.rm = TRUE)
mean(x$TNHA_SSMH_S_COMP, na.rm = TRUE)

max(x$TNHA_NS_DIFF, na.rm = TRUE)
max(x$SSMH_NS_DIFF, na.rm = TRUE)

#below, p1 = 6/19-8/01 , p2= 8/01-
p3 = with(x, mean(TNHA_SSMH_N_COMP[date >= "2023-06-01" & date <= "2023-09-01"]))

#CHANGE DATA RANGE ABOVE

```



SUMMER 2024. AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- SUMM24_ALLDEPTHS %>% filter(fullname == "TNHA-BASE") %>% aggregate(value~date+depth, mean)
averaged_by_station$fullname <- c("TNHA-BASE")


averaged_by_station2 <- SUMM24_ALLDEPTHS %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station2$fullname <- c("TNHA-SA")

averaged_by_station3 <- SUMM24_ALLDEPTHS %>% filter(fullname == "TNHA-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station3$fullname <- c("TNHA-SB")


averaged_by_station4 <- SUMMERGRND23_ALLDEPTHS %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date+depth, mean)
averaged_by_station4$fullname <- c("TNHA-SC")

averaged_by_station5 <- SUMM24_ALLDEPTHS %>% filter(fullname == "TNHA-SD") %>% aggregate(value~date+depth, mean)
averaged_by_station5$fullname <- c("TNHA-SD")

averaged_by_station6 <- SUMM24_ALLDEPTHS %>% filter(fullname == "SSMH-BASE") %>% aggregate(value~date+depth, mean)
averaged_by_station6$fullname <- c("SSMH-BASE")

averaged_by_station7 <- SUMM24_ALLDEPTHS %>% filter(fullname == "SSMH-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station7$fullname <- c("SSMH-SA")

averaged_by_station8 <- SUMM24_ALLDEPTHS %>% filter(fullname == "SSMH-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station8$fullname <- c("SSMH-SB")

averaged_by_station9 <- SUMM24_ALLDEPTHS %>% filter(fullname == "SSMH-SD") %>% aggregate(value~date+depth, mean)
averaged_by_station9$fullname <- c("SSMH-SD")


averaged_by_station10 <- SUMM24_ALLDEPTHS %>% filter(fullname == "SSMH-SE") %>% aggregate(value~date+depth, mean)
averaged_by_station10$fullname <- c("SSMH-SE")

averaged_by_station11 <- SUMM24_ALLDEPTHS %>% filter(fullname == "SSMH-SF") %>% aggregate(value~date+depth, mean)
averaged_by_station11$fullname <- c("SSMH-SF")

averaged_by_station12 <- SUMM24_ALLDEPTHS %>% filter(fullname == "SSMH-SG") %>% aggregate(value~date+depth, mean)
averaged_by_station12$fullname <- c("SSMH-SG")

averaged_by_station13 <- SUMM24_ALLDEPTHS %>% filter(fullname == "SSMH-SH") %>% aggregate(value~date+depth, mean)
averaged_by_station13$fullname <- c("SSMH-SH")

averaged_by_station14 <- SUMM24_ALLDEPTHS %>% filter(fullname == "SSMH-SI") %>% aggregate(value~date+depth, mean)
averaged_by_station14$fullname <- c("SSMH-SI")

averaged_by_station15 <- SUMM24_ALLDEPTHS %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06") %>% aggregate(value~date+depth, mean)
 averaged_by_station15$fullname <- c("BEO")



SUMMERGRND24_STATIONS_AVG<- rbind(averaged_by_station, averaged_by_station2, averaged_by_station3,averaged_by_station4,averaged_by_station5,averaged_by_station6,averaged_by_station7,averaged_by_station8,averaged_by_station9,averaged_by_station10,averaged_by_station11,averaged_by_station12,averaged_by_station13,averaged_by_station14,averaged_by_station15)
```


STATS: average daily difference between SPECIFIC STATIONS for Xcm depth (define in x <- )
```{r}
wide <- SUMMERGRND24_STATIONS_AVG %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_NBASE_DIFF <- wide$`TNHA-BASE` - wide$`TNHA-SC`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO`

wide$SSMH_NS_DIFF <- wide$`SSMH-SA` - wide$`SSMH-SB`
wide$SSMH_N_BEO_DIFF <- wide$`SSMH-SB` - wide$`BEO`
wide$SSMH_S_BEO_DIFF <- wide$`SSMH-SA` - wide$`BEO`

wide$TNHA_SSMH_N_COMP <- wide$`TNHA-SC`-wide$`SSMH-SB`
wide$TNHA_SSMH_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SA`
wide$TNHA_SSMH_N_S_COMP <- wide$`TNHA-SC`-wide$`SSMH-SA`
wide$TNHA_N_S_DIFF <- wide$`SSMH-SB`-wide$`TNHA-SC`


#write.csv(wide, "~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_daily_average_difference_N_vs_S.csv")

x<- wide %>% filter(depth == "30cm") 

mean(x$TNHA_NS_DIFF,na.rm = TRUE) # TNHA SA - TNHA SC (TNHA SOUTH - NORTH)
mean(x$TNHA_NBASE_DIFF, na.rm = TRUE) #TNHA BASE - TNHA SC (TNHA SOUTH NORTH (BASE))
mean(x$SSMH_NS_DIFF, na.rm = TRUE) #SSMH SA - SSMH SC (SSMH SOUTH - NORTH)
mean(x$TNHA_S_BEO_DIFF, na.rm = TRUE) # TNHA SA - BEO (TNHA SOUTH - BEO)
mean(x$TNHA_N_BEO_DIFF, na.rm = TRUE) # TNHA-SC - BEO (TNHA NORTH - BEO)

mean(x$SSMH_S_BEO_DIFF, na.rm = TRUE) # SSMH SA - BEO (SSMH SOUTH - BEO)

mean(x$SSMH_N_BEO_DIFF, na.rm = TRUE) #SSMH SB - BEO (SSMH NORTH - BEO)

mean(x$TNHA_SSMH_N_S_COMP, na.rm = TRUE) # TNHA SA - SSMH SA (TNHA SOUTH - SSMH SOUTH)
mean(x$TNHA_N_S_DIFF, na.rm = TRUE) # SSMH SB - TNHA SC (SSMH NORTH - TNHA NORTH)
mean(x$TNHA_SSMH_S_COMP, na.rm = TRUE) # TNHA SA - SSMH SA (TNHA SOUTH - SSMH SOUTH)



#below, p1 = 6/19-8/01 , p2= 8/01-
#p3 = with(x, mean(TNHA_SSMH_N_COMP[date >= "2023-06-01" & date <= "2023-09-01"]))



```
