---
title: "Differences_in_averages"
author: "Mirella Shaban"
date: "2023-10-23"
output: html_document

############################################################
THIS SCRIPT GIVES YOU THE SEASONAL AVERAGES 
1) BY SITE & 2) BY ASPECT AT THE SAME SITE
############################################################
---
load packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape) # basically like text to columns
library(naniar)
library(RColorBrewer)
library(caret)
```

GRND TMP SUMMER 2022
```{r}
grndtmp22=read.csv("~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_22_grndtmp_2023_11_14_20_50_46_UTC_1.csv")
names(grndtmp22) <- sub("....C..RX3000_BRW1", "", names(grndtmp22))
names(grndtmp22) <- sub("....C..RX3000_BRW4", "", names(grndtmp22))
names(grndtmp22) <- sub("....C..RX3000_BRW5", "", names(grndtmp22))
names(grndtmp22) <- sub("....C..RX3000_BRW6", "", names(grndtmp22))
names(grndtmp22) <- sub(".RXW.GP6.", "", names(grndtmp22))
grndtmp22$Date <-as.POSIXct(grndtmp22$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

long data frame
```{r}
grndtmp = gather(grndtmp22, variable, response, "Temperature.21198259.21206939.7":"Temperature.21401803.21398591.17") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
summer22grndtmp = cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
summer22grndtmp$variable_1 <- NULL
colnames(summer22grndtmp)[3:6] <- c("value","station", "sensor", "depth")

summer22grndtmp$depth <- as.factor(as.character(summer22grndtmp$depth))
summer22grndtmp$sensor <- as.factor(as.character(summer22grndtmp$sensor))
summer22grndtmp$station <- as.factor(summer22grndtmp$station)
summer22grndtmp$station <- revalue(summer22grndtmp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
summer22grndtmp$sensor <- revalue(summer22grndtmp$sensor, c('21206939'='TNHA-BASE','21393042'='BUECI-SC','21393044'='SSMH-SB','21393047'='TNHA-SC','21393048'='BEO-B05','21393049'='SSMH-SD','21398578'='BUECI-SF01','21398579'='BUECI-SE','21398583'='BUECI-SB','21398584'='BUECI-SD','21398585'='BUECI-BASE','21398588'='SSMH-BASE','21398590'='BUECI-SA','21398591'='BEO-B06','21398593'='TNHA-SA','21398598'='BUECI-SF02','21398599'='SSMH-SA','21398601'='TNHA-SD','21398576'='SSMH-SC','21398587'='TNHA-SB'))

summer22grndtmp$depth <- revalue(summer22grndtmp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

summer22grndtmp$depth = factor(summer22grndtmp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))



```

Daily average grnd temperature per sensor at BEO,TNHA,SSMH. Below we force NA values into each indiv. station df & stack them to form a comprehensive df with NA's (plots will disconnect when NA values are present)

-------------------
10cm
```{r}
grndmean <- summer22grndtmp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "10cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

SUMMER22_10CM <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
SUMMER22_10CM$depth <- "10cm"
SUMMER22_10CM$depth <- as.factor(SUMMER22_10CM$depth)
```

30cm
```{r}
grndmean <- summer22grndtmp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "30cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

SUMMER22_30CM <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
SUMMER22_30CM$depth <- "30cm"
SUMMER22_30CM$depth <- as.factor(SUMMER22_30CM$depth)
```

55cm
```{r}
grndmean <- summer22grndtmp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "55cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

SUMMER22_55CM <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
SUMMER22_55CM$depth <- "55cm"
SUMMER22_55CM$depth <- as.factor(SUMMER22_55CM$depth)
```

65cm
```{r}
grndmean <- summer22grndtmp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "65cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

SUMMER22_65CM <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
SUMMER22_65CM$depth <- "65cm"
SUMMER22_65CM$depth <- as.factor(SUMMER22_65CM$depth)
```

90cm
```{r}
grndmean <- summer22grndtmp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "90cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

SUMMER22_90CM <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
SUMMER22_90CM$depth <- "90cm"
SUMMER22_90CM$depth <- as.factor(SUMMER22_90CM$depth)
```



Bind all depth dataframes together
```{r}
SUMMERGRND22_ALLDEPTHS <- rbind(SUMMER22_10CM,SUMMER22_30CM,SUMMER22_55CM,SUMMER22_65CM,SUMMER22_90CM)
```








AVERAGE BY SITE
-----------
```{r}
averaged_by_site_TNHA <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-BASE" | fullname == "TNHA-SA" | fullname == "TNHA-SB"| fullname == "TNHA-SC" | fullname == "TNHA-SD") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_TNHA$station <- c("TNHA")

averaged_by_site_BEO <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "BEO-B06" | fullname == "BEO-B05") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_BEO$station <- c("BEO") 

averaged_by_site_SSMH <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "SSMH-BASE" | fullname == "SSMH-SA" | fullname == "SSMH-SB"| fullname == "SSMH-SC" | fullname == "SSMH-SD") %>% aggregate(value~date+depth+val, mean, na.rm=FALSE)
averaged_by_site_SSMH$station <- c("SSMH")

SITES_AVG_SUM22_GRND <- rbind(averaged_by_site_BEO, averaged_by_site_TNHA, averaged_by_site_SSMH)
```


STATS: average daily difference between SITES (all stations combined) for Xcm depth (define in x <- )
```{r}
wide <- SITES_AVG_SUM22_GRND %>%
    pivot_wider(names_from = station, values_from = value)

wide$TBDIFF <- wide$TNHA - wide$BEO
wide$SBDIFF <- wide$SSMH - wide$BEO
wide$TBSBDIFF <- wide$SSMH - wide$TNHA

x<- wide %>% filter(depth == "10cm") 
mean(x$TBDIFF,na.rm = TRUE)
mean(x$SBDIFF, na.rm = TRUE)
mean(x$TBSBDIFF, na.rm = TRUE)
```

AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station$fullname <- c("TNHA-SA")
averaged_by_station2 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date+depth, mean)
averaged_by_station2$fullname <- c("TNHA-SC")
averaged_by_station3 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "SSMH-SA") %>% aggregate(value~date+depth, mean)
averaged_by_station3$fullname <- c("SSMH-SA")
averaged_by_station4 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "SSMH-SB") %>% aggregate(value~date+depth, mean)
averaged_by_station4$fullname <- c("SSMH-SB")
averaged_by_station5 <- SUMMERGRND22_ALLDEPTHS %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06") %>% aggregate(value~date+depth, mean)
averaged_by_station5$fullname <- c("BEO")
TNHA_NvsS <- rbind(averaged_by_station, averaged_by_station2, averaged_by_station3,averaged_by_station4,averaged_by_station5)
```


STATS: average daily difference between SPECIFIC STATIONS for Xcm depth (define in x <- )
```{r}
wide <- TNHA_NvsS %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO`

wide$SSMH_NS_DIFF <- wide$`SSMH-SA` - wide$`SSMH-SB`
wide$SSMH_N_BEO_DIFF <- wide$`SSMH-SB` - wide$`BEO`
wide$SSMH_S_BEO_DIFF <- wide$`SSMH-SA` - wide$`BEO`

wide$TNHA_SSMH_N_COMP <- wide$`TNHA-SC`-wide$`SSMH-SB`
wide$TNHA_SSMH_S_COMP <- wide$`TNHA-SA`-wide$`SSMH-SA`

#write.csv(wide, "~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_daily_average_difference_N_vs_S.csv")

x<- wide %>% filter(depth == "30cm") 

mean(x$TNHA_NS_DIFF,na.rm = TRUE)
mean(x$SSMH_NS_DIFF, na.rm = TRUE)


mean(x$TNHA_S_BEO_DIFF, na.rm = TRUE)
mean(x$SSMH_S_BEO_DIFF, na.rm = TRUE)

max(x$TNHA_NS_DIFF, na.rm = TRUE)
max(x$SSMH_NS_DIFF, na.rm = TRUE)

x<- wide %>% filter(depth == "90cm") 
#below, p1 = 6/19-8/01 , p2= 8/01-
p1 = with(x, mean(TNHA_SSMH_N_COMP[date >= "2022-06-19" & date <= "2022-07-08"]))
p1

```


------------------------------------------------
AIR TEMPERATURES

AIR SUMMER 2022
```{r}
air22 =read.csv("~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_22_airtmp_2023_11_14_20_57_51_UTC_1.csv")
names(air22) <- sub("Temperature..RXW.THC.", "" ,names(air22))
names(air22) <- sub("Temperature..S.THC.", "" ,names(air22))
names(air22) <- sub("S.", "" ,names(air22))
names(air22) <- sub("....C..RX3000_BRW1", "" ,names(air22))
names(air22) <- sub("....C..RX3000_BRW4", "" ,names(air22))
names(air22) <- sub("....C..RX3000_BRW5", "" ,names(air22))
names(air22) <- sub("....C..RX3000_BRW6", "" ,names(air22))
air22$Date <- as.POSIXct(air22$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

AIR FALL 2022
```{r}
airfall22 =read.csv("~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Fall_22_airtmp_2023_11_14_21_43_57_UTC_1.csv")
names(airfall22) <- sub("Temperature..RXW.THC.", "" ,names(airfall22))
names(airfall22) <- sub("Temperature..S.THC.", "" ,names(airfall22))
names(airfall22) <- sub("S.", "" ,names(airfall22))
names(airfall22) <- sub("....C..RX3000_BRW1", "" ,names(airfall22))
names(airfall22) <- sub("....C..RX3000_BRW4", "" ,names(airfall22))
names(airfall22) <- sub("....C..RX3000_BRW5", "" ,names(airfall22))
names(airfall22) <- sub("....C..RX3000_BRW6", "" ,names(airfall22))
airfall22$Date <- as.POSIXct(airfall22$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

Air SUMMER 2023
```{r}
air_2023 =read.csv("~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/UPDATED_SUMMER_23_DATA/New_Summer_2023_AIRTMP_2023_11_01_18_31_41_UTC_1.csv")
names(air_2023) <- sub("Temperature..RXW.THC.", "" ,names(air_2023))
names(air_2023) <- sub("Temperature..S.THC.", "" ,names(air_2023))
names(air_2023) <- sub("S.", "" ,names(air_2023))
names(air_2023) <- sub("....C..RX3000_BRW1", "" ,names(air_2023))
names(air_2023) <- sub("....C..RX3000_BRW4", "" ,names(air_2023))
names(air_2023) <- sub("....C..RX3000_BRW5", "" ,names(air_2023))
names(air_2023) <- sub("....C..RX3000_BRW6", "" ,names(air_2023))
air_2023$Date <- as.POSIXct(air_2023$Date,format="%m/%d/%y %H:%M",tz="UTC")
```



############# DATA CLEANING

```{r}
airtemp = gather(air22, variable, response, "21198259.21218018.1":"21401803.21397541.1") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

summer22airtemp= cSplit(airtemp, 'variable', sep=".", direction = "wide") 
summer22airtemp$variable_3 <- NULL
colnames(summer22airtemp)[3:5] <- c("value", "station", "sensor")
summer22airtemp$value <- as.numeric(as.character(summer22airtemp$value)) 
summer22airtemp$sensor <- as.factor(as.character(summer22airtemp$sensor))
summer22airtemp$station <- as.factor(summer22airtemp$station)

summer22airtemp$station <- revalue(summer22airtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) 

summer22airtemp$sensor <- revalue(summer22airtemp$sensor, c('21187245'='BUECI-SE','21218018'='TNHA-BASE','21362254'='BUECI-SB','21362256'='BUECI-SD','21380919'='TNHA-SA','21390849'='SSMH-BASE','21390851'='BUECI-BASE','21397541'='BEO-BASE','21398665'='SSMH-SB','21398666'='TNHA-SD','21398668'='BUECI-SC','21398670'='SSMH-SA','21398671'='BUECI-SA','21398674'='TNHA-SC','21398676'='TNHA-SB','21397542'='BUECI-BASE??','21398659'='BUECI-__','21398661'='BUECI-__','21398664'='TNHA-??'))

### need to correct the names for bueci stations, i dont have the metadata clearly on here for the air teperature sensors
```

```{r}
airtemp = gather(airfall22, variable, response, "21198259.21218018.1":"21401803.21397541.1") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

fall22airtemp= cSplit(airtemp, 'variable', sep=".", direction = "wide") 
fall22airtemp$variable_3 <- NULL
colnames(fall22airtemp)[3:5] <- c("value", "station", "sensor")
fall22airtemp$value <- as.numeric(as.character(fall22airtemp$value)) 
fall22airtemp$sensor <- as.factor(as.character(fall22airtemp$sensor))
fall22airtemp$station <- as.factor(fall22airtemp$station)

fall22airtemp$station <- revalue(fall22airtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) 

fall22airtemp$sensor <- revalue(fall22airtemp$sensor, c('21187245'='BUECI-SE','21218018'='TNHA-BASE','21362254'='BUECI-SB','21362256'='BUECI-SD','21380919'='TNHA-SA','21390849'='SSMH-BASE','21390851'='BUECI-BASE','21397541'='BEO-BASE','21398665'='SSMH-SB','21398666'='TNHA-SD','21398668'='BUECI-SC','21398670'='SSMH-SA','21398671'='BUECI-SA','21398674'='TNHA-SC','21398676'='TNHA-SB','21397542'='BUECI-BASE??','21398659'='BUECI-__','21398661'='BUECI-__','21398664'='TNHA-??'))
```

```{r}
airtempsum23 = gather(air_2023, variable, response, "21198259.21380919.1_TNHA":"21401801.21398667.1_MH") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

summer23airtemp= cSplit(airtempsum23, 'variable', sep=".", direction = "wide") 
summer23airtemp= cSplit(summer23airtemp, 'variable_3', sep="1_", direction = "wide") 
summer23airtemp$variable_3_1 <- NULL
colnames(summer23airtemp)[3:6] <- c("value", "station", "sensor","site")
summer23airtemp$value <- as.numeric(as.character(summer23airtemp$value)) 
summer23airtemp$sensor <- as.factor(as.character(summer23airtemp$sensor))
summer23airtemp$station <- as.factor(summer23airtemp$station)
summer23airtemp$site <- as.factor(summer23airtemp$site)
summer23airtemp$station <- revalue(summer23airtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) 

summer23airtemp$sensor <- revalue(summer23airtemp$sensor, c('21187245'='BUECI-SE','21218018'='TNHA-BASE','21362254'='BUECI-SB','21362256'='BUECI-SD','21380919'='TNHA-SA','21390849'='SSMH-BASE','21390851'='BUECI-BASE','21397541'='BEO-BASE','21398665'='SSMH-SB','21398666'='TNHA-SD','21398668'='BUECI-SC','21398670'='SSMH-SA','21398671'='BUECI-SA','21398674'='TNHA-SC','21398676'='TNHA-SB','21397542'='BUECI-BASE??','21398659'='BUECI-__','21398661'='BUECI-__','21398664'='TNHA-??'))


#### need to adjust for the air temp sensors
```

AVERAGING



```{r}
AIRmean <- summer22airtemp %>% subset() %>% select(value,sensor,Date) %>% filter( !value == "-888.8") 

AIRmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

AIRmean$time <- format(as.Date(
  AIRmean$Date),format = "%H:%M:%S")
AIRmean$day <- as.Date(AIRmean$Date)


Daily_average_air_per_sensor <- AIRmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_air_per_sensor <-  cSplit(Daily_average_air_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_air_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_air_per_sensor$fullname <- paste(Daily_average_air_per_sensor$station,Daily_average_air_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_air_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_air_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_air_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-BASE"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

SUMMER22_air <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
```

average daily difference between TNHA and BEO SITES (all stations combined) 
```{r}
wide <- SUMMER22_air %>%
    pivot_wider(names_from = station, values_from = value)

wide$TBDIFF <- wide$TNHA - wide$BEO


mean(wide$TBDIFF,na.rm = TRUE)
mean(wide$SBDIFF, na.rm = TRUE)
mean(wide$TBSBDIFF, na.rm = TRUE)
```

This averages for SENSOR (can compare sensors at the same site, such as N & S) by day
```{r}
averaged_by_station <- SUMMER22_air %>% filter(fullname == "TNHA-SA") %>% aggregate(value~date, mean)
averaged_by_station$fullname <- c("TNHA-SA")

averaged_by_station2 <- SUMMER22_air %>% filter(fullname == "TNHA-SC") %>% aggregate(value~date, mean)
averaged_by_station2$fullname <- c("TNHA-SC")

averaged_by_station3 <- SUMMER22_air %>% filter(fullname == "BEO-BASE") %>% aggregate(value~date, mean)
averaged_by_station3$fullname <- c("BEO-BASE")
TNHA_NvsS <- rbind(averaged_by_station, averaged_by_station2, averaged_by_station3)
```

average daily difference between SPECIFIC STATIONS at TNHA and BEO
```{r}
wide <- TNHA_NvsS %>%
    pivot_wider(names_from = fullname, values_from = value)

wide$TNHA_NS_DIFF <- wide$`TNHA-SA` - wide$`TNHA-SC`
wide$TNHA_SN_DIFF <- wide$`TNHA-SC` - wide$`TNHA-SA`
wide$TNHA_N_BEO_DIFF <- wide$`TNHA-SC` - wide$`BEO-BASE`
wide$TNHA_S_BEO_DIFF <- wide$`TNHA-SA` - wide$`BEO-BASE`



mean(wide$TNHA_NS_DIFF,na.rm = TRUE)

mean(wide$TNHA_S_BEO_DIFF, na.rm = TRUE)

max(x$TNHA_NS_DIFF, na.rm = TRUE)

```
