---
title: "Diurnal_Cycles_SOLAR"
author: "Kiefer Fallin"
date: "1/28/2023"
output: html_document
---

# loading packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape)
library(naniar)
library(RColorBrewer)
library(clock)    #this is a new package that I downloaded to use the date_group() function (basically group_by() but for POSIXct vectors)
library(timechange)  #my lubridate package would not work unless I loaded this one as well
library(lubridate)
```

# solar radiation for all sites data prepping

```{r}
solarrad619_920 = read.csv("/Users/kieferfallin/Desktop/Research/Season_SOLARAD_6_19_22_9_20_22_2022_09_21_16_27_20_UTC_1.csv")
names(solarrad619_920) <- sub("olar.Radiation..RXW.LIB", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("olar.Radiation..S.LIB", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("S.", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW1", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW4", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW5", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW6", "" ,names(solarrad619_920))
solarrad619_920$Date <- as.POSIXct(solarrad619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

# creating long dataset for solar radiation (Mirella's code)

```{r}
solarrad = gather(solarrad619_920, variable, response, '21401801.21362319.1':'21198259.21398620.1') %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
solarrad= cSplit(solarrad, 'variable', sep=".", direction = "wide") # texttocols, 
solarrad$variable_3 <- NULL
colnames(solarrad)[3:5] <- c("value", "station", "sensor")
solarrad$value <- as.numeric(as.character(solarrad$value)) 
solarrad$sensor <- as.factor(as.character(solarrad$sensor))
solarrad$station <- as.factor(solarrad$station)
solarrad$station <- revalue(solarrad$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO"))
solarrad$sensor <- revalue(solarrad$sensor, c("21390411" = "BUECI-BASE", "21398618" = "BUCEI-SA", "21398624" = "BUECI-SB", "21362313" = "BUCEI-SC", "21362316" = "BUCEI-SD", "21362320" = "BUCEI-SE", "21398578" = "BUCEI-SF.01", "21398598" = "BUCEI-SF.02", 
                                              "21390413" ="SSMH-BASE", "21398622" = "SSMH-SA", "21362319" = "SSMH-SB", "21166008" = "SSMH-SC", "21393049" = "SSMH-SD" ,
                                              "21176526" = "TNHA-BASE", "21398620" = "TNHA-SA", "21398616"="TNHA-SB","21362315" = "TNHA-SD", "21362317" = "TNHA-SC", 
                                              "21390415" = "BEO-Base"))
```


# Averaging data by hour per day for sensor over 6/19-6/24 for TNHA-BASE
```{r}
# Filter for the specified sensor and date range
TNHA_BASE_solar <- solarrad %>%
  filter(sensor == "TNHA-BASE" & 
         Date >= "2022-06-19 00:00:00" & 
         Date <= "2022-06-24 23:55:00")
# for some reason this still excludes observations from 6/19 up to 04:00:00 and includes some 6/25 dates

```

# averging data for TNHA-BASE by hour per day 
```{r}
TNHA_BASE_solaravg <- TNHA_BASE_solar %>%
  group_by(hour = date_group(Date,"hour"), day = date_group(Date, "day")) %>%
  summarize(mean_solar = mean(value)) #this is the average solar radiation for every hour of every day in the date range
```

# adding columns that have just the hour and just the day,month, and year
```{r}
TNHA_BASE_solaravg$hour_only <- format(TNHA_BASE_solaravg$hour, "%H")
TNHA_BASE_solaravg$day_only <- format(TNHA_BASE_solaravg$hour, "%m/%d/%y")
```

# Getting the mean solar radiation for each hour for the week
```{r}
TNHA_BASE_daily_hourly_avg <- TNHA_BASE_solaravg %>% 
  group_by(hour_only) %>% 
  summarise(hourly_solar_avg = mean(mean_solar))%>%
  mutate(hour=as.numeric(hour_only))%>%
  select(hour,hourly_solar_avg)

head(TNHA_BASE_daily_hourly_avg,24)
```

# Doing the same thing but for all of the stations combined

## Subsetting for just 6-19 to 6-24 for TNHA sites
```{r}
TNHAsolar <- solarrad %>%
  filter(station == "TNHA" & Date >= "2022-06-19 00:00:00" & Date <= "2022-06-24 23:55:00")
```

# Creating dataset with the average solar radiation at a given hour for TNHA throughout the week
```{r}
TNHAsolaravg <- TNHAsolar %>%
  group_by(hour = date_group(Date,"hour"), day = date_group(Date, "day")) %>%
  summarize(mean_solar = mean(value))

TNHAsolaravg$hour_only <- format(TNHAsolaravg$hour, "%H")
TNHAsolaravg$day_only <- format(TNHAsolaravg$hour, "%m/%d/%y")


TNHA_daily_hourly_avg <- TNHAsolaravg %>% 
  group_by(hour_only) %>% 
  summarise(hourly_solar_avg = mean(mean_solar))%>%
  mutate(hour=as.numeric(hour_only))%>%
  select(hour,hourly_solar_avg)

head(TNHA_daily_hourly_avg,24)
```







