---
title: "LOADANDCLEAN_DIFFERENCES_IN_AVERAGES_VWC"
author: "Mirella Shaban"
date: "2024-09-08"
output: html_document
---

load packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape) # basically like text to columns
library(naniar)
library(RColorBrewer)
library(caret)
library(scales)
```

############################################################
THIS SCRIPT IMPORTS (VWC) DATA FOR SUMMER 22, FALL 22, SPRING 23, SUMMER 23, AND FALL 23* TO THEN BE ANALYZED IN DIFFERENCES_IN_AVERAGES.RMD
*Run this BEFORE YOU RUN Differences_in_averages*

1) BY SITE & 2) BY ASPECT AT THE SAME SITE
############################################################


Data loading and cleaning (summ22,fall22,summ23,fall23)
----
Ground Moisture for all sites data prepping
```{r}
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/Meteorological_Seasons_Data/RECENT_CSV_DOWNLOADS/"
mst1=read.csv(paste(filepath,"Summer_22_VWC_2023_11_14_21_00_08_UTC_1.csv",sep = ""))
mst2=read.csv(paste(filepath,"Fall_22_VWC_2023_11_14_21_46_41_UTC_1.csv", sep = ""))
mst3=read.csv(paste(filepath,"New_Summer_2023_VWC_2023_11_01_18_34_52_UTC_1.csv", sep = ""))
mst4=read.csv(paste(filepath,"Fall_2023_VWC_2023_12_05_20_06_01_UTC_1.csv", sep = ""))

#column name cleaning
names(mst1) <- sub("...m.3.m.3..RX3000_BRW1", "", names(mst1))
names(mst1) <- sub("...m.3.m.3..RX3000_BRW4", "", names(mst1))
names(mst1) <- sub("...m.3.m.3..RX3000_BRW5", "", names(mst1))
names(mst1) <- sub("...m.3.m.3..RX3000_BRW6", "", names(mst1))
names(mst1) <- sub("..RXW.GP6.", "", names(mst1))
names(mst1) <- sub("Water.Content", "", names(mst1))
mst1$Date <-as.POSIXct(mst1$Date,format="%m/%d/%y %H:%M",tz="UTC")

names(mst2) <- sub("...m.3.m.3..RX3000_BRW1", "", names(mst2))
names(mst2) <- sub("...m.3.m.3..RX3000_BRW4", "", names(mst2))
names(mst2) <- sub("...m.3.m.3..RX3000_BRW5", "", names(mst2))
names(mst2) <- sub("...m.3.m.3..RX3000_BRW6", "", names(mst2))
names(mst2) <- sub("..RXW.GP6.", "", names(mst2))
names(mst2) <- sub("Water.Content", "", names(mst2))
mst2$Date <-as.POSIXct(mst2$Date,format="%m/%d/%y %H:%M",tz="UTC")

names(mst3) <- sub("...m.3.m.3..RX3000_BRW1", "", names(mst3))
names(mst3) <- sub("...m.3.m.3..RX3000_BRW4", "", names(mst3))
names(mst3) <- sub("...m.3.m.3..RX3000_BRW5", "", names(mst3))
names(mst3) <- sub("...m.3.m.3..RX3000_BRW6", "", names(mst3))
names(mst3) <- sub("..RXW.GP6.", "", names(mst3))
names(mst3) <- sub("_TNHA", "", names(mst3))
names(mst3) <- sub("_BUECI", "", names(mst3))
names(mst3) <- sub("_SSMH", "", names(mst3))
names(mst3) <- sub("_BEO", "", names(mst3))
names(mst3) <- sub("Water.Content", "", names(mst3))

mst3$Date <-as.POSIXct(mst3$Date,format="%m/%d/%y %H:%M",tz="UTC")

names(mst4) <- sub("...m.3.m.3..RX3000_BRW1", "", names(mst4))
names(mst4) <- sub("...m.3.m.3..RX3000_BRW4", "", names(mst4))
names(mst4) <- sub("...m.3.m.3..RX3000_BRW5", "", names(mst4))
names(mst4) <- sub("...m.3.m.3..RX3000_BRW6", "", names(mst4))
names(mst4) <- sub("..RXW.GP6.", "", names(mst4))
names(mst4) <- sub("Water.Content", "", names(mst4))
names(mst4) <- sub("_TNHA", "", names(mst4))
names(mst4) <- sub("_BUECI", "", names(mst4))
names(mst4) <- sub("_SSMH", "", names(mst4))
names(mst4) <- sub("_BEO", "", names(mst4))
mst4$Date <-as.POSIXct(mst4$Date,format="%m/%d/%y %H:%M",tz="UTC")

```

summer 22
```{r}
#pivot longer vs gather
mst1_df = gather(mst1, variable, response, "21198259.21206939.1":"21401803.21398591.6") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
vwc_summ22 = cSplit(mst1_df , 'variable', sep=".", direction = "wide") 
colnames(vwc_summ22)[3:6] <- c("value","station", "sensor", "depth")
vwc_summ22$depth <- as.factor(as.character(vwc_summ22$depth)) 
vwc_summ22$station <- as.factor(vwc_summ22$station)
vwc_summ22$sensor <- as.factor(as.character(vwc_summ22$sensor))
vwc_summ22$station <- revalue(vwc_summ22$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
vwc_summ22$depth <- revalue(vwc_summ22$depth, c("1" = "0:15cm", "2" = "15:30cm", "3" = "30:45cm", "4" = "45:60cm", "5" = "60:75cm", "6" = "75:90cm"))
vwc_summ22$sensor <- revalue(vwc_summ22$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUCEI-SA", "21398583" = "BUECI-SB", "21393042" = "BUCEI-SC", "21398584" = "BUCEI-SD", "21398579" = "BUCEI-SE", "21398578" = "BUCEI-SF.01", "21398598" = "BUCEI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398587"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05")) 
```
fall 22
```{r}
mst2_df = gather(mst2, variable, response, "21198259.21206939.1":"21401803.21398591.6") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
vwc_fall22 = cSplit(mst2_df , 'variable', sep=".", direction = "wide") 
colnames(vwc_fall22)[3:6] <- c("value","station", "sensor", "depth")
vwc_fall22$depth <- as.factor(as.character(vwc_fall22$depth)) 
vwc_fall22$station <- as.factor(vwc_fall22$station)
vwc_fall22$sensor <- as.factor(as.character(vwc_fall22$sensor))
vwc_fall22$station <- revalue(vwc_fall22$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
vwc_fall22$depth <- revalue(vwc_fall22$depth, c("1" = "0:15cm", "2" = "15:30cm", "3" = "30:45cm", "4" = "45:60cm", "5" = "60:75cm", "6" = "75:90cm"))
vwc_fall22$sensor <- revalue(vwc_fall22$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUCEI-SA", "21398583" = "BUECI-SB", "21393042" = "BUCEI-SC", "21398584" = "BUCEI-SD", "21398579" = "BUCEI-SE", "21398578" = "BUCEI-SF.01", "21398598" = "BUCEI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398587"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05")) 
```

summer 23
```{r}
mst3_df = gather(mst3, variable, response, "21198259.21206939.1":"21401803.21398591.6") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
vwc_summ23 = cSplit(mst3_df , 'variable', sep=".", direction = "wide") 
colnames(vwc_summ23)[3:6] <- c("value","station", "sensor", "depth")
vwc_summ23$depth <- as.factor(as.character(vwc_summ23$depth)) 
vwc_summ23$station <- as.factor(vwc_summ23$station)
vwc_summ23$sensor <- as.factor(as.character(vwc_summ23$sensor))
vwc_summ23$station <- revalue(vwc_summ23$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
vwc_summ23$depth <- revalue(vwc_summ23$depth, c("1" = "0:15cm", "2" = "15:30cm", "3" = "30:45cm", "4" = "45:60cm", "5" = "60:75cm", "6" = "75:90cm"))
vwc_summ23$sensor <- revalue(vwc_summ23$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398587" = "TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05","21393045" = "BEO-B07", "21398576" = "BEO-B08",
                                                      "21393043" = "SSMH-SE",
                                                      "21398594" = "SSMH-SF",
                                                      "21398580" = "SSMH-SG",
                                                      "21398586" = "SSMH-SH",
                                                      "21398581" = "SSMH-SI")) 
```

fall 23
```{r}
mst4_df = gather(mst4, variable, response, "21198259.21206939.1":"21401803.21398576.6") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
vwc_fall23 = cSplit(mst4_df , 'variable', sep=".", direction = "wide") 
colnames(vwc_fall23)[3:6] <- c("value","station", "sensor", "depth")
vwc_fall23$depth <- as.factor(as.character(vwc_fall23$depth)) 
vwc_fall23$station <- as.factor(vwc_fall23$station)
vwc_fall23$sensor <- as.factor(as.character(vwc_fall23$sensor))
vwc_fall23$station <- revalue(vwc_fall23$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
vwc_fall23$depth <- revalue(vwc_fall23$depth, c("1" = "0:15cm", "2" = "15:30cm", "3" = "30:45cm", "4" = "45:60cm", "5" = "60:75cm", "6" = "75:90cm"))
vwc_fall23$sensor <- revalue(vwc_fall23$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398587" = "TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05","21393045" = "BEO-B07", "21398576" = "BEO-B08",
                                                      "21393043" = "SSMH-SE",
                                                      "21398594" = "SSMH-SF",
                                                      "21398580" = "SSMH-SG",
                                                      "21398586" = "SSMH-SH",
                                                      "21398581" = "SSMH-SI")) 
```
combine all 
```{r}
VWC22_23_combined <- rbind(vwc_summ22,vwc_fall22,vwc_summ23,vwc_fall23)
```


---

0 to 15 cm
```{r}
vwcmean <- VWC22_23_combined %>% subset() %>% select(Date,value,sensor,depth) %>% filter(depth == "0:15cm") 

vwcmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

vwcmean$time <- format(as.Date(
  vwcmean$Date),format = "%H:%M:%S")
vwcmean$day <- as.Date(vwcmean$Date)


Daily_average_vwc_per_sensor <- vwcmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_vwc_per_sensor <-  cSplit(Daily_average_vwc_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_vwc_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_vwc_per_sensor$fullname <- paste(Daily_average_vwc_per_sensor$station,Daily_average_vwc_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_vwc_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2023-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=length(date)/5)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=length(date)/5)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","BEO-B07", "BEO-B08", "FAKEVAR3"),each=length(date)/5)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
TSB_NONA$depth <- "1"
TSB_NONA$depth <- as.factor(TSB_NONA$depth)
```

15 to 30 cm
```{r}
vwcmean1 <- VWC22_23_combined %>% subset() %>% select(value,sensor,Date,depth) %>% filter(depth == "15:30cm") 

vwcmean1 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

vwcmean1$time <- format(as.Date(
  vwcmean1$Date),format = "%H:%M:%S")
vwcmean1$day <- as.Date(vwcmean1$Date)


Daily_average_vwc_per_sensor <- vwcmean1 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_vwc_per_sensor <-  cSplit(Daily_average_vwc_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_vwc_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_vwc_per_sensor$fullname <- paste(Daily_average_vwc_per_sensor$station,Daily_average_vwc_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_vwc_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2023-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=length(date)/5)
testdf <- data.frame(date,fullname,val)
TNHA_NONA15 <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=length(date)/5)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA15 <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","BEO-B07", "BEO-B08", "FAKEVAR3"),each=length(date)/5)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA15 <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA15 <- rbind(TNHA_NONA15,BEO_NONA15,SSMH_NONA15)
TSB_NONA15$depth <- "2"
TSB_NONA15$depth <- as.factor(TSB_NONA15$depth)
```

30 to 45 cm
```{r}
vwcmean2 <- VWC22_23_combined %>% subset() %>% select(value,sensor,Date,depth) %>% filter(depth == "30:45cm") 

vwcmean2 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate(value ~ sensor, mean ) %>%
  group_by(sensor)

vwcmean2$time <- format(as.Date(
  vwcmean2$Date),format = "%H:%M:%S")
vwcmean2$day <- as.Date(vwcmean2$Date)


Daily_average_vwc_per_sensor <- vwcmean2 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_vwc_per_sensor <-  cSplit(Daily_average_vwc_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_vwc_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_vwc_per_sensor$fullname <- paste(Daily_average_vwc_per_sensor$station,Daily_average_vwc_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_vwc_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2023-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=length(date)/5)
testdf <- data.frame(date,fullname,val)
TNHA_NONA30 <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=length(date)/5)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA30 <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","BEO-B07", "BEO-B08", "FAKEVAR3"),each=length(date)/5)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA30 <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA30 <- rbind(TNHA_NONA30,BEO_NONA30,SSMH_NONA30)
TSB_NONA30$depth <- "3"
TSB_NONA30$depth <- as.factor(TSB_NONA30$depth)
```


45 to 60 cm
```{r}
vwcmean3 <- VWC22_23_combined %>% subset() %>% select(value,sensor,Date,depth) %>% filter(depth == "45:60cm") 

vwcmean3 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

vwcmean3$time <- format(as.Date(
  vwcmean3$Date),format = "%H:%M:%S")
vwcmean3$day <- as.Date(vwcmean3$Date)


Daily_average_vwc_per_sensor <- vwcmean3 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_vwc_per_sensor <-  cSplit(Daily_average_vwc_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_vwc_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_vwc_per_sensor$fullname <- paste(Daily_average_vwc_per_sensor$station,Daily_average_vwc_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_vwc_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2023-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=length(date)/5)
testdf <- data.frame(date,fullname,val)
TNHA_NONA55 <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=length(date)/5)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA55 <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","BEO-B07", "BEO-B08", "FAKEVAR3"),each=length(date)/5)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA55 <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA55 <- rbind(TNHA_NONA55,BEO_NONA55,SSMH_NONA55)
TSB_NONA55$depth <- "4"
TSB_NONA55$depth <- as.factor(TSB_NONA55$depth)
```

60 to 75 cm
```{r}
vwcmean4 <- VWC22_23_combined %>% subset() %>% select(value,sensor,Date,depth) %>% filter(depth == "60:75cm") 

vwcmean4 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

vwcmean4$time <- format(as.Date(
  vwcmean4$Date),format = "%H:%M:%S")
vwcmean4$day <- as.Date(vwcmean4$Date)


Daily_average_vwc_per_sensor <- vwcmean4 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_vwc_per_sensor <-  cSplit(Daily_average_vwc_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_vwc_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_vwc_per_sensor$fullname <- paste(Daily_average_vwc_per_sensor$station,Daily_average_vwc_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_vwc_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2023-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=length(date)/5)
testdf <- data.frame(date,fullname,val)
TNHA_NONA65 <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=length(date)/5)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA65 <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","BEO-B07", "BEO-B08", "FAKEVAR3"),each=length(date)/5)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA65 <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA65 <- rbind(TNHA_NONA65,BEO_NONA65,SSMH_NONA65)
TSB_NONA65$depth <- "5"
TSB_NONA65$depth <- as.factor(TSB_NONA65$depth)
```

75 to 90 cm
```{r}
vwcmean5 <- VWC22_23_combined %>% subset() %>% select(value,sensor,Date,depth) %>% filter(depth == "75:90cm") 

vwcmean5 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

vwcmean5$time <- format(as.Date(
  vwcmean5$Date),format = "%H:%M:%S")
vwcmean5$day <- as.Date(vwcmean5$Date)


Daily_average_vwc_per_sensor <- vwcmean5 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_vwc_per_sensor <-  cSplit(Daily_average_vwc_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_vwc_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_vwc_per_sensor$fullname <- paste(Daily_average_vwc_per_sensor$station,Daily_average_vwc_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_vwc_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_vwc_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2023-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=length(date)/5)
testdf <- data.frame(date,fullname,val)
TNHA_NONA90 <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=length(date)/5)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA90 <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","BEO-B07", "BEO-B08", "FAKEVAR3"),each=length(date)/5)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA90 <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA90 <- rbind(TNHA_NONA90,BEO_NONA90,SSMH_NONA90)
TSB_NONA90$depth <- "6"
TSB_NONA90$depth <- as.factor(TSB_NONA90$depth)
```

Combining
```{r}
VWC_DLYAVG_COMB_22_23 <- rbind(TSB_NONA,TSB_NONA15,TSB_NONA30,TSB_NONA55,TSB_NONA65,TSB_NONA90)
```

Renaming depth vector values to be more intuitive
```{r}
VWC_DLYAVG_COMB_22_23 <- VWC_DLYAVG_COMB_22_23 %>%
  mutate(depth = case_when(
    depth == 1 ~ "0:15cm",
    depth == 2 ~ "15:30cm",
    depth == 3 ~ "30:45cm",
    depth == 4 ~ "45:60cm",
    depth == 5 ~ "60:75cm",
    depth == 6 ~ "75:90cm",
    TRUE ~ as.character(depth)  
  ))
```

Split into seasons for station vs station analysis
```{r}
VWC_Summer22_Daily = VWC_DLYAVG_COMB_22_23 %>% filter(date >= "2022-06-01" & date <= "2022-09-01") %>%
  mutate(across(where(is.numeric), ~ round(., 3)))
VWC_Fall22_Daily = VWC_DLYAVG_COMB_22_23 %>% filter(date >= "2022-09-01" & date <= "2022-12-01") %>%
  mutate(across(where(is.numeric), ~ round(., 3)))
VWC_Summer23_Daily = VWC_DLYAVG_COMB_22_23 %>% filter(date >= "2023-06-01" & date <= "2023-09-01") %>%
  mutate(across(where(is.numeric), ~ round(., 3)))
VWC_Fall23_Daily = VWC_DLYAVG_COMB_22_23 %>% filter(date >= "2023-09-01" & date <= "2023-12-01") %>%
  mutate(across(where(is.numeric), ~ round(., 3)))
```

Summer 22 vs summer 23
```{r}
#makes an additonal column substring of the month and day only for both 22 and 23 data
#makes an additonal column substring of the year only for both 22 and 23 data
#rbind the 22 and 23 data

VWC_Summer22_Daily$day <- substr(VWC_Summer22_Daily$date, 6,10)
VWC_Summer22_Daily$day <- as.Date(VWC_Summer22_Daily$day, format = "%m-%d")
VWC_Summer22_Daily$year <- substr(VWC_Summer22_Daily$date, 1,4)

VWC_Summer23_Daily$day <- substr(VWC_Summer23_Daily$date, 6,10)
VWC_Summer23_Daily$year <- substr(VWC_Summer23_Daily$date, 1,4)
VWC_Summer23_Daily$day <- as.Date(VWC_Summer23_Daily$day, format = "%m-%d")

SUMMER_22_23_VWC_DAILY <- rbind(VWC_Summer22_Daily,VWC_Summer23_Daily)

```

Fall 22 vs fall 23
```{r}
VWC_Fall22_Daily$day <- substr(VWC_Fall22_Daily$date, 6,10)
VWC_Fall22_Daily$day <- as.Date(VWC_Fall22_Daily$day, format = "%m-%d")
VWC_Fall22_Daily$year <- substr(VWC_Fall22_Daily$date, 1,4)

VWC_Fall23_Daily$day <- substr(VWC_Fall23_Daily$date, 6,10)
VWC_Fall23_Daily$year <- substr(VWC_Fall23_Daily$date, 1,4)
VWC_Fall23_Daily$day <- as.Date(VWC_Fall23_Daily$day, format = "%m-%d")

FALL_22_23_VWC_DAILY <- rbind(VWC_Fall22_Daily,VWC_Fall23_Daily)

```


