---
title: "Multi-depth averaging and plotting - ground temp"
author: "Mirella Shaban"
date: "2023-07-14"
output: html_document
---

Loading packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape) 
library(naniar)
library(RColorBrewer)
library(clock)
library(readr)
```


*SUMMER 2022 DATA*
10 cm depth 

Ground Temperatures for all sites data prepping
```{r}
# importing ground temp data for 6-19 through 9-20
grndtmp619_920=read.csv("https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/Meteorological_Seasons_Data/Summer_grndtmp_2023_04_06_12_21_30_UTC_1.csv")
names(grndtmp619_920) <- sub("....C..RX3000_BRW1", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW4", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW5", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW6", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub(".RXW.GP6.", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("Temperature.", "", names(grndtmp619_920))

#there may be wind and air variables in the df
grndtmp619_920 = select(grndtmp619_920, !contains(".RXW"))
grndtmp619_920 = select(grndtmp619_920, !contains(".S"))

grndtmp619_920$Date <-as.POSIXct(grndtmp619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

```{r}
grndtmp = gather(grndtmp619_920, variable, response, 3:ncol(grndtmp619_920)) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
dfgrndtemp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(dfgrndtemp)[3] <- ("value")
colnames(dfgrndtemp)[4:6] <- c("station", "sensor", "depth")

dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth))
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))
dfgrndtemp$station <- as.factor(dfgrndtemp$station)
dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c('21206939'='TNHA-BASE','21393042'='BUECI-SC','21393044'='SSMH-SB','21393047'='TNHA-SC','21393048'='BEO-B05','21393049'='SSMH-SD','21398578'='BUECI-SF01','21398579'='BUECI-SE','21398583'='BUECI-SB','21398584'='BUECI-SD','21398585'='BUECI-BASE','21398588'='SSMH-BASE','21398590'='BUECI-SA','21398591'='BEO-B06','21398593'='TNHA-SA','21398598'='BUECI-SF02','21398599'='SSMH-SA','21398601'='TNHA-SD','21398576'='SSMH-SC','21398587'='TNHA-SB'						))

dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))

dfgrndtemp <- dfgrndtemp %>% filter(!value == -888.88)
dfgrndtemp <- dfgrndtemp %>% filter(!sensor == 21187245)
dfgrndtemp <- dfgrndtemp %>% filter(!sensor == 21398676)
dfgrndtemp <- dfgrndtemp %>% filter(!sensor == 21398671)
dfgrndtemp <- dfgrndtemp %>% filter(!sensor == 21398668)
dfgrndtemp <- dfgrndtemp %>% filter(!sensor == 21398666)
dfgrndtemp <- dfgrndtemp %>% filter(!sensor == 21398660)

```


*Daily average ground temperature grouped by sensor at station "X"*
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,station,sensor,Date,depth) %>% filter( !value == "-888.8", depth %in% "10cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value,depth,station) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor_10 <- grndmean %>% 
  group_by(sensor, day, depth) %>%
  aggregate(value ~ sensor + day + depth + station, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season

#Below -- exports as a csv to folder of choice 

#write.csv(Daily_average_grnd_per_sensor_10, "~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_10.csv")

```

*Add sequential dates to where dates are missing in the .csv to avoid connecting lines and add NA values*
This bullshit is giving me issues as the Date object isn't registering as literal dates but rather characters?
*DON'T RUN*
```{r}
#Daily_average_grnd_per_sensor_1$day_test <- Daily_average_grnd_per_sensor$day_1

Daily_average_grnd_per_sensor_10 <- Daily_average_grnd_per_sensor_10 %>%
  mutate(day = as.POSIXct(day)) %>%
  complete(day = seq.POSIXt(min(day), max(day), by="day", na= NA)) %>% group_by(sensor, day) %>% cSplit("day", sep= " ", direction = "wide") 
Daily_average_grnd_per_sensor_10$day_2 <- NULL

Daily_average_grnd_per_sensor$day <- as.POSIXct(Daily_average_grnd_per_sensor$day)

### this stopped working, but originally gave me the NA values I needed for this shit to work.
```


*Write csv of filtered and subset data at multiple depths, add NA values, reimport to get rid of line connections in figures where there is no data*
#### THIS CHUNK BELOW IS NOT NECESSARY IF YOU RUN THE COMPLETE FUNCTION. THIS IS JUST HERE FOR REFERENCE BECAUSE I SPENT TIME MAKING IT..... ####### 

really annoying data cleaning. Added in NA values in excel, then imported the csv, removed quotation marks they imported with and changed the commas to periods. Made the date a date object in R and then rounded to 2 decimal places
```{r}
#write.csv(Daily_average_grnd_per_sensor_10, "~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_10.csv")

#once exported, use text to columns in excel to sepearate by " and then change the thousands separator to commas. Reimport and change commas to periods. To do this in code rather than in the files tab, run the next line below

Daily_Averages_Ground_Temps_Summer_10 <- read_csv("~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_10.csv", na = "NA")



Daily_Averages_Ground_Temps_Summer_10$day <-as.POSIXct(Daily_Averages_Ground_Temps_Summer_10$day,format="%m/%d/%y",tz="UTC")

removeQuotes <- function(x) gsub("\"", "", x)
commatoperiod <- function(y) gsub(",",".", y)

Daily_Averages_Ground_Temps_Summer_10_noquote <- Daily_Averages_Ground_Temps_Summer_10 %>%
    mutate_if(is.character, removeQuotes)
Daily_Averages_Ground_Temps_Summer_10_noquote_nocomma <- Daily_Averages_Ground_Temps_Summer_10_noquote %>%
  mutate_if(is.character, commatoperiod)

Daily_Averages_Ground_Temps_Summer_10_noquote_nocomma$day<-as.POSIXct(Daily_Averages_Ground_Temps_Summer_10_noquote_nocomma$day,format="%m/%d/%y",tz="UTC")

#give it a shorter name...
CLEAN_BEO_TNHA_10 <- Daily_Averages_Ground_Temps_Summer_10_noquote_nocomma
CLEAN_BEO_TNHA_10$day <- as.POSIXct(CLEAN_BEO_TNHA_10$day, format="%m/%d/%y",tz="UTC")
CLEAN_BEO_TNHA_10$day <- as.Date(CLEAN_BEO_TNHA_10$day, format = "%m/%d/%y", tz="UTC")
CLEAN_BEO_TNHA_10$value <- as.numeric(CLEAN_BEO_TNHA_10$value)
round(CLEAN_BEO_TNHA_10$value, digits= 2)

#removerandomtime <- function(w) gsub(" 19:00:00", "", w)
#CLEAN_BEO_TNHA_10 <- CLEAN_BEO_TNHA_10 %>% mutate_if(is.POSIXct, removerandomtime)
```


Dates before value >0
```{r}
before_thaw <- CLEAN_BEO_TNHA_10 %>% filter(value > 0 | value == 0)
```


```{r fig14, fig.height = 6, fig.width = 14}
CLEAN_BEO_TNHA_10_select <- CLEAN_BEO_TNHA_10%>% filter(sensor == "TNHA-BASE" | sensor == "TNHA-SA" | sensor == "TNHA-SC" | sensor == "BEO-B06")

CLEAN_BEO_TNHA_10_select %>%
  ggplot(aes(x=day, y=value, color = sensor), size = 0.9) +
  geom_line() +
  geom_point() +
  guides(color = guide_legend(override.aes = list(size=20)))+
  theme_bw()+ 
  theme(legend.title = element_text(size = 20),legend.text = element_text(size = 10)) +
  labs(title = "Summer 2022 daily average ground temperatures at 10cm depth, control vs. N&S facing sensors",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-2.5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 35),
        axis.title.x = element_text(size = 30),
        axis.title.y = element_text(size = 30)) +
  scale_color_manual(name = "Station",values =c("TNHA-BASE" = "#1B9E77", "TNHA-SC" ="#E7298A", "BEO-B06"= "#084594","TNHA-SA" = "#D95F02", labels = c("BEO (Control)", "TNHA-BASE (South Facing)", "TNHA-SA (South Facing)", "TNHA-SC (North Facing)")))

```


SUMMER 2022 DATA

### 30 cm

Daily average grnd temperature per sensor at station "X" 
*-- DONT RUN, IMPORT CSV "Daily_Averages_Ground_Temps_Summer_30.csv" INSTEAD, IN CHUNK BELOW -- *
```{r}
grndmean30 <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "30cm") 

grndmean30 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean30$time <- format(as.Date(
  grndmean30$Date),format = "%H:%M:%S")
grndmean30$day <- as.Date(grndmean30$Date)


Daily_average_grnd_per_sensor_30 <- grndmean30 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season

#Below -- exports as a csv to folder of choice 

#write.csv(Daily_average_grnd_per_sensor, "~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_NEW.csv")






```


*run below for figure of 30cm*
```{r fig13, fig.height = 6, fig.width = 14}
#imported this file, saved with NA;s in it
Daily_Averages_Ground_Temps_Summer_30 <- read_csv("~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_30.csv", na = "NA")

Daily_Averages_Ground_Temps_Summer_30$day <- as.Date(Daily_Averages_Ground_Temps_Summer_30$day, format = "%m/%d/%y")

Daily_average_grnd_per_sensor_30 %>% filter(sensor == "TNHA-BASE" | sensor == "TNHA-SA" | sensor == "TNHA-SC" | sensor == "TNHA-SB" |sensor == "TNHA-SD" | sensor == "BEO-B06") %>%
  ggplot(aes(x=day, y=value, color = sensor), size = 0.9) +
  geom_line() +
  geom_point() +
  guides(color = guide_legend(override.aes = list(size=20)))+
  theme_bw()+ 
  theme(legend.title = element_text(size = 20),legend.text = element_text(size = 10)) +
  labs(title = "Summer 2022 daily average ground temperatures at 55cm depth, control vs. N&S facing sensors",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-15,-10,-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-15, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 35),
        axis.title.x = element_text(size = 30),
        axis.title.y = element_text(size = 30)) +
  scale_color_manual(name = "Station",values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E", "BEO-B06"= "#084594", labels = c("BEO (Control)", "TNHA-BASE (South Facing)", "TNHA-SA (South Facing)", "TNHA-SC (North Facing)")))

```




really annoying data cleaning. Added in NA values in excel, then imported the csv, removed quotation marks they imported with and changed the commas to periods. Made the date a date object in R and then rounded to 2 decimal places

*dont run*
```{r}
TNHABEOCOMPGRND_30$day <-as.POSIXct(TNHABEOCOMPGRND_30$day,format="%m/%d/%y",tz="UTC")

removeQuotes <- function(x) gsub("\"", "", x)
commatoperiod <- function(y) gsub(",",".", y)

TNHABEOCOMPGRND_SUMMER22_noquote <- TNHABEOCOMPGRND_30 %>%
    mutate_if(is.character, removeQuotes)
TNHABEOCOMPGRND_SUMMER22_noquote_nocomma <- TNHABEOCOMPGRND_SUMMER22_noquote %>%
  mutate_if(is.character, commatoperiod)

TNHABEOCOMPGRND_SUMMER22_noquote_nocomma$day<-as.POSIXct(TNHABEOCOMPGRND_SUMMER22_noquote_nocomma$day,format="%m/%d/%y",tz="UTC")

#give it a shorter name...
CLEAN_BEO_TNHA <- TNHABEOCOMPGRND_SUMMER22_noquote_nocomma
CLEAN_BEO_TNHA$day <- as.POSIXct(CLEAN_BEO_TNHA$day, format="%m/%d/%y",tz="UTC")
CLEAN_BEO_TNHA$day <- as.Date(CLEAN_BEO_TNHA$day, format = "%m/%d/%y", tz="UTC")
CLEAN_BEO_TNHA$value <- as.numeric(CLEAN_BEO_TNHA$value)
round(CLEAN_BEO_TNHA$value, digits= 2)

#removerandomtime <- function(w) gsub(" 19:00:00", "", w)
#CLEAN_BEO_TNHA <- CLEAN_BEO_TNHA %>% mutate_if(is.POSIXct, removerandomtime)
```

```{r fig14, fig.height = 6, fig.width = 14}
CLEAN_BEO_TNHA %>%
  ggplot(aes(x=day, y=value, color = sensor), size = 0.9) +
  geom_line() +
  geom_point() +
  guides(color = guide_legend(override.aes = list(size=20)))+
  theme_bw()+ 
  theme(legend.title = element_text(size = 20),legend.text = element_text(size = 10)) +
  labs(title = "Summer 2022 daily average ground temperatures, control vs. N&S facing sensors (30cm)",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-2.5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 35),
        axis.title.x = element_text(size = 30),
        axis.title.y = element_text(size = 30)) +
  scale_color_manual(name = "Station",values =c("TNHA-BASE" = "#1B9E77", "TNHA-SC" ="#E7298A", "BEO-B06"= "#084594","TNHA-SA" = "#D95F02", labels = c("BEO (Control)", "TNHA-BASE (South Facing)", "TNHA-SA (South Facing)", "TNHA-SC (North Facing)")))

```

SUMMER 2022 DATA
##### 65 cm


Daily average grnd temperature per sensor at station "X"
```{r}
grndmean65 <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "65cm") 

grndmean65 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean65$time <- format(as.Date(
  grndmean65$Date),format = "%H:%M:%S")
grndmean65$day <- as.Date(grndmean65$Date)


Daily_average_grnd_per_sensor_65 <- grndmean65 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season

#Below -- exports as a csv to folder of choice 

#write.csv(Daily_average_grnd_per_sensor_65, "~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_65.csv")


```



```{r}
Daily_Averages_Ground_Temps_Summer_65 <- read_csv("~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_65.csv", na = "NA")

Daily_Averages_Ground_Temps_Summer_65$day<-as.POSIXct(Daily_Averages_Ground_Temps_Summer_65$day,format="%m/%d/%y",tz="UTC")

removeQuotes <- function(x) gsub("\"", "", x)
commatoperiod <- function(y) gsub(",",".", y)

Daily_Averages_Ground_Temps_Summer_65_noquote <- Daily_Averages_Ground_Temps_Summer_65 %>%
    mutate_if(is.character, removeQuotes)
Daily_Averages_Ground_Temps_Summer_65_noquote_nocomma <- Daily_Averages_Ground_Temps_Summer_65_noquote %>%
  mutate_if(is.character, commatoperiod)

Daily_Averages_Ground_Temps_Summer_65_noquote_nocomma$day<-as.POSIXct(Daily_Averages_Ground_Temps_Summer_65_noquote_nocomma$day,format="%m/%d/%y",tz="UTC")

#give it a shorter name...
CLEAN_BEO_TNHA_65 <- Daily_Averages_Ground_Temps_Summer_65_noquote_nocomma
CLEAN_BEO_TNHA_65$day <- as.POSIXct(CLEAN_BEO_TNHA_65$day, format="%m/%d/%y",tz="UTC")
CLEAN_BEO_TNHA_65$day <- as.Date(CLEAN_BEO_TNHA_65$day, format = "%m/%d/%y", tz="UTC")
CLEAN_BEO_TNHA_65$value <- as.numeric(CLEAN_BEO_TNHA_65$value)
round(CLEAN_BEO_TNHA_65$value, digits= 2)
```

```{r fig3, fig.height = 6, fig.width = 14}
CLEAN_BEO_TNHA_65 %>% filter(sensor == "TNHA-BASE" | sensor == "TNHA-SA" | sensor == "TNHA-SC" | sensor == "BEO-B06") %>%
  ggplot(aes(x=day, y=value, color = sensor), size = 0.9) +
  geom_line() +
  geom_point() +
  guides(color = guide_legend(override.aes = list(size=20)))+
  theme_bw()+ 
  theme(legend.title = element_text(size = 20),legend.text = element_text(size = 10)) +
  labs(title = "Summer 2022 daily average ground temperatures at 65cm depth, control vs. N&S facing sensors",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-2.5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 35),
        axis.title.x = element_text(size = 30),
        axis.title.y = element_text(size = 30)) +
  scale_color_manual(name = "Station",values =c("TNHA-BASE" = "#1B9E77", "TNHA-SC" ="#E7298A", "BEO-B06"= "#084594","TNHA-SA" = "#D95F02", labels = c("BEO (Control)", "TNHA-BASE (South Facing)", "TNHA-SA (South Facing)", "TNHA-SC (North Facing)")))

```

#####

SUMMER 2022 DATA
##### 90 cm


Daily average grnd temperature per sensor at station "X"
```{r}
grndmean90 <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "90cm") 

grndmean90 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean90$time <- format(as.Date(
  grndmean90$Date),format = "%H:%M:%S")
grndmean90$day <- as.Date(grndmean90$Date)


Daily_average_grnd_per_sensor_90 <- grndmean90 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season

#Below -- exports as a csv to folder of choice 

#write.csv(Daily_average_grnd_per_sensor_90, "~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_90.csv")


```

```{r}
Daily_Averages_Ground_Temps_Summer_90 <- read_csv("~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_90.csv", na = "NA")

Daily_Averages_Ground_Temps_Summer_90$day <-as.POSIXct(Daily_Averages_Ground_Temps_Summer_90$day,format="%m/%d/%y",tz="UTC")

removeQuotes <- function(x) gsub("\"", "", x)
commatoperiod <- function(y) gsub(",",".", y)

Daily_Averages_Ground_Temps_Summer_90_noquote <- Daily_Averages_Ground_Temps_Summer_90 %>%
    mutate_if(is.character, removeQuotes)
Daily_Averages_Ground_Temps_Summer_90_noquote_nocomma <- Daily_Averages_Ground_Temps_Summer_90_noquote %>%
  mutate_if(is.character, commatoperiod)

Daily_Averages_Ground_Temps_Summer_90_noquote_nocomma$day<-as.POSIXct(Daily_Averages_Ground_Temps_Summer_90_noquote_nocomma$day,format="%m/%d/%y",tz="UTC")

#give it a shorter name...
CLEAN_BEO_TNHA_90 <- Daily_Averages_Ground_Temps_Summer_90_noquote_nocomma
CLEAN_BEO_TNHA_90$day <- as.POSIXct(CLEAN_BEO_TNHA_90$day, format="%m/%d/%y",tz="UTC")
CLEAN_BEO_TNHA_90$day <- as.Date(CLEAN_BEO_TNHA_90$day, format = "%m/%d/%y", tz="UTC")
CLEAN_BEO_TNHA_90$value <- as.numeric(CLEAN_BEO_TNHA_90$value)
round(CLEAN_BEO_TNHA_90$value, digits= 2)
```

```{r fig13, fig.height = 6, fig.width = 14}

#Daily_Averages_Ground_Temps_Summer_90 <- read_csv("Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer_90.csv", na = "NA")


CLEAN_BEO_TNHA_90 %>% filter(sensor == "TNHA-BASE" | sensor == "TNHA-SA" | sensor == "TNHA-SC" | sensor == "BEO-B06") %>%
  ggplot(aes(x=day, y=value, color = sensor), size = 0.9) +
  geom_line() +
  geom_point() +
  guides(color = guide_legend(override.aes = list(size=20)))+
  theme_bw()+ 
  theme(legend.title = element_text(size = 20),legend.text = element_text(size = 10)) +
  labs(title = "Summer 2022 daily average ground temperatures at 90cm depth, control vs. N&S facing sensors",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-2.5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 35),
        axis.title.x = element_text(size = 30),
        axis.title.y = element_text(size = 30)) +
  scale_color_manual(name = "Station",values =c("TNHA-BASE" = "#1B9E77", "TNHA-SC" ="#E7298A", "BEO-B06"= "#084594","TNHA-SA" = "#D95F02", labels = c("BEO (Control)", "TNHA-BASE (South Facing)", "TNHA-SA (South Facing)", "TNHA-SC (North Facing)"))) 

```

Dates before value >0
```{r}
before_thaw <- Daily_average_grnd_per_sensor_65 %>% filter(value > 0, sensor == "SSMH-SD")
```



Subsetting TNHA-SA, TNHA-SC, & BEO-B06 for 07-04-2022 through 09-01-2022 for means over season where no data gaps occur & same the start date 

```{r}
#10cm
subset10new <- CLEAN_BEO_TNHA_10[CLEAN_BEO_TNHA_10$day >= "2022-07-04" & CLEAN_BEO_TNHA_10$day <= "2022-09-01", ]
aggregate(value~sensor, subset10new, mean)
#30cm
subset30new <- Daily_average_grnd_per_sensor_30[Daily_average_grnd_per_sensor_30$day >= "2022-07-04" & Daily_average_grnd_per_sensor_30$day <= "2022-09-01", ]
aggregate(value~sensor, subset30new, mean)
#65cm
subset65new <- CLEAN_BEO_TNHA_65[CLEAN_BEO_TNHA_65$day >= "2022-07-04" & CLEAN_BEO_TNHA_65$day <= "2022-09-01", ]
aggregate(value~sensor, subset65new, mean)
#90cm
subset90new <- CLEAN_BEO_TNHA_90[CLEAN_BEO_TNHA_90$day >= "2022-07-04" & CLEAN_BEO_TNHA_90$day <= "2022-09-01", ]
aggregate(value~sensor, subset90new, mean)

```


####
Fall Data


*Ground Temperatures for all sites data prepping*
```{r}
# importing ground temp data for 6-19 through 9-20
grndtmp921_1119=read.csv("~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Fall_grndtmp_2023_04_06_12_19_52_UTC_1.csv")
names(grndtmp921_1119) <- sub("....C..RX3000_BRW1", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub("....C..RX3000_BRW4", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub("....C..RX3000_BRW5", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub("....C..RX3000_BRW6", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub(".RXW.GP6.", "", names(grndtmp921_1119))
grndtmp921_1119$Date <-as.POSIXct(grndtmp921_1119$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

 

*GROUND TEMPERATURE CSV ADJUSTMENTS, LONG DATAFRAME*
```{r}
grndtmp = gather(grndtmp921_1119, variable, response, Temperature.21198259.21206939.7:Temperature.21401800.21398576.17) %>%
  mutate(variable = fct_inorder(variable))
dfgrndtemp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(dfgrndtemp)[3] <- ("value")
colnames(dfgrndtemp)[5:7] <- c("station", "sensor", "depth")

dfgrndtemp$variable_1 <- NULL
dfgrndtemp$variable_5 <- NULL
dfgrndtemp$variable_6 <- NULL
dfgrndtemp$variable_7 <- NULL
dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth))
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))
dfgrndtemp$station <- as.factor(dfgrndtemp$station)
dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398587" = "TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))

dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))

dfgrndtemp <- dfgrndtemp %>% filter(!value == -888.88)

```


*****

*Daily average grnd temperature per sensor at station "X" at depth "X"*
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "10cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season

#Below -- exports as a csv to folder of choice 

#write.csv(Daily_average_grnd_per_sensor, "~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer.csv")

```


Dates before value >0
```{r}
before_thaw <- Daily_average_grnd_per_sensor %>% filter(value < 0 | value == 0 )
```



```{r fig14, fig.height = 6, fig.width = 14}
Daily_average_grnd_per_sensor %>% filter(sensor == "TNHA-BASE" | sensor == "TNHA-SA" | sensor == "TNHA-SC" |sensor == "TNHA-SB" | sensor == "TNHA-SD" | sensor == "BEO-B06") %>%
  ggplot(aes(x=day, y=value, color = sensor), size = 0.9) +
  geom_line() +
  geom_point() +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(title = "Fall 2022 daily average ground temperatures at 10cm depth, control vs. N&S facing sensors",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-15,-10,-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-15, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 23),
        axis.title.x = element_text(size = 30),
        axis.title.y = element_text(size = 30)) +
  scale_color_manual(name = "Station",values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E", "BEO-B06"= "#084594"))


  #scale_color_manual(name = "Station",values =c("TNHA-BASE" = "#1B9E77", "TNHA-SC" ="#E7298A", "BEO-B06"= "#084594","TNHA-SA" = "#D95F02", labels = c("BEO (Control)", "TNHA-BASE (South Facing)", "TNHA-SA (South Facing)", "TNHA-SC (North Facing)")))
```




Spring Data ##########

Ground Temperatures for all sites data prepping
```{r}
# importing ground temp data for 6-19 through 9-20
grndtmp619_920=read.csv("~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Spring_meteorological_grndtmp_2023_07_14_22_32_10_UTC_1.csv")
names(grndtmp619_920) <- sub("....C..RX3000_BRW1", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW4", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW5", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW6", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub(".RXW.GP6.", "", names(grndtmp619_920))
grndtmp619_920$Date <-as.POSIXct(grndtmp619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

```{r}
grndtmp = gather(grndtmp619_920, variable, response, Temperature.21198259.21206939.7_TNHA:Temperature.21401803.21398591.17_BEO) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
dfgrndtemp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(dfgrndtemp)[3] <- ("value")
colnames(dfgrndtemp)[5:7] <- c("station", "sensor", "depth")

dfgrndtemp$variable_5 <- NULL
dfgrndtemp$variable_6 <- NULL
dfgrndtemp$variable_7 <- NULL
dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth))
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))
dfgrndtemp$station <- as.factor(dfgrndtemp$station)
dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))

dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7_TNHA" = "3.5cm", "8_TNHA" = "10cm", "9_TNHA" = "20cm", "10_TNHA" = "30cm", "11_TNHA" = "40cm", "12_TNHA" = "50cm", "13_TNHA" = "55cm", "14_TNHA" = "65cm", "15_TNHA" = "75cm", "16_TNHA" = "85cm", "17_TNHA" = "90cm", "7_BEO" = "3.5cm", "8_BEO" = "10cm", "9_BEO" = "20cm", "10_BEO" = "30cm", "11_BEO" = "40cm", "12_BEO" = "50cm", "13_BEO" = "55cm", "14_BEO" = "65cm", "15_BEO" = "75cm", "16_BEO" = "85cm", "17_BEO" = "90cm","7_BUECI" = "3.5cm", "8_BUECI" = "10cm", "9_BUECI" = "20cm", "10_BUECI" = "30cm", "11_BUECI" = "40cm", "12_BUECI" = "50cm", "13_BUECI" = "55cm", "14_BUECI" = "65cm", "15_BUECI" = "75cm", "16_BUECI" = "85cm", "17_BUECI" = "90cm","7_SSMH" = "3.5cm", "8_SSMH" = "10cm", "9_SSMH" = "20cm", "10_SSMH" = "30cm", "11_SSMH"= "40cm", "12_SSMH" = "50cm", "13_SSMH" = "55cm", "14_SSMH" = "65cm", "15_SSMH" = "75cm", "16_SSMH" = "85cm", "17_SSMH" = "90cm"))

dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))

dfgrndtemp <- dfgrndtemp %>% filter(!value == -888.88)
```
*Daily average grnd temperature per sensor at station "X" at depth "X"*
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "65cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season
  

#Below -- exports as a csv to folder of choice 

#write.csv(Daily_average_grnd_per_sensor, "~/Desktop/UVA/RESEARCH/Barrow/csvfilesfromR/Daily_Averages_Ground_Temps_Summer.csv")

```





```{r fig14, fig.height = 6, fig.width = 14}
Daily_average_grnd_per_sensor %>% filter(sensor == "TNHA-BASE" | sensor == "TNHA-SA" | sensor == "TNHA-SC" | sensor == "BEO-B06") %>%
  ggplot(aes(x=day, y=value, color = sensor), size = 0.9) +
  geom_line() +
  geom_point() +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(title = "Spring 2023 daily average ground temperatures at 65cm depth, control vs. N&S facing sensors",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-20,-15,-10,-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-20, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 23),
        axis.title.x = element_text(size = 30),
        axis.title.y = element_text(size = 30)) +
  scale_color_manual(name = "Station",values =c("TNHA-BASE" = "#1B9E77", "TNHA-SC" ="#E7298A", "BEO-B06"= "#084594","TNHA-SA" = "#D95F02", labels = c("BEO (Control)", "TNHA-BASE (South Facing)", "TNHA-SA (South Facing)", "TNHA-SC (North Facing)")))
```
