---
title: "N-factor calculations"
author: "Mirella Shaban"
date: "2023-06-09"
output: html_document
---

*Formula for n-factor:*
nt = DDTs/DDTa = ∫(0 to thetas) (Ts-Tf)dt / ∫(0 to thetaa) (Ta-Tf)dt = Σ(0 to theta s) T overbar S / Σ(0 to theta a) T overbar a

nt = thaw season n-factor (dimensionless)
DDTs and DDTa = thaawing degree day sums for soil surface and air temp (deg C days)
Tf = temperaure of freezing point (0 deg C)
Ts and Ta = temp of soil surface and air (deg C)
theta s and theta a = periods of thawing season for soil surface and air (days)
T overbars and T overbar a = mean daily soil-surface and air temperatures (deg C)

DDTs and DDTa calculated by summing mean daily temperatures above 0 deg C

Our stab at the n-factor will be taking the slope of the line of XY plots of air temperature versus ground temperature 

*LOADING SCRIPT COMES FROM XY_MULTISIDE_COMPARISONS.RMD*

* Run the LOADANDCLEAN......RMD scripts for air temp and ground temp before running this! NO PLOTTING IN THIS SCRIPT, ONLY LM FITS!*

NAMES:
#############################
*SUMMER*
*2022
DAILY_AIR_SUMM22
SUMMERGRND22_ALLDEPTHS


*2023
DAILY_AIR_SUMM23
SUMMERGRND23_ALLDEPTHS

*2024
DAILY_AIR_SUMM24
SUMM24_ALLDEPTHS

##############################
*FALL*
*2022 
*2023
##############################


change name of value for air temp and ground temp to distinguish

```{r}
colnames(DAILY_AIR_SUMM22)[4] <- c("airtemp")
colnames(SUMMERGRND22_ALLDEPTHS)[4] <- c("grndtmp")
```


-------Merge Data Frames-------

```{r}
averages = DAILY_AIR_SUMM22 %>% full_join(SUMMERGRND22_ALLDEPTHS, by=c("date","sensor","station", "val", "fullname"))

```



TNHA
```{r}
XY_grnd_air= averages %>% filter(station == "TNHA", !fullname == "TNHA-SB", !fullname == "TNHA-SA", !fullname == "TNHA-SC", !fullname == "TNHA-SD", depth == "10cm") 

print(XY_grnd_air)

fit1 <- lm(grndtmp ~ airtemp, data = XY_grnd_air)
summary(fit1)
corfit <- cor(XY_grnd_air$airtemp,XY_grnd_air$grndtmp, use = 'complete.obs')

print("Correlation coefficint")
print(corfit)
```

SSMH
```{r}
XY_grnd_air= averages %>% filter(station == "SSMH", !fullname == "SSMH-BASE", !fullname == "SSMH-SE", !fullname == "SSMH-SD", !fullname == "SSMH-SB", !fullname == "SSMH-SH", depth == "10cm") 

print(XY_grnd_air)

fit1 <- lm(grndtmp ~ airtemp, data = XY_grnd_air)
summary(fit1)
corfit <- cor(XY_grnd_air$airtemp,XY_grnd_air$grndtmp, use = 'complete.obs')

print("Correlation coefficint")
print(corfit)
```

BEO
```{r}
XY_grnd_air= averages %>% filter(station == "BEO", depth == "10cm") 

print(XY_grnd_air)

fit1 <- lm(grndtmp ~ airtemp, data = XY_grnd_air)
summary(fit1)
corfit <- cor(XY_grnd_air$airtemp,XY_grnd_air$grndtmp, use = 'complete.obs')

print("Correlation coefficint")
print(corfit)
```