---
title: "Stacked_Averaged_Figures"
author: "Mirella Shaban"
date: "2024-01-24"
output: html_document
---

Loading packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape) 
library(naniar)
library(RColorBrewer)
library(clock)
```

Summer 2022
Ground Temperatures for all sites data prepping
```{r}
# importing ground temp data for 6-19 through 9-20
grndtmp619_920=read.csv("~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Summer_22_grndtmp_2023_11_14_20_50_46_UTC_1.csv")
names(grndtmp619_920) <- sub("....C..RX3000_BRW1", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW4", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW5", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW6", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub(".RXW.GP6.", "", names(grndtmp619_920))
grndtmp619_920$Date <-as.POSIXct(grndtmp619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

LONG DATAFRAMES
```{r}
grndtmp = gather(grndtmp619_920, variable, response, Temperature.21198259.21206939.7:Temperature.21401803.21398591.17) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))
dfgrndtemp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(dfgrndtemp)[3] <- ("value")
colnames(dfgrndtemp)[5:7] <- c("station", "sensor", "depth")

dfgrndtemp$variable_5 <- NULL
dfgrndtemp$variable_6 <- NULL
dfgrndtemp$variable_7 <- NULL
dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth))
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))
dfgrndtemp$station <- as.factor(dfgrndtemp$station)
dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398587"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))

dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))

dfgrndtemp <- dfgrndtemp %>% filter(!value == -888.88)
```


Daily average grnd temperature per sensor at BEO,TNHA,SSMH. Below we force NA values into each indiv. station df & stack them to form a comprehensive df with NA's (plots will disconnect when NA values are present)
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "10cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
TSB_NONA$depth <- "10cm"
TSB_NONA$depth <- as.factor(TSB_NONA$depth)
```

30cm
```{r}
grndmean1 <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "30cm") 

grndmean1 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean1$time <- format(as.Date(
  grndmean1$Date),format = "%H:%M:%S")
grndmean1$day <- as.Date(grndmean1$Date)


Daily_average_grnd_per_sensor <- grndmean1 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA30 <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA30 <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA30 <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA30 <- rbind(TNHA_NONA30,BEO_NONA30,SSMH_NONA30)
TSB_NONA30$depth <- "30cm"
TSB_NONA30$depth <- as.factor(TSB_NONA30$depth)
```

55

```{r}
grndmean2 <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "55cm") 

grndmean2 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean2$time <- format(as.Date(
  grndmean2$Date),format = "%H:%M:%S")
grndmean2$day <- as.Date(grndmean2$Date)


Daily_average_grnd_per_sensor <- grndmean2 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA55 <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA55 <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA55 <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA55 <- rbind(TNHA_NONA55,BEO_NONA55,SSMH_NONA55)
TSB_NONA55$depth <- "55cm"
TSB_NONA55$depth <- as.factor(TSB_NONA55$depth)
```

65

```{r}
grndmean3 <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "65cm") 

grndmean3 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean3$time <- format(as.Date(
  grndmean3$Date),format = "%H:%M:%S")
grndmean3$day <- as.Date(grndmean3$Date)


Daily_average_grnd_per_sensor <- grndmean3 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA65 <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA65 <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA65 <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA65 <- rbind(TNHA_NONA65,BEO_NONA65,SSMH_NONA65)
TSB_NONA65$depth <- "65cm"
TSB_NONA65$depth <- as.factor(TSB_NONA65$depth)
```

90
```{r}
grndmean4 <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "90cm") 

grndmean4 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean4$time <- format(as.Date(
  grndmean4$Date),format = "%H:%M:%S")
grndmean4$day <- as.Date(grndmean4$Date)


Daily_average_grnd_per_sensor <- grndmean4 %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)


Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")


date <- rep(seq(as.Date("2022-06-19"), as.Date("2022-09-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=75)
testdf <- data.frame(date,fullname,val)
TNHA_NONA90 <- left_join(testdf,TNHA_DAILY, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=75)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA90 <- left_join(testdf2,SSMH_DAILY, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=75)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA90 <- left_join(testdf3,BEO_DAILY, by=c("date","fullname"))

TSB_NONA90 <- rbind(TNHA_NONA90,BEO_NONA90,SSMH_NONA90)
TSB_NONA90$depth <- "90cm"
TSB_NONA90$depth <- as.factor(TSB_NONA90$depth)
```

Bind all depth dataframes together
```{r}
TSB_NONE_ALLDEPTHS <- rbind(TSB_NONA,TSB_NONA30,TSB_NONA55,TSB_NONA65,TSB_NONA90)
```


FIGURES

```{r fig13, fig.height = 30, fig.width = 26}
TSB_NONE_ALLDEPTHS %>% 
  ggplot(aes(x=date, y=value, color = fullname), size = 0.9) +
  geom_line() +
  geom_point() +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  facet_wrap(depth~.,nrow=5) +
  theme(legend.title = element_text(size = 25),legend.text = element_text(size = 20)) +
  labs(title = "Daily average ground temperatures at TNHA, SSMH, & BEO",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 30),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SC" ="#E7298A", "TNHA-SD"="#66A61E","BEO-B05"="#6BAED6", "BEO-B06"= "#084594","SSMH-BASE"="#807DBA", "SSMH-SA"="#A6761D", "SSMH-SB"= "#4292C6", "SSMH-SD"="#7A0177")) 


#scale_y_continuous(breaks=c(-10,-7.5,-5,-2.5,0,2.5,5), limits=c(-10, 5)) + ---> this is the scale for 90cm and 65cm....

```

Just N and S sensors at TNHA and SSMH
```{r fig14, fig.height = 23, fig.width = 24}
TSB_NONE_ALLDEPTHS %>% filter(fullname == "TNHA-SA" | fullname == "TNHA-SC" | fullname == "BEO-B06" | fullname == "SSMH-SA" | fullname == "SSMH-SB") %>% subset(depth == "10cm" | depth == "55cm" | depth == "90cm") %>%
  ggplot(aes(x=date, y=value, color = fullname), size = 0.9) +
  geom_line(size=0.9) +
  geom_point(size=1.5) +
  guides(color = guide_legend(override.aes = list(size=15)))+
  theme_bw() +
  facet_wrap(depth~.,nrow=5) +
  theme(legend.title = element_text(size = 25),legend.text = element_text(size = 20)) +
  labs(title = "Daily average ground temperatures, summer 2022",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid"),
        strip.text= element_text(size=35, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5, size = 30),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SC" ="#E7298A", "TNHA-SD"="#66A61E","BEO-B05"="#6BAED6", "BEO-B06"= "#084594","SSMH-BASE"="#807DBA", "SSMH-SA"="#A6761D", "SSMH-SB"= "#4292C6", "SSMH-SD"="#7A0177")) 



```


Fall 2022
Ground Temperatures for all sites data prepping
```{r}
# importing ground temp data for 6-19 through 9-20
grndtmp921_1119=read.csv("~/Desktop/UVA/RESEARCH/Barrow/Meteorological_Seasons_Data/Fall_22_grndtmp_2023_11_14_21_31_58_UTC_1.csv")
names(grndtmp921_1119) <- sub("....C..RX3000_BRW1", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub("....C..RX3000_BRW4", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub("....C..RX3000_BRW5", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub("....C..RX3000_BRW6", "", names(grndtmp921_1119))
names(grndtmp921_1119) <- sub(".RXW.GP6.", "", names(grndtmp921_1119))
grndtmp921_1119$Date <-as.POSIXct(grndtmp921_1119$Date,format="%m/%d/%y %H:%M",tz="UTC")
```




GROUND TEMPERATURE CSV ADJUSTMENTS, LONG DATAFRAME
```{r}
grndtmp = gather(grndtmp921_1119, variable, response, Temperature.21198259.21206939.7:Temperature.21401803.21398591.17) %>%
  mutate(variable = fct_inorder(variable))
dfgrndtemp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(dfgrndtemp)[3] <- ("value")
colnames(dfgrndtemp)[5:7] <- c("station", "sensor", "depth")

dfgrndtemp$variable_1 <- NULL
dfgrndtemp$variable_5 <- NULL
dfgrndtemp$variable_6 <- NULL
dfgrndtemp$variable_7 <- NULL
dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth))
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))
dfgrndtemp$station <- as.factor(as.character(dfgrndtemp$station))
dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398587"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))

dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))

dfgrndtemp <- dfgrndtemp %>% filter(!value == -888.88)

```


Daily average grnd temperature per sensor at BEO,TNHA,SSMH. Below we force NA values into each indiv. station df & stack them to form a comprehensive df with NA's (plots will disconnect when NA values are present)

10cm
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "10cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)

Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY10 <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY10<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY10<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")

date <- rep(seq(as.Date("2022-09-01"), as.Date("2022-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=92)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY10, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=92)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY10, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=92)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY10, by=c("date","fullname"))

TSB_NONA10 <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
TSB_NONA10$depth <- "10cm"
TSB_NONA10$depth <- as.factor(TSB_NONA10$depth)
#TSB_NONA is our finalized df for plotting with all 3 of 4 stations having a measurement or NA value for each daily average. Adjust the depth of grndmean, and adjust the season via the csv uploaded to this script, to make different figures
```

30cm
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "30cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)

Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY30 <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY30<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY30<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")

date <- rep(seq(as.Date("2022-09-01"), as.Date("2022-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=92)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY30, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=92)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY30, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=92)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY30, by=c("date","fullname"))

TSB_NONA30 <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
TSB_NONA30$depth <- "30cm"
TSB_NONA30$depth <- as.factor(TSB_NONA30$depth)
#TSB_NONA is our finalized df for plotting with all 3 of 4 stations having a measurement or NA value for each daily average. Adjust the depth of grndmean, and adjust the season via the csv uploaded to this script, to make different figures
```

55cm
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "55cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)

Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY55 <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY55<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY55<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")

date <- rep(seq(as.Date("2022-09-01"), as.Date("2022-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=92)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY55, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=92)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY55, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=92)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY55, by=c("date","fullname"))

TSB_NONA55 <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
TSB_NONA55$depth <- "55cm"
TSB_NONA55$depth <- as.factor(TSB_NONA55$depth)
#TSB_NONA is our finalized df for plotting with all 3 of 4 stations having a measurement or NA value for each daily average. Adjust the depth of grndmean, and adjust the season via the csv uploaded to this script, to make different figures
```

65cm
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "65cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)

Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY65 <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY65<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY65<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")

date <- rep(seq(as.Date("2022-09-01"), as.Date("2022-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=92)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY65, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=92)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY65, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=92)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY65, by=c("date","fullname"))

TSB_NONA65 <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
TSB_NONA65$depth <- "65cm"
TSB_NONA65$depth <- as.factor(TSB_NONA65$depth)
#TSB_NONA is our finalized df for plotting with all 3 of 4 stations having a measurement or NA value for each daily average. Adjust the depth of grndmean, and adjust the season via the csv uploaded to this script, to make different figures
```

90cm
```{r}
grndmean <- dfgrndtemp %>% subset() %>% select(value,sensor,Date,depth) %>% filter( !value == "-888.8", depth == "90cm") 

grndmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

grndmean$time <- format(as.Date(
  grndmean$Date),format = "%H:%M:%S")
grndmean$day <- as.Date(grndmean$Date)


Daily_average_grnd_per_sensor <- grndmean %>% 
  group_by(sensor, day) %>%
  aggregate(value ~ sensor + day, mean)

Daily_average_grnd_per_sensor <-  cSplit(Daily_average_grnd_per_sensor, 'sensor', sep="-", direction = "wide")
colnames(Daily_average_grnd_per_sensor)[1:4] <- c("date","value","station","sensor")
Daily_average_grnd_per_sensor$fullname <- paste(Daily_average_grnd_per_sensor$station,Daily_average_grnd_per_sensor$sensor,sep="-")


TNHA_DAILY90 <- Daily_average_grnd_per_sensor %>% filter(station == "TNHA")
SSMH_DAILY90<- Daily_average_grnd_per_sensor %>% filter(station == "SSMH")
BEO_DAILY90<- Daily_average_grnd_per_sensor %>% filter(station == "BEO")

date <- rep(seq(as.Date("2022-09-01"), as.Date("2022-12-01"), by="day"),5)
val <- rep(1,length(date))
fullname <- rep(c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD"),each=92)
testdf <- data.frame(date,fullname,val)
TNHA_NONA <- left_join(testdf,TNHA_DAILY90, by=c("date","fullname"))

fullname <- rep(c("SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD"),each=92)
testdf2 <- data.frame(date,fullname,val)
SSMH_NONA <- left_join(testdf2,SSMH_DAILY90, by=c("date","fullname"))

fullname <- rep(c("BEO-B05", "BEO-B06","FAKEVAR", "FAKEVAR2", "FAKEVAR3"),each=92)
testdf3 <- data.frame(date,fullname,val)
BEO_NONA <- left_join(testdf3,BEO_DAILY90, by=c("date","fullname"))

TSB_NONA90 <- rbind(TNHA_NONA,BEO_NONA,SSMH_NONA)
TSB_NONA90$depth <- "90cm"
TSB_NONA90$depth <- as.factor(TSB_NONA90$depth)
#TSB_NONA is our finalized df for plotting with all 3 of 4 stations having a measurement or NA value for each daily average. Adjust the depth of grndmean, and adjust the season via the csv uploaded to this script, to make different figures
```
#change names of outputs above ^

Bind all depth dataframes together
```{r}
TSB_NONE_ALLDEPTHS <- rbind(TSB_NONA10,TSB_NONA30,TSB_NONA55,TSB_NONA65,TSB_NONA90)
```

```{r fig13, fig.height = 23, fig.width = 24}
TSB_NONE_ALLDEPTHS %>% filter(fullname == "TNHA-SA" | fullname == "TNHA-SC" | fullname == "SSMH-SA"| fullname == "SSMH-SB" | fullname == "BEO-B06") %>% subset(depth == "10cm" | depth == "55cm" | depth == "90cm") %>%
  ggplot(aes(x=date, y=value, color = fullname), size = 0.9) +
  geom_line(size=0.9) +
  geom_point(size=1.5) +
  guides(color = guide_legend(override.aes = list(size=15)))+
  theme_bw() +
  facet_wrap(depth~.,nrow=5) +
  theme(legend.title = element_text(size = 25),legend.text = element_text(size = 20)) +
  labs(title = "Daily average ground temperatures, fall 2022",y="Temperature (°C)",x="Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-5, 15)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20),
        strip.text= element_text(size=35, face = "bold"),
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 30),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25)) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SC" ="#E7298A", "TNHA-SD"="#66A61E","BEO-B05"="#6BAED6", "BEO-B06"= "#084594","SSMH-BASE"="#807DBA", "SSMH-SA"="#A6761D", "SSMH-SB"= "#4292C6", "SSMH-SD"="#7A0177")) 


#scale_y_continuous(breaks=c(-10,-7.5,-5,-2.5,0,2.5,5), limits=c(-10, 5)) + ---> this is the scale for 90cm and 65cm....

```




Winter 2022-2023


Spring 2023


Summer 2023

Fall 2023


