---
title: "Air Temp, Ground Temp, Solar Radiation, VWC & Soil Temp  - Summer Season (June-September) 2022, Barrow, AK"
author: "Mirella Shaban"
date: "11/07/2022", last updated "01/23/2023"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
 Loading packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape) # basically like text to columns
library(naniar)
library(RColorBrewer)
```


Ground Temperatures for all sites data prepping
```{r}
# importing ground temp data for 6-19 through 9-20
grndtmp619_920=read.csv("C:/Users/qwe2qh/UVA/Utqiagvik/Season_GRNDTMP_6_19_22_9_20_22_2022_09_21_16_19_26_UTC_1.csv")
 
# data cleaning for dataset
names(grndtmp619_920) <- sub("....C..RX3000_BRW1", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW4", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW5", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub("....C..RX3000_BRW6", "", names(grndtmp619_920))
names(grndtmp619_920) <- sub(".RXW.GP6.", "", names(grndtmp619_920))
grndtmp619_920$Date <-as.POSIXct(grndtmp619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```


Ground Moisture for all sites data prepping
```{r}
mst619_920=read.csv("C:/Users/qwe2qh/UVA/Utqiagvik/Season_VWC_6_19_22_9_20_22_2022_09_21_16_11_28_UTC_1.csv")

names(mst619_920) <- sub("...m.3.m.3..RX3000_BRW1", "", names(mst619_920))
names(mst619_920) <- sub("...m.3.m.3..RX3000_BRW4", "", names(mst619_920))
names(mst619_920) <- sub("...m.3.m.3..RX3000_BRW5", "", names(mst619_920))
names(mst619_920) <- sub("...m.3.m.3..RX3000_BRW6", "", names(mst619_920))
names(mst619_920) <- sub("..RXW.GP6.", "", names(mst619_920))
mst619_920$Date <-as.POSIXct(mst619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

Solar radiation for all sites data prepping
```{r}
solarrad619_920 = read.csv("C:/Users/qwe2qh/UVA/Utqiagvik/Season_SOLARAD_6_19_22_9_20_22_2022_09_21_16_27_20_UTC_1.csv")

names(solarrad619_920) <- sub("olar.Radiation..RXW.LIB", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("olar.Radiation..S.LIB", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("S.", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW1", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW4", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW5", "" ,names(solarrad619_920))
names(solarrad619_920) <- sub("...W.m.2..RX3000_BRW6", "" ,names(solarrad619_920))
solarrad619_920$Date <- as.POSIXct(solarrad619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

#Upload air temp data and clean
```{r}
airtemp619_920 =read.csv("C:/Users/qwe2qh/UVA/Utqiagvik/Season_AIRTMP_6_19_22_9_20_22_2022_09_21_16_21_48_UTC_1.csv")

names(airtemp619_920) <- sub("Temperature..RXW.THC.", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("Temperature..S.THC.", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("S.", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW1", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW4", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW5", "" ,names(airtemp619_920))
names(airtemp619_920) <- sub("....C..RX3000_BRW6", "" ,names(airtemp619_920))
airtemp619_920$Date <- as.POSIXct(airtemp619_920$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

### Remove Sneaky Air Temperature!!
```{r}
grndtmp619_920 = grndtmp619_920[,-14]
```

## Return Sensor ID of Each Sensor by Placement Number
NOTE: the sensor IDs do not match up for each serial number - this is an attempt
to identify inconsistencies.
**Not necessary to run, just to understand how many sensors there are!
```{r}
c = 3
sensorID = NULL
columnNum = NULL
while(c <= ncol(grndtmp619_920)){
  sensorID = append(sensorID, names(grndtmp619_920)[c])
  columnNum = append(columnNum, c)
  c = c + 11
}
location = c(rep("TNHA",5),rep("SSMH",4),rep("BUECI",8),rep("BEO",2))
ID = c("BASE B07", "SC", "SD?", "SA", "SB", "SB", "BASE B05?/SD", "SA", "BASE B05??", "SC", "SF-01", "SE", "SB", "SD", "BASE B05", "SA", "SF-02", "BASE B06", "BASE B05")
sensors = data.frame(columnNum, sensorID, location, ID)
```

#removing ground temp zeroes occuring at multiple consecutive depths for loop
```{r}
c = 3 #this is the first column of the sensors
zeroes = 0
while(c <= ncol(grndtmp619_920)-10){
  #select one sensor at a time at 11 depths
  sensorGroup = grndtmp619_920 %>% select(2,c, c+1, c+2, c+3, c+4, c+5, c+6, c+7, c+8, c+9, c+10)
  
  #for loop: goes through each row, checking for zeroes
  for(r in 1:nrow(sensorGroup)) {
    zeroes = 0
    col = 2
    
    #while loop: counts zeroes if they are present in the current row
    #could have used a for loop, but i was experiencing unrelated errors so now it's a while loop 
    while(col <= 12){
      if(sensorGroup[r,col] == 0 | is.na(sensorGroup[r,col])) { #will not return T/F if value is N/A
        if(!is.na(sensorGroup[r,col])){
          zeroes = zeroes + 1 #count a zero
        }
      }
      col = col + 1
    }
    if(zeroes >= 3){ #if we detect a zero at 3 different depths, we count this as a false zero
      for(columnn in 2:ncol(sensorGroup)){ #go through each column, if zero then replace with NA
      if(sensorGroup[r,columnn] == 0 | is.na(sensorGroup[r,columnn])) { 
        #again, will not work if there is N/A ^
        grndtmp619_920[r,columnn+c] = NA
      }
    }
    }
  }
  c = c + 11 #each time, it increases by the number of depths (11)
}
```

--- End of basic data upload and cleaning ---

clean up ground moisture data 
```{r}
#pivot longer vs gather
dataframelong = gather(mst619_920, variable, response, Water.Content21198259.21206939.1:Water.Content21401803.21398591.6) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

# go into this dataframe, variable = all column naames, response is the numbers that will fill that column, first column to last column, etc.

df = cSplit(dataframelong, 'variable', sep=".", direction = "wide") # texttocols, 
colnames(df)[5:7] <- c("station", "sensor", "depth")
df$variable_1 <- NULL # gets rid of this column
df$depth <- as.factor(as.character(df$depth)) #taking depth from numbers and turning it into a character, then to factor (cant go from number to factor) -- when we want to sort by depth it wont take a number, only a factor **
df$station <- as.factor(df$station)
# df$sensor <- factor(df$sensor, levels = c("TNHA-BASE","TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD",
#                                     "SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD",
#                                     "BUCEI-BASE", "BUCEI-SA", "BUCEI-SB", "BUCEI-SC", "BUCEI-SD", "BUCEI-SE",
#                                     "BUCEI-SF.01", "BUCEI-SF.02",
#                                     "BEO-B05", "BEO-B06"))
df$sensor <- as.factor(as.character(df$sensor))

df$station <- revalue(df$station, c("Content21198259" = "TNHA", "Content21401800" = "BUECI", "Content21401801" = "SSMH", "Content21401803" = "BEO")) #renaming the stations
df$depth <- revalue(df$depth, c("1" = "0:15cm", "2" = "15:30cm", "3" = "30:45cm", "4" = "45:60cm", "5" = "60:75cm", "6" = "75:90cm"))

df$sensor <- revalue(df$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUCEI-SA", "21398583" = "BUECI-SB", "21393042" = "BUCEI-SC", "21398584" = "BUCEI-SD", "21398579" = "BUCEI-SE", "21398578" = "BUCEI-SF.01", "21398598" = "BUCEI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05")) 

#TNHA station B --- imposter???

#TNHA-SB : this is an assumption that this serial number is at this station. Need on the ground confirmation

# can do this same process for the seria nubers for stations, as well as for the depth columns 
# 0 = NA , left side is whats in my station column and im changing it to be NA
```

# Removing 0's (only run this chunk if you would like perfect values of 0 removed from the VWC data)
```{r}
df[df == 0.000] <- NA
```

TEMPLATE: ALL depths VWC (select sensors w/  %>% filter(!sensor =="" to remove) etc.)
```{r fig1, fig.height = 8, fig.width = 18}
#All locations, separated by station (columns) and depth (rows)
ggplot(subset(df, depth %in% c("0:15cm", "15:30cm", "30:45cm", "45:60cm", "60:75cm", "75:90cm")), aes(x=Date, y= response, color = sensor)) + 
  geom_point(size = 0.5) +
  facet_grid(depth~station) + #takes station as x value, tilda says " station by what, period means by nothing ## think anova, splitting up the data by these two parameters ,, so sensor is still my color and each 
  theme_bw() +
  labs(title= "Ground Moisture", y="Volumetric Water Content ("~m^3~"/"~m^3~")",color="Station", caption = "Mirella Shaban") + # fix title and names of axes
  #google relabel facet labels ***
  scale_x_datetime(date_breaks = "1 weeks") +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_colour_hue(breaks=c("TNHA-BASE","TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD",
                               "SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD",
                               "BUCEI-BASE", "BUCEI-SA", "BUCEI-SB", "BUCEI-SC", "BUCEI-SD", "BUCEI-SE",
                               "BUCEI-SF.01", "BUCEI-SF.02",
                               "BEO-B05", "BEO-B06"))


# line 92 --  we're subsetting by depth specifically instead of looking at the whole dataframe -- id change depth to station and the c() values to the names of the stations --- look up conditional in R cheatsheet
```
TEMPLATE: SELECT STATIONS AND SENSORS VWC AT ALL DEPTHS
```{r fig2, fig.height = 18, fig.width = 20}
df1 <- df %>% filter(!station == "BUECI", !station == "BEO", !station == "SSMH", !sensor == "TNHA-BASE", !sensor == "TNHA-SB", !sensor == "TNHA-SA", !sensor == "TNHA-SC")
ggplot(subset(df1, depth %in% c("0:15cm", "15:30cm", "30:45cm", "45:60cm", "60:75cm", "75:90cm")), aes(x=Date, y= response, color = sensor)) + 
  geom_point(size = 0.5) +
  guides(color = guide_legend(override.aes = list(size=10))) +
  facet_grid(depth~station) + #takes station as x value, tilda says " station by what, period means by nothing ## think anova, splitting up the data by these two parameters ,, so sensor is still my color and each 
  theme_bw() +
  labs(title= "Ground Moisture", y="Volumetric Water Content ("~m^3~"/"~m^3~")",color="Station") + # fix title and names of axes
  #google relabel facet labels ***
  scale_x_datetime(date_breaks = "1 weeks") +
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 30),plot.title = element_text(hjust = 0.5, size = 30), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 20),
        axis.title = element_text(size = 20)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.key.height = unit(1,'cm')) +
  scale_colour_hue(breaks=c("TNHA-BASE","TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD","SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD",
                               "BUCEI-BASE", "BUCEI-SA", "BUCEI-SB", "BUCEI-SC", "BUCEI-SD", "BUCEI-SE",
                               "BUCEI-SF.01", "BUCEI-SF.02",
                               "BEO-B05", "BEO-B06"))

```
ONLY TNHA & BEO VWC (CHANGE STATION W/ !STATION == "")
```{r fig7, fig.height = 8, fig.width = 18}
df5 <- df %>% filter(!station == "BUECI", 
                     !station == "SSMH")
ggplot(subset(df5, depth %in% c("0:15cm","15:30cm","30:45cm", "75:90cm")), 
       aes(x=Date, y= response, color = sensor)) + 
  geom_point(size = 0.5) +
  guides(color = guide_legend(override.aes = list(size=10))) +
  facet_grid(depth~station) +
  theme_bw() +
  labs(title= "Ground Moisture, (VWC)", 
       y="Volumetric Water Content ("~m^3~"/"~m^3~")",
       color="Station", caption = "Mirella Shaban") + 
  scale_x_datetime(date_breaks = "1 weeks") +
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, 
                                 hjust = 1.0, 
                                 size = 15), 
        axis.line = element_line(colour = "black", 
                                 size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 30),
        plot.title = element_text(hjust = 0.5, size = 30), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 20),
        axis.title = element_text(size = 20)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.key.height = unit(1,'cm')) +
  scale_colour_hue(breaks=c("TNHA-BASE","TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD",
                               "BEO-B05", "BEO-B06"))

```

GROUND TEMPERATURE CSV ADJUSTMENTS, LONG DATAFRAME
```{r}
grndtmp = gather(grndtmp619_920, variable, response, Temperature.21198259.21206939.7:Temperature.21401803.21398591.17) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

# go into this dataframe, variable = all column naames, response is the numbers that will fill that column, first column to last column, etc.

dfgrndtemp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") # texttocols, 
colnames(dfgrndtemp)[3] <- ("value")
colnames(dfgrndtemp)[5:7] <- c("station", "sensor", "depth")

dfgrndtemp$V212 <- NULL
dfgrndtemp$V213 <- NULL
dfgrndtemp$variable_1 <- NULL
dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth)) #taking depth from numbers and turning it into a character, then to factor (cant go from number to factor) -- when we want to sort by depth it wont take a number, only a factor **
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))
dfgrndtemp$station <- as.factor(dfgrndtemp$station)

dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUCEI-SA", "21398583" = "BUECI-SB", "21393042" = "BUCEI-SC", "21398584" = "BUCEI-SD", "21398579" = "BUCEI-SE", "21398578" = "BUCEI-SF.01", "21398598" = "BUCEI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))

dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

#print(levels(data$column)) 
dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))

```

ALL SITE GROUND TEMPERATURE
```{r fig3, fig.height = 18, fig.width = 20}
ggplot(subset(dfgrndtemp, depth %in% c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm")),
       aes(x=Date, y= value, color = sensor)) + 
  geom_point(size = 0.5) +
  guides(color = guide_legend(override.aes = list(size=10))) +
  facet_grid(depth~station) + #takes station as x value, tilda says " station by what, period means by nothing ## think anova, splitting up the data by these two parameters ,, so sensor is still my color and each 
  theme_bw() +
  labs(y="Temperature (°C)",color="Station", caption = "Mirella Shaban") +
  scale_x_datetime(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-6, 14)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_colour_hue(breaks=c("TNHA-BASE","TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD",
                               "SSMH-BASE", "SSMH-SA", "SSMH-SB", "SSMH-SC", "SSMH-SD",
                               "BUCEI-BASE", "BUCEI-SA", "BUCEI-SB", "BUCEI-SC", "BUCEI-SD", "BUCEI-SE",
                               "BUCEI-SF.01", "BUCEI-SF.02",
                               "BEO-B05", "BEO-B06"))

#add a color palette for BUECI in scale_color_manual

```

SELECT SITES GROUND TEMP (HERE: TNHA, SSMH, BEO)
```{r fig4, fig.height = 18, fig.width = 20}

df2 <- dfgrndtemp %>% filter(!station == "BUECI")

ggplot(subset(df2, depth %in% c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm") & value >.002 | value < -0.002),
       aes(x=Date, y= value, color = sensor)) + 
  geom_miss_point(size = 0.5) +
  guides(color = guide_legend(override.aes = list(size=10))) +
  facet_grid(depth~station) + #takes station as x value, tilda says " station by what, period means by nothing ## think anova, splitting up the data by these two parameters ,, so sensor is still my color and each 
  theme_bw() +
  labs(y="Temperature (°C)",color="Station", caption = "Mirella Shaban") +
  scale_x_datetime(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-6, 14)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
   scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E",
                               "SSMH-BASE"="#807DBA", "SSMH-SA"="#A6761D", "SSMH-SB"= "#4292C6", "SSMH-SC"= "#DD3497", "SSMH-SD"="#7A0177", "BEO-B05"="#6BAED6", "BEO-B06"= "#084594"))

#aggregate based on depth + station (ggplot conditional for TNHA at all depths on one figure)
```
TEMPLATE: ALL depths ground temp (select sensors, HERE:BEO)
```{r fig82, fig.height = 30, fig.width = 25}

df2 <- dfgrndtemp %>% filter(!station == "BUECI", !station == "BEO", !station == "SSMH")

ggplot(subset(df2, depth %in% c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm") & value >.002 | value < -0.002),
       aes(x=Date, y= value, color = sensor)) + 
  geom_miss_point(size = 0.5) +
  guides(color = guide_legend(override.aes = list(size=10))) +
  facet_grid(depth~station) + #takes station as x value, tilda says " station by what, period means by nothing ## think anova, splitting up the data by these two parameters ,, so sensor is still my color and each 
  theme_bw() +
  labs(y="Temperature (°C)",color="Station") +
  scale_x_datetime(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-6, 14)) +
   theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 30),plot.title = element_text(hjust = 0.5, size = 30), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 20),
        axis.title = element_text(size = 20)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.key.height = unit(1,'cm')) +
   scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E"
                               ))

#aggregate based on depth + station (ggplot conditional for TNHA at all depths on one figure)
```



#DENSITY PLOT: Averaged TNHA ground temp data, all depths on one figure
```{r fig5, fig.height = 8, fig.width = 20}
dfgrndtemp %>%
  subset(station == c("TNHA") & value >0.002 | value< -0.003)  %>% #GET RID OF BOGUS DATA
   aggregate(value ~ depth + Date + station, mean) %>%
  ggplot()+
  geom_density_2d(aes(x=Date, y= value, color = depth), size = 1) + 
  guides(color = guide_legend(override.aes = list(size=10))) +
  theme_bw() +
  labs(y="Temperature (°C)",color="Station") +
  scale_x_datetime(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-6, 14)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0), axis.line = element_line(colour = "black", size =0.5,
                                                                                    linetype = "solid"))

```
#TNHA SSMH & BEO ground temp certain depths.
```{r}
df3 <- dfgrndtemp %>% filter(!station == "BUECI") %>% filter(!sensor == "21398595") %>% subset(depth %in% c("10cm","30cm","65cm","90cm"))
```

```{r fig6, fig.height = 8, fig.width = 18}
ggplot(subset(df3, depth %in% c("10cm","30cm","65cm","90cm") & value >.002 | value < -0.002),
       aes(x=Date, y= value, color = sensor)) + 
  geom_point(size = 0.5) +
  guides(color = guide_legend(override.aes = list(size=10))) +
  facet_grid(depth~station) + 
  theme_bw() + theme(strip.text = element_text(face = "bold"))+
  labs(y="Temperature (°C)", color="Sensor") +
  scale_x_datetime(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-5,-2.5,0,2.5,5,7.5,10,12.5,15), limits=c(-6, 15)) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 30),plot.title = element_text(hjust = 0.5, size = 30), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 20),
        axis.title = element_text(size = 20)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E",
                               "SSMH-BASE"="#807DBA", "SSMH-SA"="#A6761D", "SSMH-SB"= "#4292C6", "SSMH-SC"= "#DD3497", "SSMH-SD"="#7A0177", "BEO-B05"="#6BAED6", "BEO-B06"= "#084594"))


#COLOR PALETTES:
#TNHA = brewer.pal(n = 8, name = 'Dark2')
#SSMH = brewer.pal(n = 8, name = 'Purples'), plus color #A6761D from Dark 2 pallet for SA
#BEO = brewer.pal(n = 8, name = 'Blues')

```
TNHA only VWC
```{r fig8, fig.height = 15, fig.width = 18}
df4 <- df %>% filter(!station == "BUECI", 
                             !station =="BEO", 
                             !station == "SSMH")

ggplot(subset(df4, 
              depth %in% c("0:15cm","30:45cm","60:75cm", "75:90cm")), 
       aes(x=Date, y= response, color = sensor)) + 
  geom_point(size = 0.5) +
  guides(color = guide_legend(override.aes = list(size=10))) +
  facet_grid(depth~station) + 
  theme_bw() +
  labs(title= "Ground Moisture", y="Volumetric Water Content ("~m^3~"/"~m^3~")",color="Station") + 
  scale_x_datetime(date_breaks = "1 weeks") +
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 15),plot.title = element_text(hjust = 0.5, size = 30), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 20),
        axis.title = element_text(size = 20)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 20),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1,'cm')) +
  scale_colour_hue(breaks=c("TNHA-BASE","TNHA-SA", "TNHA-SC", "TNHA-SD"))
#SB removed as there is no data tied to it in hobolink
```

SOLAR RADIATION : LONG DATAFRAME

```{r fig9, fig.heigt = 14, fig.width = 18}


solarrad = gather(solarrad619_920, variable, response, '21401801.21362319.1':'21198259.21398620.1') %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

solarrad= cSplit(solarrad, 'variable', sep=".", direction = "wide") # texttocols, 
solarrad$variable_3 <- NULL
colnames(solarrad)[3:5] <- c("value", "station", "sensor")
solarrad$value <- as.numeric(as.character(solarrad$value)) 
solarrad$sensor <- as.factor(as.character(solarrad$sensor))
solarrad$station <- as.factor(solarrad$station)

solarrad$station <- revalue(solarrad$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO"))





solarrad$sensor <- revalue(solarrad$sensor, c("21390411" = "BUECI-BASE", "21398618" = "BUCEI-SA", "21398624" = "BUECI-SB", "21362313" = "BUCEI-SC", "21362316" = "BUCEI-SD", "21362320" = "BUCEI-SE", "21398578" = "BUCEI-SF.01", "21398598" = "BUCEI-SF.02", 
                                              "21390413" ="SSMH-BASE", "21398622" = "SSMH-SA", "21362319" = "SSMH-SB", "21166008" = "SSMH-SC", "21393049" = "SSMH-SD" ,
                                              "21176526" = "TNHA-BASE", "21398620" = "TNHA-SA", "21398616"="TNHA-SB","21362315" = "TNHA-SD", "21362317" = "TNHA-SC", 
                                              "21390415" = "BEO-Base"))

```

AVERAGING OF SOLAR RAD DATA, AVERAGE DAILY VALUES

```{r}
#TNHA AVG DAILY SOLAR ALL SENSORS COMBINED
avgdailysolar <- solarrad %>% filter(!station == "BEO", !station == "SSMH", !station == "BUECI") 
  aggregate(list(Value = avgdailysolar$value), 
          list(dailyavg = cut(avgdailysolar$Date, "1 day")), 
          mean)

#TNHA AVG DAILY SOLAR, SELECT SENSOR  
avgdailysolar <- solarrad %>% filter(!station == "BEO", !station == "SSMH", !station == "BUECI", !sensor == "TNHA-SD", !sensor == "TNHA-BASE", !sensor == "TNHA-SA", !sensor == "TNHA-SB") 
aggregate(list(Value = avgdailysolar$value), 
          list(dailyavg = cut(avgdailysolar$Date, "1 day")), 
          mean)


```
AVG DAILY SOLAR FIGURE FOR SELECT SENSOR (BASED ON WHATS IN CHUNK ABOVE)
```{r fig10, fig.height = 6, fig.width = 16}

avgdailysolar %>%
  filter(!value == 0) %>%
  aggregate(value ~ Date + sensor, mean) %>%
  ggplot() +
  geom_line(aes(x=Date, y=value, color = sensor), size = 0.9) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(y="W /"~m^2~"",color="Station") +
  scale_x_datetime(date_breaks = "1 weeks") +
  scale_y_continuous(limits=c(0, 1000)) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E"))



```

MAX MIN MEANS FOR AIRTEMP:  variable & specified station
```{r}
#library(tibbletime)
solarrad2 <- solarrad %>% subset(station == c("SSMH")) %>% select(value,sensor,Date) %>% filter(!value == "999") %>% filter


solarrad2 %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)


solarrad2 %>%
  mutate(hourly = hour(time)) %>%
  group_by(hourly) %>%
  summarize(mean_X1 = mean(X1))

solarrad2_1 <- ddply(solarrad2, 'sensor', summarize, value =mean(value), Date=head(solarrad$Date,1), Value=head(solarrad$value,1))



solarrad2$time <- as.factor(solarrad2$time)


solarrad2$time <- format(as.Date(
  solarrad2$Date),format = "%H:%M:%S")
solarrad2$day <- as.Date(solarrad2$Date)

solarrad2 %>% filter(time == 11:00:00 : 16:00:00)


solarrad23 <- solarrad2 %>%
  group_by(sensor) %>%
  mutate(
    Maxval = max(value, na.rm = T),
    Minval = min(value, na.rm = T),
    mean_val = mean(value, na.rm = T)
  ) %>%
  arrange(sensor)
```


#air temp plots ((EDIT))
```{r}
airtemp = gather(airtemp619_920, variable, response, "21198259.21380919.1":"21401803.21397541.1") %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

airtemp= cSplit(airtemp, 'variable', sep=".", direction = "wide") 
airtemp$variable_3 <- NULL
colnames(airtemp)[3:5] <- c("value", "station", "sensor")
airtemp$value <- as.numeric(as.character(airtemp$value)) 
airtemp$sensor <- as.factor(as.character(airtemp$sensor))
airtemp$station <- as.factor(airtemp$station)

airtemp$station <- revalue(airtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) 


airtemp$sensor <- revalue(airtemp$sensor, c("21390851" = "BUECI-BASE", "21398671" = "BUCEI-SA", "21362254" = "BUECI-SB", "21398668" = "BUCEI-SC", "21398660" = "BUCEI-SD", "21187245" = "BUCEI-SE",
                                            "21390849" ="SSMH-BASE","21393049" = "SSMH-SD", "21398670" = "SSMH-SA", "21398665" = "SSMH-SB",
                                            "21218018" = "TNHA-BASE", "21380919" = "TNHA-SA" , "21398676"="TNHA-SB","21398666" = "TNHA-SD", "21398674" = "TNHA-SC", "21397541" = "BEO-BASE"))
#TNHA-A and TNHA-B are GUESSES - need metadata on ground
```

#Daily average air temperature per sensor at station X
```{r}
tnha_airmean <- airtemp %>% subset(station == c("SSMH")) %>% select(value,sensor,Date) %>% filter(!value == "0.0") 

tnha_airmean %>%
  mutate(date = day(Date))  %>%
  group_by(date,sensor,value) %>%
  summarize(mean_X1 = mean(value)) %>% 
  aggregate( value ~ sensor, mean ) %>%
  group_by(sensor)

tnha_airmean$time <- format(as.Date(
  tnha_airmean$Date),format = "%H:%M:%S")
tnha_airmean$day <- as.Date(tnha_airmean$Date)


bigbet <- tnha_airmean %>% 
  group_by(sensor, day) %>% filter(!sensor == '21398664') %>%
  aggregate(value ~ sensor + day, mean)
#this gives us the daily mean air temp for each sensor at TNHA in summer season

```
#Weekly Max,Min,Median averages 
```{r fig20, fig.heigt = 4, fig.width = 8}

maxminmedair<- tnha_airmean %>% filter(!sensor == '21398664') %>%
  group_by(day,sensor) %>%
  summarise(Max = max(value, na.rm = T),Min = min(value, na.rm = T), p = median(value,na.rm = T))
#above is daily max, min, med per sensor

maxminmedair$Week_Day <- as.numeric(format(maxminmedair$day, format='%w'))

# Adjust end-of-week date (first saturday from the original Date)
maxminmedair$End_of_Week <- maxminmedair$day + (6 - maxminmedair$Week_Day)


Weekly_Airtemp_MaxMinMed <- maxminmedair %>% group_by(sensor, time = as.Date(End_of_Week)) %>% summarise(across(c(Max, Min, p), mean))



Weekly_Airtemp_MaxMinMed %>% 
ggplot(aes(x = time, y = p, ymin = Min, ymax = Max, color = sensor)) + 
    geom_pointrange() +
  geom_errorbar() +
  guides(color = guide_legend(override.aes = list(size=0.5)))+
  theme_bw() +
  labs(y="Temperature (°C)",x="End of Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E"))

# make it geom bar and gridwrap or facetwrap it   
```


```{r}

BASEMAXTEMP <- Weekly_Airtemp_MaxMinMed %>% filter(!sensor=="TNHA-SB", !sensor == "TNHA-BASE", !sensor == "TNHA-SA", !sensor == "TNHA-SC") 

 ggplot() + 
  geom_bar(data = BASEMAXTEMP, aes(x=time,y=Max),stat = "identity",position = "identity", fill="#66A61E") +
   geom_bar(data = BASEMAXTEMP, aes(x=time, y=Min),stat = "identity", positon = "identity", fill="grey")+
  labs(title="Max/min air temperatures (Summer)") +
  guides(color = guide_legend(override.aes = list(size=0.5)))+
  theme_bw() +
  labs(y="Temperature (°C)",x="End of Week",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  ylim(0,20) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) 

```

AIR TEMPERATURE : LONG DATAFRAME
```{r}
newairtemp = gather(Weekly_Airtemp_MaxMinMed, variable, response, Max, Min, p) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

# go into this dataframe, variable = all column naames, response is the numbers that will fill that column, first column to last column, etc.

newairtemp1= cSplit(newairtemp, 'variable', sep=".", direction = "wide") # texttocols, 

dfgrndtemp$depth <- as.factor(as.character(dfgrndtemp$depth)) #taking depth from numbers and turning it into a character, then to factor (cant go from number to factor) -- when we want to sort by depth it wont take a number, only a factor **
dfgrndtemp$sensor <- as.factor(as.character(dfgrndtemp$sensor))


dfgrndtemp$station <- revalue(dfgrndtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) #renaming the stations
dfgrndtemp$sensor <- revalue(dfgrndtemp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUCEI-SA", "21398583" = "BUECI-SB", "21393042" = "BUCEI-SC", "21398584" = "BUCEI-SD", "21398579" = "BUCEI-SE", "21398578" = "BUCEI-SF.01", "21398598" = "BUCEI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398576"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21393048" = "BEO-B06", "21398591" = "BEO-B05"))

dfgrndtemp$depth <- revalue(dfgrndtemp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

#print(levels(data$column)) 
dfgrndtemp$depth = factor(dfgrndtemp$depth, c("3.5cm","10cm", "20cm", "30cm", "40cm", "50cm", "55cm", "65cm", "75cm", "85cm", "90cm"))
#can i convert air temp to this format?
```


#DAILY AVERAGE VALUE: MEAN DAILY AIR TEMPERATURE per sensor at TNHA for summer season
```{r}
bigbet %>% 
  ggplot() +
  geom_line(aes(x=day, y=value, color = sensor), size = 0.9) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(y="Temperature (°C)",x="Day",color="Station") +
  scale_x_date(date_breaks = "1 weeks") +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45, hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.caption = element_text(face ="italic")) +
  scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E"))

```


#air temp average for TNHA,BEO,SSMH
```{r fig13, fig.height = 6, fig.width = 16}
airtemp %>%
  subset(station == c("BEO","SSMH","TNHA")) %>% filter(!sensor == "21398664") %>%
  aggregate(value ~ Date + station + sensor , mean) %>%
  ggplot() +
  geom_line(aes(x=Date, y=value, color = sensor), size = 0.5) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() +
  labs(y="Temperature (°C)",color="Station") +
  scale_x_datetime(date_breaks = "1 weeks") +
  scale_y_continuous(breaks=c(-2.5,0,2.5,5,7.5,10,12.5,15,20,25),limits=c(-2.5, 25)) +
  theme(axis.text.y =element_text(face="bold") ,axis.text.x=element_text(angle = 45,
hjust = 1.0,face = "bold"), axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) + theme(plot.title = element_text(hjust = 0.5), 
         plot.caption = element_text(face ="italic")) 
   #scale_color_manual(values =c("BEO-BASE"= "#084594"))
  
  
```



#Degree day calculation
```{r}

```

Estimate of variation between sensors at 10cm between sensors at one site -- use TNHA
```{r}
#select 10cm depth from TNHA ground temp dataframe
# get max and min values from all of the sensors combined (what sensor read highest and at what val, which read lowest and at what val)
#get max and min of each sensor
#get average max and average min of each sensor

# or compatre maxs and mins on extreme days (looking at figures) between the sensors

```

MAX MIN MEAN VALUES GROUND TEMP 
```{r}
df2_tnha <- dfgrndtemp %>% filter(!station == "BUECI", !station == "BEO", !station == "SSMH", depth == "10cm")

tnhagrnd_maxmin <- df2_tnha %>% # at 10cm depths for TNHA what are each sensors maximum and minimum values
  group_by(sensor) %>%
  mutate(
    Maxval = max(value, na.rm = T),
    Minval = min(value, na.rm = T),
    mean_val = mean(value, na.rm = T)
  ) %>%
  arrange(sensor)

tnhagrnd_maxmin %>% distinct(sensor,Maxval,Minval)
```


