---
title: "X-Y Multisite Comparison Plots"
author: "Mirella Shaban"
date: '2024-01-24'
output: html_document
---

-------------IMPORTANT FILE PATHS/VALUES-------------

```{r}
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/Meteorological_Seasons_Data/" #Set to prefix of your file location

#Daily Step
datemin = "2022-06-19"
datemax = "2022-09-19"

#Depth of Interest
bestdepth = 8 #8 == 10cm
depthcm = "10 cm"
```


-------------DATA SETUP AND CLEANING-------------

```{r}

library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape) # basically like text to columns
library(naniar)
library(RColorBrewer)
library(caret)
```

Ground Temperatures for all sites data prepping
```{r}
# importing ground temp data for 6-19 through 9-20
grndtmp=read.csv(paste(filepath,"Summer_grndtmp_2023_04_06_12_21_30_UTC_1.csv",sep = ""))
 
# data cleaning for dataset
names(grndtmp) <- sub("....C..RX3000_BRW1", "", names(grndtmp))
names(grndtmp) <- sub("....C..RX3000_BRW4", "", names(grndtmp))
names(grndtmp) <- sub("....C..RX3000_BRW5", "", names(grndtmp))
names(grndtmp) <- sub("....C..RX3000_BRW6", "", names(grndtmp))
names(grndtmp) <- sub(".RXW.GP6.", "", names(grndtmp))
names(grndtmp) <- sub("Temperature.", "", names(grndtmp))
grndtmp$Date <-as.POSIXct(grndtmp$Date,format="%m/%d/%y %H:%M",tz="UTC")
#Air Temp Sensor - Remove
grndtmp = grndtmp[,-14]
#THESE ARE BLANK COLUMNS - rewrite if using a different data set
#NOTE: these have gaps in the BEGINNING part of data collection - may not always be blank
#grndtmp = grndtmp[,-c(26:36)]
#grndtmp = grndtmp[,-c(92:201)]
raw_ground = grndtmp
grndtmp$.RXW.THC.21198259.21398666.1 <- NULL
grndtmp$.RXW.THC.21198259.21398676.1 <- NULL
```


Ground Moisture for all sites data prepping
```{r}
mst=read.csv(paste(filepath,"Summer_VWC_2023_04_06_12_33_54_UTC_1.csv",sep = ""))

names(mst) <- sub("...m.3.m.3..RX3000_BRW1", "", names(mst))
names(mst) <- sub("...m.3.m.3..RX3000_BRW4", "", names(mst))
names(mst) <- sub("...m.3.m.3..RX3000_BRW5", "", names(mst))
names(mst) <- sub("...m.3.m.3..RX3000_BRW6", "", names(mst))
names(mst) <- sub("..RXW.GP6.", "", names(mst))
mst$Date <-as.POSIXct(mst$Date,format="%m/%d/%y %H:%M",tz="UTC")
```

Solar radiation for all sites data prepping
```{r}
solarrad = read.csv(paste(filepath,"Summer_solar_2023_04_06_13_16_22_UTC_1.csv",sep = ""))

names(solarrad) <- sub("olar.Radiation..RXW.LIB", "" ,names(solarrad))
names(solarrad) <- sub("olar.Radiation..S.LIB", "" ,names(solarrad))
names(solarrad) <- sub("S.", "" ,names(solarrad))
names(solarrad) <- sub("...W.m.2..RX3000_BRW1", "" ,names(solarrad))
names(solarrad) <- sub("...W.m.2..RX3000_BRW4", "" ,names(solarrad))
names(solarrad) <- sub("...W.m.2..RX3000_BRW5", "" ,names(solarrad))
names(solarrad) <- sub("...W.m.2..RX3000_BRW6", "" ,names(solarrad))
solarrad$Date <- as.POSIXct(solarrad$Date,format="%m/%d/%y %H:%M",tz="UTC")
```


```{r}
airtemp=read.csv(paste(filepath,"Summer_airtemp_2023_04_06_11_59_55_UTC_1.csv",sep = ""))

names(airtemp) <- sub("Temperature..RXW.THC.", "" ,names(airtemp))
names(airtemp) <- sub("Temperature..S.THC.", "" ,names(airtemp))
names(airtemp) <- sub("S.", "" ,names(airtemp))
names(airtemp) <- sub("....C..RX3000_BRW1", "" ,names(airtemp))
names(airtemp) <- sub("....C..RX3000_BRW4", "" ,names(airtemp))
names(airtemp) <- sub("....C..RX3000_BRW5", "" ,names(airtemp))
names(airtemp) <- sub("....C..RX3000_BRW6", "" ,names(airtemp))
airtemp$Date <- as.POSIXct(airtemp$Date,format="%m/%d/%y %H:%M",tz="UTC")
```


-------Convert to Long DataFrame-------

#Solar Radiation
```{r fig9, fig.heigt = 14, fig.width = 18}
raw_solar = solarrad

solarrad = gather(solarrad, variable, response, '21198259.21176526.1':'21401800.21390411.1') %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

solarrad= cSplit(solarrad, 'variable', sep=".", direction = "wide") # texttocols, 
solarrad$variable_3 <- NULL
colnames(solarrad)[3:5] <- c("value", "station", "sensor")
solarrad$value <- as.numeric(as.character(solarrad$value)) 
solarrad$sensor <- as.factor(as.character(solarrad$sensor))
solarrad$station <- as.factor(solarrad$station)

solarrad$station <- revalue(solarrad$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO"))





solarrad$sensor <- revalue(solarrad$sensor, c("21390411" = "BUECI-BASE", "21398618" = "BUECI-SA", "21398624" = "BUECI-SB", "21362313" = "BUECI-SC", "21362316" = "BUECI-SD", "21362320" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", 
                                              "21390413" ="SSMH-BASE", "21398622" = "SSMH-SA", "21362319" = "SSMH-SB", "21166008" = "SSMH-SC", "21393049" = "SSMH-SD" ,
                                              "21176526" = "TNHA-BASE", "21398620" = "TNHA-SA", "21398616"="TNHA-SB","21362315" = "TNHA-SD", "21362317" = "TNHA-SC", 
                                              "21390415" = "BEO-BASE"))

solarrad <- solarrad %>% filter(!value == -888.88)
```

```{r fig9, fig.heigt = 14, fig.width = 18}
grndtmp = raw_ground

grndtmp = gather(grndtmp, variable, response, 3:ncol(grndtmp)) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

grndtmp= cSplit(grndtmp, 'variable', sep=".", direction = "wide") # texttocols, 
colnames(grndtmp)[3:6] <- c("value", "station", "sensor","depth")
grndtmp$value <- as.numeric(as.character(grndtmp$value)) 
grndtmp$sensor <- as.factor(as.character(grndtmp$sensor))
grndtmp$station <- as.factor(grndtmp$station)
grndtmp$depth <- as.factor(grndtmp$depth)
grndtmp$variable_4 <- NULL
grndtmp$variable_5 <- NULL
grndtmp$variable_6 <- NULL
grndtmp$station <- revalue(grndtmp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO"))
grndtmp$depth <- revalue(grndtmp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11"
                                          = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" =
                                            "75cm", "16" = "85cm", "17" = "90cm"))


#Find Sensor Values

grndtmp$sensor <- revalue(grndtmp$sensor, c("21398585" = "BUECI-BASE", "21398590" = "BUECI-SA", "21398583" = "BUECI-SB", "21393042" = "BUECI-SC", "21398584" = "BUECI-SD", "21398579" = "BUECI-SE", "21398578" = "BUECI-SF.01", "21398598" = "BUECI-SF.02", "21398588" ="SSMH-BASE","21393049" = "SSMH-SD", "21398599" = "SSMH-SA", "21393044" = "SSMH-SB", "21393049" = "SSMH-SD" , "21206939" = "TNHA-BASE", "21398593" = "TNHA-SA", "21398587"="TNHA-SB","21398601" = "TNHA-SD", "21393047" = "TNHA-SC", "21398591" = "BEO-BASE"))

#BEO-BASE here is actually BEO B05 for the XY plots to work
grndtmp <- grndtmp %>% filter(!value == -888.88, !depth == "THC")
```



```{r fig9, fig.heigt = 14, fig.width = 18}
raw_air = airtemp

airtemp = gather(airtemp, variable, response, 3:ncol(airtemp)) %>%
  mutate(variable = fct_inorder(variable))
airtemp= cSplit(airtemp, 'variable', sep=".", direction = "wide") 
airtemp$variable_3 <- NULL
colnames(airtemp)[3:5] <- c("value", "station", "sensor")
airtemp$value <- as.numeric(as.character(airtemp$value)) 
airtemp$sensor <- as.factor(as.character(airtemp$sensor))
airtemp$station <- as.factor(airtemp$station)
airtemp$station <- revalue(airtemp$station, c("21198259" = "TNHA", "21401800" = "BUECI", "21401801" = "SSMH", "21401803" = "BEO")) 
airtemp$sensor <- revalue(airtemp$sensor, c("21390851" = "BUECI-BASE", "21398671" = "BUECI-SA", "21362254" = "BUECI-SB", "21398668" = "BUECI-SC", "21398660" = "BUECI-SD", "21187245" = "BUECI-SE",
                                            "21390849" ="SSMH-BASE","21393049" = "SSMH-SD", "21398670" = "SSMH-SA", "21398665" = "SSMH-SB",
                                            "21218018" = "TNHA-BASE", "21380919" = "TNHA-SA" , "21398676"="TNHA-SB","21398666" = "TNHA-SD", "21398674" = "TNHA-SC", "21397541" = "BEO-BASE"))
#TNHA-A and TNHA-B are GUESSES - need metadata on ground

airtemp <- airtemp %>% filter(!value == -888.88)

```
-------Convert to Daily Values-------

```{r}
avgdailysolar = data.frame(Date=NA, value=NA, station=NA, sensor=NA)[numeric(0), ]
sensorList = c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD", "BEO-BASE", "SSMH-BASE", "SSMH-SA", "SSMH-SB")
#adjusted to include other stations -- change back to just TNHA***

for(sensorName in sensorList){
  daily = solarrad %>% filter(sensor==sensorName) %>% filter(as.Date(Date, "%Y-%m-%d") <= datemax) %>%                 filter(as.Date(Date, "%Y-%m-%d") >= datemin)

  daily$Date = floor_date(daily$Date, unit = "days", week_start = getOption("lubridate.week.start",      7))
           
  daily = aggregate(value ~ Date, daily, mean, na.rm=T)
  sensor = rep(sensorName, nrow(daily))
  station = rep(sub(c("-BASE|-SA|-SB|-SC|-SD"), "" ,sensorName), nrow(daily))
  daily = data.frame(daily, station, sensor)
  avgdailysolar = rbind(avgdailysolar, daily)
}

colnames(avgdailysolar)[2] = "solar"

```
#adjust sensorList for BEO-B05, BEO-B06 if comparing ground temp to ground temp for XY. Set equal to BEO-BASE if comparing air temp to ground temp xY
```{r}
grndtmp_10cm <- grndtmp %>% filter(depth == "10cm")
avgdailyground = data.frame(Date=NA, value=NA, station=NA, sensor=NA)[numeric(0), ]
sensorList = c("TNHA-BASE", "TNHA-SA", "TNHA-SB","TNHA-SC","TNHA-SD", "BEO-BASE", "SSMH-BASE", "SSMH-SA", "SSMH-SB")


for(sensorName in sensorList){
    
    daily = grndtmp_10cm %>% filter(sensor==sensorName) %>% filter(as.Date(Date, "%Y-%m-%d") <= datemax) %>%
      filter(as.Date(Date, "%Y-%m-%d") >= datemin)
if(nrow(daily) > 0){
  daily$Date = floor_date(daily$Date, unit = "days", week_start = getOption("lubridate.week.start",      7))

  daily = aggregate(value ~ Date, daily, mean, na.rm=T)
  sensor = rep(sensorName, nrow(daily))
  station = rep(sub(c("-BASE|-SA|-SB|-SC|-SD|-B05|-B06"), "" ,sensorName), nrow(daily))
  daily = data.frame(daily, station, sensor)
  avgdailyground = rbind(avgdailyground, daily)
}

  
}

colnames(avgdailyground)[2] = "grndtmp"

```

```{r}
avgdailyair = data.frame(Date=NA, value=NA, station=NA, sensor=NA)[numeric(0), ]
sensorList = c("TNHA-BASE", "TNHA-SA", "TNHA-SB", "TNHA-SC", "TNHA-SD", "BEO-BASE", "SSMH-BASE", "SSMH-SA", "SSMH-SB")


for(sensorName in sensorList){
  daily = airtemp %>% filter(sensor==sensorName) %>% filter(as.Date(Date, "%Y-%m-%d") <= datemax) %>%                 filter(as.Date(Date, "%Y-%m-%d") >= datemin)

  daily$Date = floor_date(daily$Date, unit = "days", week_start = getOption("lubridate.week.start",      7))

  daily = aggregate(value ~ Date, daily, mean, na.rm=T)
  sensor = rep(sensorName, nrow(daily))
  station = rep(sub(c("-BASE|-SA|-SB|-SC|-SD"), "" ,sensorName), nrow(daily))
  daily = data.frame(daily, station, sensor)
  avgdailyair= rbind(avgdailyair, daily)
}

colnames(avgdailyair)[2] = "airtemp"
```
----- AMPLITUDE -----

```{r}
grndamp <- grndtmp %>% filter(grndtmp$depth == "10cm", !value == 0.0) 
grndamp <- grndamp %>%
  group_by(Date) %>%
  mutate(
    Max_grnd = max(value, na.rm = T),
    Min_grnd = min(value, na.rm = T),
    mean_grnd = mean(value, na.rm = T),
    amplitude = (Max_grnd-Min_grnd)
  ) %>%
   arrange(sensor)

grndamp$Date <- as.Date(grndamp$Date)

grndamp_averages <- grndamp %>% group_by(Date, sensor, Date = as.Date(Date)) %>% summarise(across(c(Max_grnd, Min_grnd, mean_grnd, amplitude), mean))

#grndamp2 <- grndamp %>% tidyr::pivot_wider(id_cols = Date, names_from = sensor, values_from = amplitude, values_fn = function(x) mean(x, na.rm = T))

#grndamp2 does the same as grndamp1 but as a wide dataframe !!!!


# consider using xts for apply.weekly, apply.mothly etc. actions to aggregate data over time
```

-------Merge Data Frames-------

```{r}
averages = avgdailysolar %>% full_join(avgdailyground, by=c("Date","sensor","station"))
solar = averages[,2]
averages = averages[,-2]
averages = data.frame(averages, solar)

averages = averages %>% full_join(avgdailyair, by=c("Date","sensor","station"))
averages = averages %>% full_join(grndamp_averages, by=c("Date", "sensor"))

```

-------Graphs for One Week-------

```{r}
model = train(solar~airtemp, averages, method="lm", na.action = na.exclude)
summary(model)

pred <- predict(model, averages)

#averages = data.frame(averages, pred)

#one week time span
```

```{r}
model1 = train(grndtmp~airtemp, averages, method="lm", na.action = na.exclude)
summary(model1)
pred1 <- predict(model1,averages)
head(pred)
#averages = data.frame(averages,pred1)
```

# Plotting XY figures of stations grnd tmp vs air tmp against eachother for later gap filling comparisons

```{r}
XY_grnd_air= averages %>% filter(station == "TNHA" | station == "BEO", !sensor == "TNHA-SD", !sensor == "TNHA-SB", !sensor == "TNHA-SA", !sensor == "TNHA-SC",!sensor == "TNHA-BASE") 

  ggplot(data = XY_grnd_air, aes(x=airtemp, y = amplitude, color=sensor)) +
  geom_point()+
    guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() + 
    labs(title= "BEO-B05 10cm @ depth AMP, 6/19/22-9/19/22", x = "Air Temperature (°C)", y="Ground Temperature (°C) @ 10cm",color="Station") + 
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 15),plot.title = element_text(hjust = 0.5, size = 15), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 15),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1,'cm')) +
    scale_x_continuous(breaks=c(0,2,4,6,8,10,12)) +
  scale_y_continuous(breaks=c(0,2,4,6,8,10,12), limits = c(0,12)) +
    geom_smooth(method=lm, se=TRUE) + #add a line of best fit 
    scale_color_manual(values =c("BEO-BASE"= "#084594"))
  
  #"TNHA-BASE" = "#1B9E77","TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E",
  
  #add date min and max
```


linear regression of XY plot above
```{r}
#adjust the XY_grnd_air filter in the plotting chunk above to adjust which sensors you're comparing
fit1 <- lm(amplitude ~ airtemp, data = XY_grnd_air)
summary(fit1)
corfit <- cor(XY_grnd_air$airtemp,XY_grnd_air$amplitude, use = 'complete.obs')
print("Correlation coefficint")
print(corfit)
#summary(fit1)$r.squared
```

Comparing BEO (control) vs TNHA 
-- note that you must set BEO-B06 or BEO-B05 (one at a time) = to "BEO-BASE" in the ground temperature long datarfame conversion chunk above. Name of air temp and ground temp station/sensor must MATCH for XY comparison
```{r}
XY_grnd_air= averages %>% filter(station == "TNHA" | station == "BEO", !sensor == "TNHA-BASE", !sensor == "TNHA-SD", !sensor == "TNHA-SB") 

  ggplot(data = XY_grnd_air, aes(x=airtemp, y = grndtmp, color=sensor)) +
  geom_point()+
    guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() + 
    labs(title= "TNHA N/S aspects @ 10cm depth", x = "Air Temperature (°C)", y="Ground Temperature (°C)",color="Station") + 
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 15),plot.title = element_text(hjust = 0.5, size = 15), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 15),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1,'cm')) +
    geom_smooth(method=lm, se=TRUE) + #add a line of best fit 
    scale_color_manual(values =c("TNHA-SA" = "#D95F02","TNHA-SC"="#E7298A" ,"BEO-BASE"= "#084594"))
  
#color references:  "TNHA-BASE" = "#1B9E77","TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E",

```

All TNHA sites vs BEO-B06 comparisons

```{r}
XY_grnd_air= averages %>% filter(station == "TNHA" | station == "BEO") 

  ggplot(data = XY_grnd_air, aes(x=airtemp, y = grndtmp, color=sensor)) +
  geom_point()+
    guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() + 
    labs(title= "TNHA vs. BEO @ 10cm depth, summer 2022", x = "Air temperature (°C)", y="Ground temperature (°C)",color="Station") + 
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 15),plot.title = element_text(hjust = 0.5, size = 15), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 15),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1,'cm')) +
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12), limits = c(0,12)) +
    scale_y_continuous(breaks = c(0,2,4,6,8,10,12), limits = c(0,12))+
    geom_smooth(method=lm, se=TRUE) + #add a line of best fit +
    geom_abline(intercept = 0, slope = 1, colour = "red") +
    scale_color_manual(name = "Station",values =c("TNHA-BASE" = "#1B9E77", "TNHA-SA" = "#D95F02","TNHA-SC" ="#E7298A", "TNHA-SD"="#66A61E","BEO-BASE"= "#084594", labels = c( "TNHA-BASE (South Facing)", "TNHA-A (South Facing)", "TNHA-C (North Facing)", "TNHA-D (East Facing)","BEO (Control)")))
  
  #
      scale_color_manual(values =c("TNHA-BASE" = "#1B9E77","TNHA-SA" = "#D95F02", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E","BEO-BASE"= "#084594"))
  
#color references:  "TNHA-BASE" = "#1B9E77","TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E",

```

All SSMH sites (w/airtemp sensors) vs BEO-B06 comparisons

```{r}
XY_grnd_air= averages %>% filter(station == "SSMH" | station == "BEO") 

  ggplot(data = XY_grnd_air, aes(x=airtemp, y = grndtmp, color=sensor)) +
  geom_point()+
    guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() + 
    labs(title= "SSMH vs. BEO @ 10cm depth, summer 2022", x = "Air temperature (°C)", y="Ground temperature (°C)",color="Station") + 
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 15),plot.title = element_text(hjust = 0.5, size = 15), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 15),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1,'cm')) +
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12), limits = c(0,12)) +
    scale_y_continuous(breaks = c(0,2,4,6,8,10,12), limits = c(0,12))+
    geom_smooth(method=lm, se=TRUE) + #add a line of best fit +
    geom_abline(intercept = 0, slope = 1, colour = "red") +
    scale_color_manual(values =c("BEO-BASE"= "#084594","SSMH-BASE"="#807DBA", "SSMH-SA"="#A6761D", "SSMH-SB"= "#4292C6"))
  
#color references:  "TNHA-BASE" = "#1B9E77","TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E",

```



--- AMPLTIUDE GRAPHS ---
```{r}

XY_grnd_avgs = averages %>% filter(station == "TNHA"| station == "BEO", !sensor == "TNHA-SA", !sensor == "TNHA-BASE", !sensor == "TNHA-SB") 
  ggplot(data = XY_grnd_avgs, aes(x=airtemp, y = amplitude, color=sensor)) +
  geom_point()+
  theme_bw() + 
    labs(title= "TNHA-SC vs. TNHA-SD (°C) @ 10cm, 6/19/22-9/19/22", x = "Air Temperature", y="Ground Temperature Amplitude (°C) @ 10cm",color="Station") + 
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 15),plot.title = element_text(hjust = 0.5, size = 15), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 15),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1,'cm')) +
    geom_smooth(method=lm, se=TRUE)

```
SOLAR RADIATION VS GROUND TEMP

```{r}
XY_grnd_sol= averages %>% filter(station == "TNHA" | station == "BEO", !sensor == "TNHA-SD", !sensor == "TNHA-SC", !sensor == "TNHA-BASE", !sensor == "TNHA-SB",!sensor == "BEO-BASE") 

  ggplot(data = XY_grnd_sol, aes(x=solar, y = amplitude, color=sensor)) +
    geom_point()+
    guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw() + 
    labs(title= "TNHA-SA 10cm @ depth AMP, 6/19/22-9/19/22", y = "Ground Temperature (°C)", x="W /"~m^2~"",color="Station") + 
  theme(axis.text.y = element_text(size = 15)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1.0, size = 15), 
        axis.line = element_line(colour = "black", size =0.5, 
                                 linetype = "solid")) +
  theme(strip.text.y = element_text(size = 15),plot.title = element_text(hjust = 0.5, size = 15), 
        plot.caption = element_text(face ="italic"),
        strip.text = element_text(size = 15),
        axis.title = element_text(size = 15)) +
  theme(axis.title.y = element_text(margin = margin(r=15))) +
  theme(axis.title.x = element_text(margin = margin(t=15))) +
  theme(legend.title = element_text(size = 15),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1,'cm')) +
   scale_x_continuous(breaks=c(0,100,200,300), limits = c(0,300)) +
  scale_y_continuous(breaks=c(0,2,4,6,8,10,12), limits = c(0,12)) +
    geom_smooth(method=lm, se=TRUE) + #add a line of best fit 
    scale_color_manual(values =c("TNHA-SA"= "#D95F02"))
  
  #"TNHA-BASE" = "#1B9E77","TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E",
  
  #add date min and max
```


linear regression of XY plot above
```{r}
#adjust the XY_grnd_air filter in the plotting chunk above to adjust which sensors you're comparing
fit1 <- lm(solar ~ amplitude, data = XY_grnd_sol)
summary(fit1)
corfit <- cor(XY_grnd_sol$amplitude,XY_grnd_sol$solar, use = 'complete.obs')
print("Correlation coefficint")
print(corfit)
#summary(fit1)$r.squared
```
