# Average Gap Filled Data
Get statistics and prep for data visualization.

```{r}
filepath <- "/Users/emvanmetre/Desktop/Arctic/Gap_Filled_Data/Transposed/"
```

```{r}
library(dplyr)
```



## BEO

Load ground temperature, air temperature, volumetric water content, solar radiation, wind direction, and wind speed datasets. (From Prep_Data.Rmd)
```{r}
gt <- read.csv(paste0(filepath, "groundtemp_gap_filled_renamed_BEO.csv"))
at <- read.csv(paste0(filepath, "airtemp_gap_filled_renamed_BEO.csv"))
vwc <- read.csv(paste0(filepath, "vwc_gap_filled_renamed_BEO.csv"))
sol <- read.csv(paste0(filepath, "solar_gap_filled_renamed_BEO.csv"))
ws <- read.csv(paste0(filepath, "windspeed_gap_filled_renamed_BEO.csv"))
wd <- read.csv(paste0(filepath, "winddirection_gap_filled_renamed_BEO.csv"))
```

```{r}
ground_hourly <- gt %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor, depth) %>% summarize(mean = mean(value, na.rm = T))
ground_hourly$hour[ground_hourly$hour == ""] = "00"
ground_hourly$Date <- as.POSIXct(paste0(ground_hourly$date, " ", ground_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(ground_hourly)[8] = "groundtemp"
ground_hourly <- ground_hourly[,c(3:9)]

vwc_hourly <- vwc %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor, depth) %>% summarize(mean = mean(value, na.rm = T))
vwc_hourly$hour[vwc_hourly$hour == ""] = "00"
vwc_hourly$Date <- as.POSIXct(paste0(vwc_hourly$date, " ", vwc_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(vwc_hourly)[8] = "vwc"
vwc_hourly <- vwc_hourly[,c(3:9)]

air_hourly <- at %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
air_hourly$hour[air_hourly$hour == ""] = "00"
air_hourly$Date <- as.POSIXct(paste0(air_hourly$date, " ", air_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(air_hourly)[7] = "airtemp"
air_hourly <- air_hourly[,c(3:8)]

solar_hourly <- sol %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
solar_hourly$hour[solar_hourly$hour == ""] = "00"
solar_hourly$Date <- as.POSIXct(paste0(solar_hourly$date, " ", solar_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(solar_hourly)[7] = "solar"
solar_hourly <- solar_hourly[,c(3:8)]

wspeed_hourly <- ws %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
wspeed_hourly$hour[wspeed_hourly$hour == ""] = "00"
wspeed_hourly$Date <- as.POSIXct(paste0(wspeed_hourly$date, " ", wspeed_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(wspeed_hourly)[7] = "windspeed"
wspeed_hourly <- wspeed_hourly[,c(3:8)]

wdir_hourly <- wd %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
wdir_hourly$hour[wdir_hourly$hour == ""] = "00"
wdir_hourly$Date <- as.POSIXct(paste0(wdir_hourly$date, " ", wdir_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(wdir_hourly)[7] = "winddir"
wdir_hourly <- wdir_hourly[,c(3:8)]
```
**subject to change**
Ground Depth
7 = "3.5cm" [1]
8 = "10cm" [1]
9 = "20cm" [2]
10 = "30cm" [3] **
11 = "40cm" [3]
12 = "50cm"[4]
13 = "55cm" [4]
14 = "65cm" [5]
15 = "75cm" [6] **
16 = "85cm" [6]
17 = "90cm" [6]

VWC Depths
1 = "0:15cm"
2 = "15:30cm"
3 = "30:45cm"
4 = "45:60cm"
5 = "60:75cm"
6 = "75:90cm"

Join variables
```{r}
ground_hourly <- ground_hourly %>% mutate(vwcdepth = case_when(depth == 7 ~ 1,
          depth == 8 ~ 1,
          depth == 9 ~ 2,
          depth == 10 ~ 3,
          depth == 11 ~ 3,
          depth == 12 ~ 4,
          depth == 13 ~ 4,
          depth == 14 ~ 5,
          depth == 15 ~ 6,
          depth == 16 ~ 6,
          depth == 17 ~ 6)) %>% filter(depth != 1)
colnames(vwc_hourly)[5] = "vwcdepth"
beo <- full_join(ground_hourly, vwc_hourly, by=c("Date", "fullname", "site", "station", "vwcdepth")) 
colnames(beo)[5] = "grounddepth"
beo <- full_join(beo, air_hourly, by=c("Date", "fullname", "site", "station"))
beo <- full_join(beo, solar_hourly, by=c("Date", "fullname", "site", "station"))
beo <- full_join(beo, wspeed_hourly, by=c("Date", "fullname", "site", "station"))
# beo <- full_join(beo, wdir_hourly, by=c("Date", "fullname")) - no data
```

```{r}
write.csv(beo, paste0(filepath, "BEO", "_hourly_gap_filled", ".csv"))
```


### Get Sensors

```{r}
ground_sensors <- ground_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

air_sensors <- air_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

vwc_sensors <- vwc_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

solar_sensors <- solar_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

winddir_sensors <- wdir_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

windspeed_sensors <- wspeed_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

all_ground_sensors <- ground_sensors
all_air_sensors <- air_sensors
all_vwc_sensors <- vwc_sensors
all_solar_sensors <- solar_sensors
all_wdir_sensors <- winddir_sensors
all_wspeed_sensors <- windspeed_sensors
```


## TNHA

Load ground temperature, air temperature, volumetric water content, solar radiation, wind direction, and wind speed datasets. (From Prep_Data.Rmd)
```{r}
gt <- read.csv(paste0(filepath, "groundtemp_gap_filled_renamed_TNHA.csv"))
at <- read.csv(paste0(filepath, "airtemp_gap_filled_renamed_TNHA.csv"))
vwc <- read.csv(paste0(filepath, "vwc_gap_filled_renamed_TNHA.csv"))
sol <- read.csv(paste0(filepath, "solar_gap_filled_renamed_TNHA.csv"))
ws <- read.csv(paste0(filepath, "windspeed_gap_filled_renamed_TNHA.csv"))
wd <- read.csv(paste0(filepath, "winddirection_gap_filled_renamed_TNHA.csv"))
```

```{r}
ground_hourly <- gt %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor, depth) %>% summarize(mean = mean(value, na.rm = T))
ground_hourly$hour[ground_hourly$hour == ""] = "00"
ground_hourly$Date <- as.POSIXct(paste0(ground_hourly$date, " ", ground_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(ground_hourly)[8] = "groundtemp"
ground_hourly <- ground_hourly[,c(3:9)]

vwc_hourly <- vwc %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor, depth) %>% summarize(mean = mean(value, na.rm = T))
vwc_hourly$hour[vwc_hourly$hour == ""] = "00"
vwc_hourly$Date <- as.POSIXct(paste0(vwc_hourly$date, " ", vwc_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(vwc_hourly)[8] = "vwc"
vwc_hourly <- vwc_hourly[,c(3:9)]

air_hourly <- at %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
air_hourly$hour[air_hourly$hour == ""] = "00"
air_hourly$Date <- as.POSIXct(paste0(air_hourly$date, " ", air_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(air_hourly)[7] = "airtemp"
air_hourly <- air_hourly[,c(3:8)]

solar_hourly <- sol %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
solar_hourly$hour[solar_hourly$hour == ""] = "00"
solar_hourly$Date <- as.POSIXct(paste0(solar_hourly$date, " ", solar_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(solar_hourly)[7] = "solar"
solar_hourly <- solar_hourly[,c(3:8)]

wspeed_hourly <- ws %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
wspeed_hourly$hour[wspeed_hourly$hour == ""] = "00"
wspeed_hourly$Date <- as.POSIXct(paste0(wspeed_hourly$date, " ", wspeed_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(wspeed_hourly)[7] = "windspeed"
wspeed_hourly <- wspeed_hourly[,c(3:8)]

wdir_hourly <- wd %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
wdir_hourly$hour[wdir_hourly$hour == ""] = "00"
wdir_hourly$Date <- as.POSIXct(paste0(wdir_hourly$date, " ", wdir_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(wdir_hourly)[7] = "winddir"
wdir_hourly <- wdir_hourly[,c(3:8)]
```

**subject to change**
Ground Depth
7 = "3.5cm" [1]
8 = "10cm" [1]
9 = "20cm" [2]
10 = "30cm" [3] **
11 = "40cm" [3]
12 = "50cm"[4]
13 = "55cm" [4]
14 = "65cm" [5]
15 = "75cm" [6] **
16 = "85cm" [6]
17 = "90cm" [6]

Join variables
```{r}
ground_hourly <- ground_hourly %>% mutate(vwcdepth = case_when(depth == 7 ~ 1,
          depth == 8 ~ 1,
          depth == 9 ~ 2,
          depth == 10 ~ 3,
          depth == 11 ~ 3,
          depth == 12 ~ 4,
          depth == 13 ~ 4,
          depth == 14 ~ 5,
          depth == 15 ~ 6,
          depth == 16 ~ 6,
          depth == 17 ~ 6)) %>% filter(depth != 1)
colnames(vwc_hourly)[5] = "vwcdepth"
tnha <- full_join(ground_hourly, vwc_hourly, by=c("Date", "fullname", "site", "station", "vwcdepth")) 
colnames(tnha)[5] = "grounddepth"
tnha <- full_join(tnha, air_hourly, by=c("Date", "fullname", "site", "station"))
tnha <- full_join(tnha, solar_hourly, by=c("Date", "fullname", "site", "station"))
tnha <- full_join(tnha, wspeed_hourly, by=c("Date", "fullname", "site", "station"))
tnha <- full_join(tnha, wdir_hourly, by=c("Date", "fullname", "site", "station"))
```

```{r}
write.csv(tnha, paste0(filepath, "TNHA", "_hourly_gap_filled", ".csv"))
```

### Get Sensors

```{r}
ground_sensors <- ground_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

air_sensors <- air_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

vwc_sensors <- vwc_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

solar_sensors <- solar_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

winddir_sensors <- wdir_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

windspeed_sensors <- wspeed_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

all_ground_sensors <- rbind(all_ground_sensors, ground_sensors)
all_air_sensors <- rbind(all_air_sensors, air_sensors)
all_vwc_sensors <- rbind(all_vwc_sensors, vwc_sensors)
all_solar_sensors <- rbind(all_solar_sensors, solar_sensors)
all_wdir_sensors <- winddir_sensors
all_wspeed_sensors <- rbind(all_wspeed_sensors, windspeed_sensors)
```


## SSMH

Load ground temperature, air temperature, volumetric water content, solar radiation, wind direction, and wind speed datasets. (From Prep_Data.Rmd)
```{r}
gt <- read.csv(paste0(filepath, "groundtemp_gap_filled_renamed_SSMH.csv"))
at <- read.csv(paste0(filepath, "airtemp_gap_filled_renamed_SSMH.csv"))
vwc <- read.csv(paste0(filepath, "vwc_gap_filled_renamed_SSMH.csv"))
sol <- read.csv(paste0(filepath, "solar_gap_filled_renamed_SSMH.csv"))
ws <- read.csv(paste0(filepath, "windspeed_gap_filled_renamed_SSMH.csv"))
wd <- read.csv(paste0(filepath, "winddirection_gap_filled_renamed_SSMH.csv"))
```

```{r}
ground_hourly <- gt %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor, depth) %>% summarize(mean = mean(value, na.rm = T))
ground_hourly$hour[ground_hourly$hour == ""] = "00"
ground_hourly$Date <- as.POSIXct(paste0(ground_hourly$date, " ", ground_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(ground_hourly)[8] = "groundtemp"
ground_hourly <- ground_hourly[,c(3:9)]

vwc_hourly <- vwc %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor, depth) %>% summarize(mean = mean(value, na.rm = T))
vwc_hourly$hour[vwc_hourly$hour == ""] = "00"
vwc_hourly$Date <- as.POSIXct(paste0(vwc_hourly$date, " ", vwc_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(vwc_hourly)[8] = "vwc"
vwc_hourly <- vwc_hourly[,c(3:9)]

air_hourly <- at %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
air_hourly$hour[air_hourly$hour == ""] = "00"
air_hourly$Date <- as.POSIXct(paste0(air_hourly$date, " ", air_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(air_hourly)[7] = "airtemp"
air_hourly <- air_hourly[,c(3:8)]

solar_hourly <- sol %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
solar_hourly$hour[solar_hourly$hour == ""] = "00"
solar_hourly$Date <- as.POSIXct(paste0(solar_hourly$date, " ", solar_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(solar_hourly)[7] = "solar"
solar_hourly <- solar_hourly[,c(3:8)]

wspeed_hourly <- ws %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
wspeed_hourly$hour[wspeed_hourly$hour == ""] = "00"
wspeed_hourly$Date <- as.POSIXct(paste0(wspeed_hourly$date, " ", wspeed_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(wspeed_hourly)[7] = "windspeed"
wspeed_hourly <- wspeed_hourly[,c(3:8)]

wdir_hourly <- wd %>% mutate(hour = substr(Date, 12, 13), date = substr(Date, 1, 10)) %>% group_by(date, hour, fullname, site, station, sensor) %>% summarize(mean = mean(value, na.rm = T))
wdir_hourly$hour[wdir_hourly$hour == ""] = "00"
wdir_hourly$Date <- as.POSIXct(paste0(wdir_hourly$date, " ", wdir_hourly$hour, ":00:00"), format="%Y-%m-%d %H:%M:%S",tz="UTC")
colnames(wdir_hourly)[7] = "winddir"
wdir_hourly <- wdir_hourly[,c(3:8)]
```

**subject to change**
Ground Depth
7 = "3.5cm" [1]
8 = "10cm" [1]
9 = "20cm" [2]
10 = "30cm" [3] **
11 = "40cm" [3]
12 = "50cm"[4]
13 = "55cm" [4]
14 = "65cm" [5]
15 = "75cm" [6] **
16 = "85cm" [6]
17 = "90cm" [6]

Join variables
```{r}
ground_hourly <- ground_hourly %>% mutate(vwcdepth = case_when(depth == 7 ~ 1,
          depth == 8 ~ 1,
          depth == 9 ~ 2,
          depth == 10 ~ 3,
          depth == 11 ~ 3,
          depth == 12 ~ 4,
          depth == 13 ~ 4,
          depth == 14 ~ 5,
          depth == 15 ~ 6,
          depth == 16 ~ 6,
          depth == 17 ~ 6)) %>% filter(depth != 1)
colnames(vwc_hourly)[5] = "vwcdepth"
ssmh <- full_join(ground_hourly, vwc_hourly, by=c("Date", "fullname", "site", "station", "vwcdepth")) 
colnames(ssmh)[5] = "grounddepth"
ssmh <- full_join(ssmh, air_hourly, by=c("Date", "fullname", "site", "station"))
ssmh <- full_join(ssmh, solar_hourly, by=c("Date", "fullname", "site", "station"))
ssmh <- full_join(ssmh, wspeed_hourly, by=c("Date", "fullname", "site", "station"))
ssmh <- full_join(ssmh, wdir_hourly, by=c("Date", "fullname", "site", "station"))
```

```{r}
write.csv(ssmh, paste0(filepath, "SSMH", "_hourly_gap_filled", ".csv"))
```


### Get Sensors

```{r}
ground_sensors <- ground_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

air_sensors <- air_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

vwc_sensors <- vwc_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

solar_sensors <- solar_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

winddir_sensors <- wdir_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

windspeed_sensors <- wspeed_hourly %>% select(sensor, fullname, Date) %>% group_by(sensor, fullname) %>% summarize(mindate = substr(min(Date), 1, 10), maxdate = substr(max(Date), 1, 10))

all_ground_sensors <- rbind(all_ground_sensors, ground_sensors)
all_air_sensors <- rbind(all_air_sensors, air_sensors)
all_vwc_sensors <- rbind(all_vwc_sensors, vwc_sensors)
all_solar_sensors <- rbind(all_solar_sensors, solar_sensors)
all_wdir_sensors <- rbind(all_wdir_sensors, winddir_sensors)
all_wspeed_sensors <- rbind(all_wspeed_sensors, windspeed_sensors)
```
## Export Sensors

```{r}
write.csv(all_ground_sensors, paste0(filepath, "all_ground_sensors_10-9-24.csv"))
write.csv(all_air_sensors, paste0(filepath, "all_air_sensors_10-9-24.csv"))
write.csv(all_solar_sensors, paste0(filepath, "all_solar_sensors_10-9-24.csv"))
write.csv(all_vwc_sensors, paste0(filepath, "all_vwc_sensors_10-9-24.csv"))
write.csv(all_wdir_sensors, paste0(filepath, "all_wdir_sensors_10-9-24.csv"))
write.csv(all_wspeed_sensors, paste0(filepath, "all_wspeed_sensors_10-9-24.csv"))
```


## Combine All Sites

```{r}
all_sites <- rbind(beo, tnha)
all_sites <- rbind(all_sites, ssmh)
```

```{r}
write.csv(all_sites, paste0(filepath, "all_sites", "_hourly_gap_filled", ".csv"))
```

