# Transpose Data
## NOTE: THIS FILE IS NO LONGER NEEDED BUT THE CODE IS PRESERVED FOR REFERENCE OR POTENTIAL USE IN THE FUTURE

### Load Libraries
```{r}
library(dplyr)
library(tidyr)
library(splitstackshape)
```




## Files

### Moisture
 [1] "Moist_data_CLEAN_BEO_from_2022_06_13_to_2024_08_31_no_gps.csv"    
 [2] "Moist_data_CLEAN_SSMH_from_2022_06_08_to_2024_08_28_no_gps.csv"   
 [3] "Moist_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv"  
 
### Other
 [4] "Other_data_CLEAN_SSMH_from_2022_06_08_to_2024_09_20_no_gps.csv"   
 [5] "Other_sensors_data_CLEAN_BEO_from_2022_06_13_to_2024_08_31_no_gps.csv" 
 [6] "Other_sensors_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv"
 
### Solar
 [7] "Solar_data_CLEAN_BEO_from_2022_06_13_to_2024_08_31_no_gps.csv"    
 [8] "Solar_data_CLEAN_SSMH_from_2022_06_08_to_2024_09_20_no_gps.csv"   
 [9] "Solar_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv"  

### Temperature
[10] "Temp_data_CLEAN_BEO_from_2022_06_13_to_2024_08_31_no_gps.csv"     
[11] "Temp_data_CLEAN_SSMH_from_2022_06_08_to_2024_08_28_no_gps.csv"    
[12] "Temp_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv" 

### Wind
[13] "Wind_data_CLEAN_BEO_from_2022_06_13_to_2024_08_31_no_gps.csv"     
[14] "Wind_data_CLEAN_SSMH_from_2022_06_08_to_2024_09_20_no_gps.csv"    
[15] "Wind_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv" 

### Set Filepath

List all the files in the folder
```{r}
filepath <- "/Users/emvanmetre/Desktop/Arctic/Gap_Filled_Data/"
exportpath <- "/Users/emvanmetre/Desktop/Arctic/Gap_Filled_Data/Transposed/"
filenames <- list.files(filepath, pattern="*.csv", full.names=F)
filenames
```

## Function

Load and Transpose
```{r}
transpose <- function(filepath, file) {
  # Load data
  df <- read.csv(paste0(filepath, file))
  names(df) <- sub("Solar.Radiation..S.LIB.", "", names(df))
  names(df) <- sub("...W.m.2..RX3000_BRW6", "", names(df))
  names(df) <- sub("Wind.Direction..S.WDA.21401803.", "", names(df))
  names(df) <- sub("Wind.Speed..S.WSB.", "", names(df))
  names(df) <- sub("......RX3000_BRW6", "", names(df))
  names(df) <- sub("_BEO", "", names(df))
  df <- df[, names(df) !="21401803.21390415.1.1"]
  names(df) <- sub("21401803.", "", names(df))
  # Transpose data
  df_long <- gather(df, sensor, value, 2:ncol(df))
  df_long <- cSplit(df_long, 'sensor', sep=".", direction = "wide")
  df_long$sensor_1 <- gsub('X', '', df_long$sensor_1)
  df_long <- select(df_long, Date, sensor_1, sensor_2, value)
  colnames(df_long)[2:3] <- c("sensor", "depth")
  df_long$Date <- as.POSIXct(df_long$Date,format="%Y-%m-%d %H:%M:%S",tz="UTC")
  return(df_long)
}
```



## Moisture

**List moisture files**
```{r}
vwc_files <- filenames[substr(filenames, 1, 5) == "Moist"]
vwc_files
```


### BEO

```{r}
file <- "Moist_data_CLEAN_BEO_from_2022_06_13_to_2024_08_31_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "vwc", "_gap_filled_", "BEO", ".csv"))
```


### TNHA

```{r}
file <- "Moist_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "vwc", "_gap_filled_", "TNHA", ".csv"))
```


### SSMH

```{r}
file <- "Moist_data_CLEAN_SSMH_from_2022_06_08_to_2024_08_28_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "vwc", "_gap_filled_", "SSMH", ".csv"))
```




## Temperature
(Ground and Air Combined)

**List temp files**
```{r}
temp_files <- filenames[substr(filenames, 1, 4) == "Temp"]
temp_files
```


### BEO

```{r}
file <- "Temp_data_CLEAN_BEO_from_2022_06_13_to_2024_08_31_no_gps.csv" 

# Call Transpose Function
final <- transpose(filepath, file)
final <- final %>% filter(sensor != "21390850")
write.csv(final, paste0(exportpath, "temp", "_gap_filled_", "BEO", ".csv"))
```

### TNHA

```{r}
file <- "Temp_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "temp", "_gap_filled_", "TNHA", ".csv"))
```

### SSMH

```{r}
file <- "Temp_data_CLEAN_SSMH_from_2022_06_08_to_2024_08_28_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "temp", "_gap_filled_", "SSMH", ".csv"))
```




## Solar Radiation

**List solar files**
```{r}
solar_files <- filenames[substr(filenames, 1, 5) == "Solar"]
solar_files
```


### BEO

```{r}
file <- "Solar_data_CLEAN_BEO_from_2022_06_13_to_2024_08_31_no_gps.csv" 

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "solar", "_gap_filled_", "BEO", ".csv"))
```

### TNHA

```{r}
file <- "Solar_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "solar", "_gap_filled_", "TNHA", ".csv"))
```

### SSMH

```{r}
file <- "Solar_data_CLEAN_SSMH_from_2022_06_08_to_2024_09_20_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "solar", "_gap_filled_", "SSMH", ".csv"))
```




## Wind

wind: sensor-1 is windspeed sensor-3 is winddirection

**List wind files**
```{r}
wind_files <- filenames[substr(filenames, 1, 4) == "Wind"]
wind_files
```


### BEO

```{r}
file <- "Wind_speed_data_CLEAN_BEO_from_2022_06_13_to_2024-11-15_no_gps.csv"   
# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "wind_speed", "_gap_filled_", "BEO", ".csv"))


file <- "Wind_direction_data_BEO_from_2022_06_13_to_2024-11-15_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "wind_direction", "_gap_filled_", "BEO", ".csv"))
```

### TNHA

```{r}
file <- "Wind_speed_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv" 

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "wind_speed", "_gap_filled_", "TNHA", ".csv"))

file <- "Wind_direction_data_CLEAN_TNHA_from_2022_05_17_to_2024_09_26_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "wind_direction", "_gap_filled_", "TNHA", ".csv"))
```

### SSMH

```{r}
file <- "Wind_speed_data_CLEAN_SSMH_from_2022_05_17_to_2024-11-15_no_gps.csv" 

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "wind_speed", "_gap_filled_", "SSMH", ".csv"))

file <- "Wind_direction_data_CLEAN_SSMH_from_2022_05_17_to_2024-11-15_no_gps.csv"

# Call Transpose Function
final <- transpose(filepath, file)
write.csv(final, paste0(exportpath, "wind_direction", "_gap_filled_", "SSMH", ".csv"))
```

