---
title: "Visualize Data"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

Load Data
```{r}
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/Analysis_Ready_Data/"
df <- read.csv(paste0(filepath, "daily_2022_2024.csv"))
df$fullname = as.factor(df$fullname)
df$date <- as.Date(df$date)
```


Load Library
```{r}
library(ggplot2)
library(dplyr)
```


```{r}
tnha_2022 <- df %>% filter(date < "2023-01-01") %>% filter(site == "TNHA") %>% filter(grounddepth == 8)
ggplot(tnha_2022) + geom_line(aes(date, groundtemp, color=fullname)) + scale_x_date(date_breaks = "1 month",
             date_labels = "%B")
```
Based on the appearance of the data, we cannot use values that were added from the gap filling algorithm prior to the first date of collection. I believe the data started on June 19, 2022 or 2022-06-19.

```{r}
beo_2022 <- df %>% filter(date < "2023-01-01") %>% filter(site == "BEO") %>% filter(grounddepth == 8)
ggplot(beo_2022) + geom_line(aes(date, groundtemp, color=fullname))
```
I would say this holds for BEO because we are not really getting an accurate representation of how the data should be acting.

```{r}
ssmh_2022 <- df %>% filter(date < "2023-01-01") %>% filter(site == "SSMH") %>% filter(grounddepth == 8)
ggplot(ssmh_2022) + geom_line(aes(date, groundtemp, color=fullname))
```

Some of the data for SSMH look inaccurate, so we should be careful when proceeding for these stations. However, it also looks okay prior to the initial collection date.