---
title: "PCA - Spring 2025"
author: "Elizabeth Van Metre"
date: "2025-02-04"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

========

# SETUP

========

## FILEPATH:

```{r}
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/Analysis_Ready_Data/"
```

## PICK DATE RANGES:

```{r}
years = c("2022", "2023")
seasons = c("Spring", "Summer","Fall")
sites = c("TNHA", "BEO", "SSMH")

# put season values for the season that has the start of the data
date_start = "2022-06-01" # data starts in June 2022 (YEAR-MO-DY) where day is always 01
# put season values for the season that has the last of the data
date_end = "2023-11-01" # data ends after November of 2023 (will get data up UNTIL date_end not after)
```

## PICK OUTPUT:

```{r}
scree = F # scree plot
eigen = T # eigenvectors and eigenvalues
```

### Dates (for selecting date ranges; no need to edit)

```{r}
spring_months = c("March", "April", "May")
summer_months = c("June","July","August")
fall_months = c("September", "October", "November")
winter_months = c("December", "January", "February")
```

```{r}
spring_dates = data.frame(months=spring_months, start=c("-03-01","-04-01","-05-01"), end=c("-04-01","-05-01","-06-01"))

summer_dates = data.frame(months=summer_months, start=c("-06-01","-07-01","-08-01"), end=c("-07-01","-08-01","-09-01"))

fall_dates = data.frame(months=fall_months, start=c("-09-01","-10-01","-11-01"), end=c("-10-01","-11-01","-12-01"))

winter_dates = data.frame(months=winter_months, start=c("-12-01","-01-01","-02-01"), end=c("-01-01","-02-01","-03-01"))
```

```{r}
all_dates = data.frame(matrix(nrow = 0, ncol = 4))
for(yur in years){
  # spring
  if("Spring" %in% seasons){
    for(i in c(1:3)){
      curdate = as.Date(paste0(yur, spring_dates[i, 2]))
      if(curdate >= date_start && curdate < date_end){
        all_dates <- rbind(all_dates, c(spring_dates[i,1], paste0(yur, spring_dates[i, 2]), paste0(yur, spring_dates[i, 3]),"Spring"))
      }
    }
  }
  if("Summer" %in% seasons){
    for(i in c(1:3)){
      curdate = as.Date(paste0(yur, summer_dates[i, 2]))
      if(curdate >= date_start && curdate < date_end){
        all_dates <- rbind(all_dates, c(summer_dates[i,1], paste0(yur, summer_dates[i, 2]), paste0(yur, summer_dates[i, 3]),"Summer"))
      }
    }
  }
  if("Fall" %in% seasons){
    for(i in c(1:3)){
      curdate = as.Date(paste0(yur, fall_dates[i, 2]))
      if(curdate >= date_start && curdate < date_end){
        all_dates <- rbind(all_dates, c(fall_dates[i,1], paste0(yur, fall_dates[i, 2]), paste0(yur, fall_dates[i, 3]),"Fall"))
      }
    }
  }
  if("Winter" %in% seasons){
    for(i in c(1:3)){
      curdate = as.Date(paste0(yur, winter_dates[i, 2]))
      if(curdate >= date_start && curdate < date_end){
        all_dates <- rbind(all_dates, c(winter_dates[i,1], paste0(yur, winter_dates[i, 2]), paste0(yur, winter_dates[i, 3]),"Winter"))
      }
    }
  }
  
}
colnames(all_dates) = c("months","start","end","szn")
```


## Packages

```{r message=FALSE}
library(dplyr)
library(lubridate)
library(tidyverse)
```

=========================

# DEFINE FUNCTIONS 

=========================

Temporal Range: Season Vertical Spatial Range: 30-45 cm Horizontal
Spatial Range: stations across site (TNHA, SSMH, BEO) --\> Average Total
Site --\> North vs South (except for BEO)

### Filter by Site and Join Tables

```{r}
pick_site <- function(cursite){

  big_df <<- daily %>% filter(site == cursite)
  
  if(szn == "Winter"){
    big_df <<- big_df %>% select(-solar)
  }
  return(big_df)
}
```

### Filter by Date Range

```{r}
pick_dates <- function(datemin, datemax, big_df){
  pca_df <<- big_df %>% filter(Date >= datemin) %>% filter(Date < datemax)
  
  # get rid of NAs
  pca_df <<- na.omit(pca_df)
  pca_df <<- unique(pca_df)
  return(pca_df)
}

```

## Calculate PCA

```{r}
calc_pca <- function(pca_df){
  pca <<- prcomp(pca_df[,6:(ncol(pca_df)-2)], center=TRUE, scale.=TRUE)

  #take out variables
  sd <- pca$sdev
  loads <<- pca$rotation
  rownames(loads) <<- colnames(pca_df[6:(ncol(pca_df)-2)])
  scores <<- pca$x
  
  var <- sd^2
  varPercent <- var/sum(var) * 100
  
  return(list("pca"=pca, "loads"=loads))
}
```

### Make Scree Plot

```{r}
make_scree <- function(pca){
  sd <- pca$sdev
  
  var <- sd^2
  varPercent <- var/sum(var) * 100
  
  barplot(varPercent, xlab="PC", ylab="Percent Variance", names.arg=1:length(varPercent), 
          las=1, ylim=c(0, max(varPercent)), col="gray")
  abline(h=1/ncol(pca_df[5:ncol(pca_df)])*100, col="red")
}
```

### Display Eigenvectors and Eigenvalues

```{r}
make_eigen <- function(pca){
  eigenvectors <- pca$rotation
  print("Eigenvectors (Loadings):")
  print(eigenvectors)
  
  print("Loadings Cutoff:")
  sqrt(1/ncol(pca_df[5:ncol(pca_df)])) # cutoff for "important" loadings
  
  # Access the eigenvalues (variances of the principal components)
  eigenvalues <- (pca$sdev)^2
  print("Eigenvalues:")
  print(eigenvalues)
}
```

===============

# PCA PLOTS 

===============

Load all data
```{r}
daily <- read.csv(paste0(filepath, "daily_2022_2024.csv"))
daily <- daily %>% select(-X) %>% select(-X.1) # get rid of index columns
daily$Date <- as.POSIXct(daily$date, format="%Y-%m-%d") # format dates
daily$fullname[daily$site == "BEO"] <- "BEO-BASE"
daily <- daily %>% filter(fullname == "TNHA-SA" | fullname == "TNHA-SC" | fullname == "SSMH-SB" | fullname == "SSMH-SA" | fullname == "BEO-BASE") %>% select(-c(windspeed, winddir, date)) %>% mutate(aspect = case_when(fullname == "TNHA-SC" | fullname == "SSMH-SB" ~ "North", fullname == "TNHA-SA" | fullname == "SSMH-SA" ~ "South", .default = "N/A")) %>% filter(grounddepth == 8) %>% filter(Date >= "2022-06-19") # create "aspect" column and filter for top depth of soil and start date of when we started collecting data
# note: the data before June 19, 2022 was estimated by our gap-filling script and should be disregarded due to extrapolation
```


```{r, fig.width=8, fig.height=8}
cursite = "BEO"
make_pca <- function(pca_df, szn, yr, cursite){
  if (cursite == "BEO")
  {
    SOUTH <<- pca_df$site == cursite
    # n <- "BEO"
  } else {
    SOUTH <<- pca_df$site == cursite & pca_df$aspect == "South"
    NORTH <<- pca_df$site == cursite & pca_df$aspect == "North"
    s <- pca_df$fullname[SOUTH][1]
    n <- pca_df$fullname[NORTH][1]
  }

    scaling <- 2
  textNudge <- 1.1
  limNudge <- 1.3
  
  xlimit <- seq(floor(min(scores[,1])*limNudge),ceiling(max(scores[,1])*limNudge), 1)
  ylimit <- seq(floor(min(scores[,2])*limNudge),ceiling(max(scores[,2])*limNudge), 1)
  
  plot(scores[, 1], scores[, 2], xlab="PCA 1", ylab="PCA 2", type="n", asp=1, 
       las=1, xaxt='n', yaxt='n')
  
  axis(side = 1, at=xlimit)
  axis(side = 2, at=ylimit)

  if (cursite == "BEO")
  {
    nvstext <- " "
  } else {
    nvstext <- " North v. South "
  }
  
  mindate = format(as.Date(min(pca_df$Date)), format="%B %d %Y")
  maxdate = format(as.Date(max(pca_df$Date)), format="%B %d %Y")
  
  title(paste0(szn, " ", yr," Principal Component Analysis:\n", site, nvstext, "\n(", mindate," - ", maxdate, ")"), adj=0.5)
  

  points(scores[SOUTH, 1], scores[SOUTH, 2], pch=16, cex=1, col="mediumturquoise")
   
  if(cursite != "BEO"){
    points(scores[NORTH, 1], scores[NORTH, 2], pch=16, cex=1, col="salmon")
     legend(x = "topright",          # Position
       legend = c(paste0(s, " (south)"), paste0(n, " (north)")),  # Legend texts
       col = c("mediumturquoise","salmon"),
       pch = 19)  #colors
  
  } else{
    legend(x = "topright",          # Position
       legend = "BEO",  # Legend texts
       col = "mediumturquoise",
       pch = 19) 
    
  }
  
   
  
  arrows(0, 0, loads[, 1]* scaling, loads[, 2]* scaling, length=0.1, angle=20, col="darkred", lwd=1.4)
   
  text(loads[1, 1]*scaling*(textNudge+0.4), loads[1, 2]*scaling*textNudge, rownames(loads)[1],   col="darkred", cex=1) # ground label
  
  text(loads[2, 1]*scaling*textNudge, loads[2, 2]*scaling*textNudge, rownames(loads)[2],   col="darkred", cex=1) # vwc label
  
  if(nrow(loads) > 2){
    text(loads[3, 1]*scaling*(textNudge+0.2), loads[3, 2]*scaling*textNudge, rownames(loads)[3],   col="darkred", cex=1) # airtemp label
  
    if(nrow(loads)>3){
      text(loads[4, 1]*scaling*textNudge-0.2, loads[4, 2]*scaling*textNudge, rownames(loads)[4],   col="darkred", cex=1) # solar or wind label
      
      if(nrow(loads)>4){
  
  text(loads[5, 1]*scaling*textNudge, loads[5, 2]*scaling*textNudge, rownames(loads)[5],   col="darkred", cex=1) # solar or wind label
    }

    }
  
  
  }
  
  
 
  #text(-3, 1]*scaling*textNudge, 1, "TNHA-SA \n(south)", col="mediumturquoise")
  #text(1, 1, "TNHA-SC \n(north)", col="salmon")
}
```

```{r, fig.width=6, fig.height=6}
for(i in c(1:nrow(all_dates))){
  
  month <- all_dates$months[i]
  startdate <- all_dates$start[i]
  enddate <- all_dates$end[i]
  szn <<- all_dates$szn[i]
  yr <<- substr(all_dates$start[i], 1, 4)
  
  for(site in sites){
    big_df <- pick_site(site)
    pca_df <- pick_dates(startdate, enddate, big_df)
    
    if(nrow(pca_df) > 4){
      p <- calc_pca(pca_df)
      pca <- p$pca
      loads <- p$loads
      if(scree == T){
        make_scree(pca)
      }
      if(eigen == T){
        make_eigen(pca)
      }
      make_pca(pca_df, szn, yr, site)
              
    }
  }
  
}
```

If you want to just make one plot (for testing):
```{r, fig.show='hide'}
  # look at all_dates and pick a row to use (set to i)

i <- 8
site <- "BEO"

month <- all_dates$months[i]
startdate <- all_dates$start[i]
enddate <- all_dates$end[i]
szn <<- all_dates$szn[i]
yr <<- substr(all_dates$start[i], 1, 4)

big_df <- pick_site(site)
pca_df <- pick_dates(startdate, enddate, big_df)


if(nrow(pca_df > 0)){
    p <- calc_pca(pca_df)
    pca <- p$pca
    loads <- p$loads # not used at the moment
    if(scree == T){
      make_scree(pca)
    }
    if(eigen == T){
      make_eigen(pca)
    }
    make_pca(pca_df, szn, yr, site)
            
  }


```

