---
title: "HTML Template for Utqiagvik Research at UVA"
author: "Template created by Elizabeth Van Metre"
date: "5/31/2025"
output: 
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

# R Markdown Guide

You can use this as a template for R Markdown files to output HTML files in a clean, concise, and consistent way.

There are many options you can use in your document, detailed in the following sections.
You can also find a cheatsheet directly from posit (the creators of RStudio) [here](https://rstudio.github.io/cheatsheets/html/rmarkdown.html).

## HTML Doc Settings

This template is using the following settings:

-   theme: yeti - theme for the html document

-   toc: true - the page displays a table of contents

-   toc_float: true - the table of contents floats on the side of the page so you can see it when you scroll

-   df_print: paged - dataframes are printed with page options for the user to navigate

-   code_folding: hide - default to hidden code, but user can show code if desired (for cleanliness)

# Formatting

## Chunks

Chunk settings can be set either in an individual chunk or in the initial r setup, using `knitr::opts_chunk$set()`.

The following chunk options are most helpful, but more can be found [here](https://rstudio.github.io/cheatsheets/html/rmarkdown.html#chunk-options).

| Option | Default | Description |
|-----------------|-----------------|--------------------------------------|
| `echo` | `TRUE` | Display code in the html output. |
| `error` | `FALSE` | Display error messages in the render, when an error occurs. |
| `include` | `TRUE` | Include chunk in the output after running. |
| `message` | `TRUE` | Display code messages in the output. |
| `warning` | `TRUE` | Display code warnings in the output. |
| `results` | `"markup"` | `"asis"` (pass through results), `"hide"` (don't display results), `"hold"` (put results below all code), `"markup"` (will format output with markdown, if applicable) |

## Graphs & Figures

| Option | Default | Description |
|-----------------|-----------------|--------------------------------------|
| `fig.align` | "default" | `"left"`, `"right"`, or `"center"` |
| `fig.alt` | NULL | alt text for a figure (good for accessibility!) |
| `fig.cap` | NULL | figure caption |
| `fig.path` | "figure/" | prefix for generating file paths |
| `fig.width` & `fig.height` | 7 | plot dimensions (inches) |
| `out.width` |  | rescales output width, e.g. `"75%"`, `"300px"` |

## Tables

| Option | Value Type | Description |
|-------------------|-------------------|-----------------------------------|
| max.print | number | The number of rows to print. |
| rows.print | number | The number of rows to display. |
| cols.print | number | The number of columns to display. |
| cols.min.print | number | The minimum number of columns to display. |
| pages.print | number | The number of pages to display under page navigation. |
| paged.print | boolean | When set to FALSE turns off paged tables. |
| rownames.print | boolean | When set to FALSE turns off row names. |

An Example Code Chunk with `cols.print=3, rows.print=6`:

```{r cols.print=3, rows.print=6}
mtcars
```
