---
title: "Double Y Axis"
author: "Elizabeth Van Metre"
date: "2023-09-25"
output: html_document
---


```{r}
#Load files from Github
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/elizabeth%20seasonal%20analysis/sourcedata/"

gtfile = "grndtemp_daily.csv"
airfile = "airtemp_daily.csv"
vwcfile = "vwc_daily.csv"
solfile = "solar_daily.csv"
windspeed = "windspeed_daily.csv"
winddir = "winddir_daily.csv"

#30 cm depth
#15-30 VWC
#average sites
```


##Loading packages
```{r}
library(dplyr)
library(lubridate)
library(tidyverse)
library(ggplot2)
library(patchwork) 
#library(hrbrthemes) used in EMR, not sure what it is
```

## Coloring

### TNHA
scale_color_manual(values =c("TNHA-BASE" = "#1B9E77" ,"TNHA-SA" = "#D95F02", "TNHA-SB"="#7570B3", "TNHA-SC" = "#E7298A", "TNHA-SD"="#66A61E"))

Solar - Orange (#D95F02)
Air - Pink (#E7298A)
Ground - Lavender (#7570B3)
VWC - teal (#1B9E77)
Wind - Green (#66A61E)


Load Data
```{r}
grndtmp = read.csv(paste0(filepath, gtfile))
airtmp = read.csv(paste0(filepath, airfile))
vwc = read.csv(paste0(filepath, vwcfile))
solar = read.csv(paste0(filepath, solfile))
windspeed = read.csv(paste0(filepath, wspeed))
winddir = read.csv(paste0(filepath, wdir))
```


## Picking

```{r}
tnha_ground = grndtmp %>% filter(station == "TNHA") %>% filter(depth == 10)
tnha_vwc = vwc %>% filter(station == "TNHA") %>% filter(depth == 2)
tnha_air = airtmp %>% filter(station == "TNHA")
tnha_wind = windspeed %>% filter(station == "TNHA")
tnha_solar = solar %>% filter(station == "TNHA")
tnha = full_join(tnha_ground, tnha_air, by=c("date", "station", "sensor")) %>% select(c(2:4,6,9))
colnames(tnha)[2:5] = c("station","sensor","groundtemp", "airtemp")
tnha = full_join(tnha, tnha_wind, by=c("date", "station", "sensor")) %>% select(c(1:5,8))
colnames(tnha)[c(2:3,6)] = c("station","sensor","windspeed")
tnha = full_join(tnha, tnha_solar, by=c("date", "station", "sensor")) %>% select(c(1:6,9))
colnames(tnha)[c(2:3,7)] = c("station","sensor","solar")
tnha = full_join(tnha, tnha_vwc, by=c("date", "station", "sensor")) %>% select(c(1:7,10))
colnames(tnha)[8] = c("vwc")
```

Average for Each Day All Sensors
```{r}
tnha2 = tnha %>% group_by(date) %>% summarize(groundtemp = mean(groundtemp, na.rm=T), 
                                             airtemp = mean(airtemp, na.rm=T),
                                             windspeed = mean(windspeed, na.rm=T), 
                                             solar = mean(solar, na.rm=T),
                                             vwc = mean(vwc, na.rm=T))
tnha2 <- tnha2 %>% mutate_all(~ifelse(is.nan(.), NA, .))
```



Get Rid of Rows with NAs
```{r}
#tnha = na.omit(tnha)
tnha2$date = as.Date(tnha2$date)
```


Fixing Gaps
```{r}
d = rep(seq(as.Date(tnha2$date[1]), as.Date(tnha2$date[nrow(tnha2)]), by="day"),5)
val = rep(1, length(date))
df = data.frame(d, val)
colnames(df) = c("date","value")
df2 = left_join(df, tnha2, by=c("date"))
```




X-Axis: Date
Y-Axis 1: Solar Radiation
Y-Axis 2: Ground Temperature (+Air Temperature?)

**TNHA missing data in winter?

```{r fig8, fig.height = 6, fig.width = 8}
#Coefficient to scale Y-Axes
coeff1 = 10
coeff2 = 1/20

ggplot(tnha2, aes(x=date)) +
  
  geom_line( aes(y=solar, color="#D95F02"), linewidth=0.4) +
  geom_line( aes(y=(groundtemp+10)*coeff1, color="#7A0177"), linewidth=0.4) +
  geom_line( aes(y=(airtemp+10)*coeff1, color="#6BAED6"), linewidth=0.4) +
  
  
  scale_x_date(date_breaks = "1 month",date_labels = "%b %Y")  +
  
  scale_y_continuous(
    
    
    # Features of the first axis
    name = "Solar Radiation ("~W/m^2~")", 
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./coeff1, name="Temperature (°C)") ###making the 1 negative in that function switches the values so positive is on top
  ) + 
  labs(x="Date")+
  

  theme_bw()+
  

#Solar - Orange (#D95F02)
#Air - Pink (#E7298A)
#Ground - Lavender (#7570B3)
#VWC - teal (#1B9E77)

  
  scale_color_manual(name=" ",labels=c("#7A0177"="Daily Average Ground Temperature (°C)","#6BAED6"="Daily Average Air Temperature (°C)","#D95F02"="Daily Average Solar Radiation ("~W/m^2~")"),values=c("#7A0177" = "#7A0177","#D95F02" = "#D95F02","#6BAED6" = "#6BAED6"))+

  theme(
    axis.title.y = element_text(color ='black', size=10, hjust = 0.5, family="Times New Roman"),
    axis.title.y.right = element_text(color = 'black', size=10, hjust = 0.5, family="Times New Roman"),
    axis.text.x = element_text(angle = 90, size = 8, vjust = 1, hjust=1, family="Times New Roman"),
    axis.title.x = element_text(color = 'black', size=10, hjust = 0.5,vjust = -1, family="Times New Roman"),
    plot.title = element_text(hjust = 0.5, size=16, family="Times New Roman"),
    text=element_text(family="Times New Roman")
  ) +

  ggtitle("Daily Average Solar Radiation, Air, and Ground Temperatures in Barrow") +
 theme(axis.text.x= element_text(size=7))+
   theme(axis.text.x = element_text(angle = 45)) + 
  theme(axis.text.y = element_text(size=7))+ ###these lines are to adjust the axis size and to change the angle 
  theme(legend.position="bottom")+
  

  guides(color=guide_legend(nrow=3, byrow=TRUE))
```



X-Axis: Date
Y-Axis 1: VWC
Y-Axis 2: Ground Temperature +Air Temperature

```{r fig8, fig.height = 6, fig.width = 8}
#Coefficient to scale Y-Axes
coeff1 = 1/20

ggplot(tnha2, aes(x=date)) +
  
  geom_line( aes(y=vwc, color="#E7298A"), linewidth=0.4) +
  geom_line( aes(y=(groundtemp+10)*coeff1, color="#7A0177"), linewidth=0.4) +
  geom_line( aes(y=(airtemp+10)*coeff1, color="#6BAED6"), linewidth=0.4) +
  
  
  scale_x_date(date_breaks = "1 month",date_labels = "%b %Y")  +
  
  scale_y_continuous(
    
    
    # Features of the first axis
    name = "Volumetric Water Content ("~m^3~"/"~m^3~")", 
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./coeff1, name="Temperature (°C)") ###making the 1 negative in that function switches the values so positive is on top
  ) + 
  labs(x="Date")+
  

  theme_bw()+
  

#Solar - Orange (#D95F02)
#Air - Pink (#E7298A)
#Ground - Lavender (#7570B3)
#VWC - teal (#1B9E77)

  
  scale_color_manual(name=" ",labels=c("#7A0177"="Daily Average Ground Temperature (°C)","#6BAED6"="Daily Average Air Temperature (°C)","#E7298A"="Daily Average Volumetric Water Content ("~m^3~"/"~m^3~")"),values=c("#7A0177" = "#7A0177","#E7298A" = "#E7298A","#6BAED6" = "#6BAED6"))+

  theme(
    axis.title.y = element_text(color ='black', size=10, hjust = 0.5, family="Times New Roman"),
    axis.title.y.right = element_text(color = 'black', size=10, hjust = 0.5, family="Times New Roman"),
    axis.text.x = element_text(angle = 90, size = 8, vjust = 1, hjust=1, family="Times New Roman"),
    axis.title.x = element_text(color = 'black', size=10, hjust = 0.5,vjust = -1, family="Times New Roman"),
    plot.title = element_text(hjust = 0.5, size=16, family="Times New Roman"),
    text=element_text(family="Times New Roman")
  ) +

  ggtitle("Daily Average Ground Moisture, Air, and Ground Temperatures in Barrow") +
 theme(axis.text.x= element_text(size=7))+
   theme(axis.text.x = element_text(angle = 45)) + 
  theme(axis.text.y = element_text(size=7))+ ###these lines are to adjust the axis size and to change the angle 
  theme(legend.position="bottom")+
  

  guides(color=guide_legend(nrow=3, byrow=TRUE))
```


X-Axis: Date
Y-Axis 1: Wind Speed
Y-Axis 2: Ground Temperature +Air Temperature

```{r fig8, fig.height = 6, fig.width = 8}
#Coefficient to scale Y-Axes
coeff1 = 1/20

ggplot(tnha2, aes(x=date)) +
  
  geom_line( aes(y=windspeed, color="#66A61E"), linewidth=0.4) +
  geom_line( aes(y=(groundtemp+10)*coeff1, color="#7A0177"), linewidth=0.4) +
  geom_line( aes(y=(airtemp+10)*coeff1, color="#6BAED6"), linewidth=0.4) +
  
  
  scale_x_date(date_breaks = "1 month",date_labels = "%b %Y")  +
  
  scale_y_continuous(
    
    
    # Features of the first axis
    name = "Wind Speed (units)", 
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./coeff1, name="Temperature (°C)") ###making the 1 negative in that function switches the values so positive is on top
  ) + 
  labs(x="Date")+
  

  theme_bw()+
  

#Solar - Orange (#D95F02)
#Air - Pink (#E7298A)
#Ground - Lavender (#7570B3)
#VWC - teal (#1B9E77)

  
  scale_color_manual(name=" ",labels=c("#7A0177"="Daily Average Ground Temperature (°C)","#6BAED6"="Daily Average Air Temperature (°C)","#66A61E"="Daily Average Wind Speed (units)"),values=c("#7A0177" = "#7A0177","#66A61E" = "#66A61E","#6BAED6" = "#6BAED6"))+

  theme(
    axis.title.y = element_text(color ='black', size=10, hjust = 0.5, family="Times New Roman"),
    axis.title.y.right = element_text(color = 'black', size=10, hjust = 0.5, family="Times New Roman"),
    axis.text.x = element_text(angle = 90, size = 8, vjust = 1, hjust=1, family="Times New Roman"),
    axis.title.x = element_text(color = 'black', size=10, hjust = 0.5,vjust = -1, family="Times New Roman"),
    plot.title = element_text(hjust = 0.5, size=16, family="Times New Roman"),
    text=element_text(family="Times New Roman")
  ) +

  ggtitle("Daily Average Wind Speed, Air, and Ground Temperatures in Barrow") +
 theme(axis.text.x= element_text(size=7))+
   theme(axis.text.x = element_text(angle = 45)) + 
  theme(axis.text.y = element_text(size=7))+ ###these lines are to adjust the axis size and to change the angle 
  theme(legend.position="bottom")+
  

  guides(color=guide_legend(nrow=3, byrow=TRUE))
```

