---
title: "PCA - Spring 2024"
author: "Elizabeth Van Metre"
date: "2024-03-27"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

========

# SETUP

========

## FILEPATH:

```{r}
filepath = "/Users/emvanmetre/Desktop/Arctic/X-Y_Plots/data/processed/Combined/"
```

## PICK DATE RANGES:

```{r}
years = c("2022", "2023")
seasons = c("Spring", "Summer","Fall")
sites = c("TNHA", "BEO", "SSMH")

# put season values for the season that has the start of the data
date_start = "2022-06-01" # data starts in June 2022 (YEAR-MO-DY) where day is always 01
# put season values for the season that has the last of the data
date_end = "2023-11-01" # data ends after November of 2023 (will get data up UNTIL date_end not after)
```

## PICK OUTPUT:

```{r}
scree = F # scree plot
eigen = T # eigenvectors and eigenvalues
```

### Dates (for selecting date ranges; no need to edit)

```{r}
spring_months = c("March", "April", "May")
summer_months = c("June","July","August")
fall_months = c("September", "October", "November")
winter_months = c("December", "January", "February")
```

```{r}
spring_dates = data.frame(months=spring_months, start=c("-03-01","-04-01","-05-01"), end=c("-04-01","-05-01","-06-01"))

summer_dates = data.frame(months=summer_months, start=c("-06-01","-07-01","-08-01"), end=c("-07-01","-08-01","-09-01"))

fall_dates = data.frame(months=fall_months, start=c("-09-01","-10-01","-11-01"), end=c("-10-01","-11-01","-12-01"))

winter_dates = data.frame(months=winter_months, start=c("-12-01","-01-01","-02-01"), end=c("-01-01","-02-01","-03-01"))
```

```{r}
all_dates = data.frame(matrix(nrow = 0, ncol = 4))
for(yur in years){
  # spring
  if("Spring" %in% seasons){
    for(i in c(1:3)){
      curdate = as.Date(paste0(yur, spring_dates[i, 2]))
      if(curdate >= date_start && curdate < date_end){
        all_dates <- rbind(all_dates, c(spring_dates[i,1], paste0(yur, spring_dates[i, 2]), paste0(yur, spring_dates[i, 3]),"Spring"))
      }
    }
  }
  if("Summer" %in% seasons){
    for(i in c(1:3)){
      curdate = as.Date(paste0(yur, summer_dates[i, 2]))
      if(curdate >= date_start && curdate < date_end){
        all_dates <- rbind(all_dates, c(summer_dates[i,1], paste0(yur, summer_dates[i, 2]), paste0(yur, summer_dates[i, 3]),"Summer"))
      }
    }
  }
  if("Fall" %in% seasons){
    for(i in c(1:3)){
      curdate = as.Date(paste0(yur, fall_dates[i, 2]))
      if(curdate >= date_start && curdate < date_end){
        all_dates <- rbind(all_dates, c(fall_dates[i,1], paste0(yur, fall_dates[i, 2]), paste0(yur, fall_dates[i, 3]),"Fall"))
      }
    }
  }
  if("Winter" %in% seasons){
    for(i in c(1:3)){
      curdate = as.Date(paste0(yur, winter_dates[i, 2]))
      if(curdate >= date_start && curdate < date_end){
        all_dates <- rbind(all_dates, c(winter_dates[i,1], paste0(yur, winter_dates[i, 2]), paste0(yur, winter_dates[i, 3]),"Winter"))
      }
    }
  }
  
}
colnames(all_dates) = c("months","start","end","szn")
```


## Packages

```{r message=FALSE}
library(dplyr)
library(lubridate)
library(tidyverse)
```

=========================

# DEFINE FUNCTIONS 

=========================

Temporal Range: Season Vertical Spatial Range: 30-45 cm Horizontal
Spatial Range: stations across site (TNHA, SSMH, BEO) --\> Average Total
Site --\> North vs South (except for BEO)

### Filter by Site and Join Tables

```{r}
pick_site <- function(cursite){
  gtfile = paste0("GroundTemperature_",szn,yr,"_DAILY.csv")
  airfile = paste0("AirTemperature_",szn,yr,"_DAILY.csv")
  vwcfile = paste0("VWC_",szn,yr,"_DAILY.csv")
  solfile = paste0("Solar_",szn,yr,"_DAILY.csv")
  windspeed = paste0("WindSpeed_",szn,yr,"_DAILY.csv")
  winddir = paste0("WindDirection_",szn,yr,"_DAILY.csv")
  
  grndtmp <<- read.csv(paste0(filepath, gtfile))
  airtmp <<- read.csv(paste0(filepath, airfile))
  vwc <<- read.csv(paste0(filepath, vwcfile))
  solar <<- read.csv(paste0(filepath, solfile))
  wind <<- read.csv(paste0(filepath, windspeed))
  
  pca_ground = grndtmp %>% filter(site == cursite) %>% filter(depth == "30cm")
  pca_air = airtmp %>% filter(site == cursite)
  pca_wind = wind %>% filter(site == cursite)
  pca_solar = solar %>% filter(site == cursite)
  pca_vwc = vwc %>% filter(site==cursite) %>% filter(depth == "30-45cm")
  
  big_df <<- full_join(pca_ground, pca_air, by=c("Date", "site", "station", "fullname")) %>% 
    select(Date, site, station, fullname, avg.x, avg.y) %>% rename("groundtemp"="avg.x", "airtemp"="avg.y")
  big_df <<- full_join(big_df, pca_wind, by=c("Date", "site", "station", "fullname")) %>% 
    select(Date, site, station, fullname, groundtemp, airtemp, avg) %>% rename("windspeed" = avg)
  big_df <<- full_join(big_df, pca_solar, by=c("Date", "site", "station", "fullname")) %>% 
    select(Date, site, station, fullname, groundtemp, airtemp, windspeed, avg)  %>% rename("solar" = avg)
  big_df <<- full_join(big_df, pca_vwc, by=c("Date", "site", "station", "fullname")) %>% 
    select(Date, site, station, fullname, groundtemp, airtemp, windspeed, solar, avg)  %>% rename("vwc" = avg)
  
  if(szn == "Winter"){
    big_df <<- big_df %>% select(-solar)
  }
  return(big_df)
}
```

### Filter by Date Range

```{r}
pick_dates <- function(datemin, datemax, big_df){
  pca_df <<- big_df %>% filter(Date >= datemin) %>% filter(Date < datemax)
  
  # get rid of NAs
  pca_df <<- na.omit(pca_df)
  pca_df <<- unique(pca_df)
  return(pca_df)
}

```

## Calculate PCA

```{r}
calc_pca <- function(pca_df){
  pca <<- prcomp(pca_df[,5:ncol(pca_df)], center=TRUE, scale.=TRUE)

  #take out variables
  sd <- pca$sdev
  loads <<- pca$rotation
  rownames(loads) <<- colnames(pca_df[5:ncol(pca_df)])
  scores <<- pca$x
  
  var <- sd^2
  varPercent <- var/sum(var) * 100
  
  return(list("pca"=pca, "loads"=loads))
}
```

### Make Scree Plot

```{r}
make_scree <- function(pca){
  sd <- pca$sdev
  
  var <- sd^2
  varPercent <- var/sum(var) * 100
  
  barplot(varPercent, xlab="PC", ylab="Percent Variance", names.arg=1:length(varPercent), 
          las=1, ylim=c(0, max(varPercent)), col="gray")
  abline(h=1/ncol(pca_df[5:ncol(pca_df)])*100, col="red")
}
```

### Display Eigenvectors and Eigenvalues

```{r}
make_eigen <- function(pca){
  eigenvectors <- pca$rotation
  print("Eigenvectors (Loadings):")
  print(eigenvectors)
  
  print("Loadings Cutoff:")
  sqrt(1/ncol(pca_df[5:ncol(pca_df)])) # cutoff for "important" loadings
  
  # Access the eigenvalues (variances of the principal components)
  eigenvalues <- (pca$sdev)^2
  print("Eigenvalues:")
  print(eigenvalues)
}
```

===============

# PCA PLOTS 

===============

```{r, fig.width=6, fig.height=6}
make_pca <- function(pca_df, szn, yr, site){
  if(site == "TNHA"){
    SOUTH <<- pca_df$fullname == "TNHA-SA"
    NORTH <<- pca_df$fullname == "TNHA-SC"
    s <- "TNHA-SA"
    n <- "TNHA-SC"
  } else{
    if(site == "SSMH"){
      SOUTH <<- pca_df$fullname == "SSMH-SA"
      NORTH <<- pca_df$fullname == "SSMH-SB"
      s <- "SSMH-SA"
      n <- "SSMH-SB"
    } else {
      SOUTH <<- pca_df$fullname == "BEO-BASE"
      n <- "BEO"
    }
  }
  
    scaling <- 2
  textNudge <- 1.1
  limNudge <- 1.3
  
  xlimit <- seq(floor(min(scores[,1])*limNudge),ceiling(max(scores[,1])*limNudge), 1)
  ylimit <- seq(floor(min(scores[,2])*limNudge),ceiling(max(scores[,2])*limNudge), 1)
  
  plot(scores[, 1], scores[, 2], xlab="PCA 1", ylab="PCA 2", type="n", asp=1, 
       las=1, xaxt='n', yaxt='n')
  
  axis(side = 1, at=xlimit)
  axis(side = 2, at=ylimit)
  
  title(paste0(szn, " ", yr," Principal Component Analysis: ",site," North v. South\n",format(as.Date(min(pca_df$Date)), format="%B %d %Y")," - ",format(as.Date(max(pca_df$Date)), format="%B %d %Y")), adj=0.5)
  
 
   
  points(scores[SOUTH, 1], scores[SOUTH, 2], pch=16, cex=0.7, col="mediumturquoise")
   
  if(site != "BEO"){
    points(scores[NORTH, 1], scores[NORTH, 2], pch=16, cex=0.7, col="salmon")
     legend(x = "topright",          # Position
       legend = c(paste0(s, " (south)"), paste0(n, " (north)")),  # Legend texts
       col = c("mediumturquoise","salmon"),
       pch = 19)  #colors
  
  } else{
    legend(x = "topright",          # Position
       legend = "BEO",  # Legend texts
       col = "mediumturquoise",
       pch = 19) 
    
  }
  
   
  
  arrows(0, 0, loads[, 1]* scaling, loads[, 2]* scaling, length=0.1, angle=20, col="darkred")
   
  text(loads[1, 1]*scaling*textNudge, loads[1, 2]*scaling*textNudge, rownames(loads)[1],   col="darkred", cex=0.7) # ground label
  
  text(loads[2, 1]*scaling*textNudge, loads[2, 2]*scaling*textNudge+0.2, rownames(loads)[2],   col="darkred", cex=0.7) # air label
  
  if(nrow(loads) > 2){
    text(loads[3, 1]*scaling*textNudge, loads[3, 2]*scaling*textNudge, rownames(loads)[3],   col="darkred", cex=0.7) # wind label
  
    if(nrow(loads)>3){
      text(loads[4, 1]*scaling*textNudge-0.2, loads[4, 2]*scaling*textNudge, rownames(loads)[4],   col="darkred", cex=0.7) # solar label
      
      if(nrow(loads)>4){
  
  text(loads[5, 1]*scaling*textNudge, loads[5, 2]*scaling*textNudge, rownames(loads)[5],   col="darkred", cex=0.7) # vwc label
    }

    }
  
  
  }
  
  
 
  #text(-3, 1]*scaling*textNudge, 1, "TNHA-SA \n(south)", col="mediumturquoise")
  #text(1, 1, "TNHA-SC \n(north)", col="salmon")
}
```

```{r, fig.width=6, fig.height=6}
for(i in c(1:nrow(all_dates))){
  
  month <- all_dates$months[i]
  startdate <- all_dates$start[i]
  enddate <- all_dates$end[i]
  szn <<- all_dates$szn[i]
  yr <<- substr(all_dates$start[i], 1, 4)
  
  for(site in sites){
    big_df <- pick_site(site)
    pca_df <- pick_dates(startdate, enddate, big_df)
    
    if(nrow(pca_df) > 4){
      p <- calc_pca(pca_df)
      pca <- p$pca
      loads <- p$loads
      if(scree == T){
        make_scree(pca)
      }
      if(eigen == T){
        make_eigen(pca)
      }
      make_pca(pca_df, szn, yr, site)
              
    }
  }
  
}
```

If you want to just make one plot (for testing):
```{r, fig.show='hide'}
  # look at all_dates and pick a row to use (set to i)

i <- 8
site <- "TNHA"

month <- all_dates$months[i]
startdate <- all_dates$start[i]
enddate <- all_dates$end[i]
szn <<- all_dates$szn[i]
yr <<- substr(all_dates$start[i], 1, 4)

big_df <- pick_site(site)
pca_df <- pick_dates(startdate, enddate, big_df)


if(nrow(pca_df > 0)){
    p <- calc_pca(pca_df)
    pca <- p$pca
    loads <- p$loads # not used at the moment
    if(scree == T){
      make_scree(pca)
    }
    if(eigen == T){
      make_eigen(pca)
    }
    make_pca(pca_df, szn, yr, site)
            
  }


```



```{r include=FALSE}
# code to make plots if dates are not set (just use year and season)

#for(year in years){
#  
#  for(season in seasons){
#    szn <<-  season
#    yr <<-  year
#    
#    if(!(paste(szn, yr) %in% do_not_include)){
#    
#      for(site in sites){
#        
#        # load files and pick current site
#        big_df <- pick_site(site)
#        for(i in seq(1,3)){ # month
#          
#          if(season == "Spring"){
#            pca_df <- pick_dates(paste0(year, spring_dates[i, 2]), #paste0(year, spring_dates[i, 3]), big_df)
#          }
#          else if(season == "Summer"){
#            pca_df <- pick_dates(paste0(year, summer_dates[i, 2]), #paste0(year, summer_dates[i, 3]), big_df)
#          }
#          else if(season == "Fall"){
#            pca_df <- pick_dates(paste0(year, fall_dates[i, 2]), paste0(year, #fall_dates[i, 3]), big_df)
#          }
#          else{
#            if(i == 1){
#              pca_df <- pick_dates(paste0(year, winter_dates[i, 2]), #paste0(year, winter_dates[i, 3]), big_df)
#            }
#            else{
#              pca_df <- pick_dates(paste0(as.numeric(year)+1, winter_dates[i, #2]), paste0(as.numeric(year)+1, winter_dates[i, 3]), big_df)
#            }
#            
#          }
#          if(nrow(pca_df > 0)){
#            p <- calc_pca(pca_df)
#            pca <- p$pca
#            loads <- p$loads
#            if(scree == T){
#              make_scree()
#            }
#            if(eigen == T){
#              make_eigen()
#            }
#            make_pca(pca_df, season, year, site)
#              
#          }
#          
#          
#        }
#        
#        
#      }
#    
#  
#      
#    }
#    
#  }
#}
```

```{r include=FALSE}

# ==============
# EXPORT PLOTS
# ==============

# Save the plot
# only run in R, don't run when knitting file

#dev.new(height=7, width=7)
#make_pca()
#dev.copy2pdf(file="PCA_TNHA_Summer2022.pdf")
#dev.off ()
```

```{r, fig.show='hide'}
#==================
## Code Graveyard
#==================

```

```{r, fig.show='hide'}
#The following code makes a bi-plot, which is similar to the code above, however it is not grouped at all. It is a bit harder to read, but it is very straightforward code for a simpler PCA.
```

```{r, fig.show='hide'}
#   ### Bi-plots
#   dev.new(height=7, width=7)
#   biplot(scores[, 1:2], loadings[, 1:2], cex=0.7)
```

```{r, fig.show='hide'}
#NOTE: this is our code graveyard for code that can be used to find other PCA results, including k-means clustering. Above, we use clusters of TNHA-North and TNHA-South. Here, we use k-means clustering to group data points together and compare that to the different stations at TNHA. This can be useful, however we are particularly interested in identifying differences between north- and south-facing sensors.
```

```{r, fig.show='hide', warning=FALSE}
#   ### K Means Clustering
#   pc_data <- summer_tnha_pca$x
#   
#   # Select the first two principal components
#   pc_to_use <- pc_data[, 1:2]
```

```{r, fig.show='hide', warning=FALSE}
#   # Run k-means for different numbers of clusters
#   wcss <- numeric()
#   for (k in 1:10) {
#     kmeans_result <- kmeans(pc_to_use, centers = k)
#     wcss[k] <- kmeans_result$tot.withinss
#   }
#   
#   # Plot the results
#   plot(1:10, wcss, type = "b", xlab = "Number of Clusters", ylab = "Within-Cluster Sum of #   Squares")
```

```{r, fig.show='hide', warning=FALSE}
#   # We can use 2-3 clusters (where the change in slope drops off)
#   # Perform k-means clustering
#   kmeans_result <- kmeans(pc_to_use, centers = 3)  # Change 'centers' based on your specific #   case
```

```{r, fig.show='hide', warning=FALSE}
#   # Add cluster assignments to the data
#   clustered_data <- data.frame(pc_to_use, cluster = kmeans_result$cluster)
#   
#   # Plot the clusters
#   ggplot(clustered_data, aes(x = PC1, y = PC2, color = factor(cluster))) +
#     geom_point() +
#     theme_minimal() +
#     labs(title = "Clusters in Principal Component Space")
```

Find Each Data Point

```{r, fig.show='hide', warning=FALSE}
#   #Want to look at clustered data
#   
#   cluster_sensors = data.frame(clustered_data, summer_tnha$station)
#   
#   cluster1 = filter(cluster_sensors, cluster == 1)
#   cluster2 = filter(cluster_sensors, cluster == 2)
#   cluster3 = filter(cluster_sensors, cluster == 3)
#   
#   cluster_sensors$cluster = as.factor(cluster_sensors$cluster)
#   
#   ggplot(cluster_sensors, aes(x = summer_tnha.station, fill = cluster)) +
#     geom_bar(position = "dodge") +
#     theme_minimal() +
#     labs(title = "Number of Instances in Each Group",
#          x = "Category",
#          y = "Count")
#   
```

```{r, fig.show='hide', warning=FALSE}
#   p = ggplot(cluster_sensors, aes(x = PC1, y = PC2)) +
#     geom_point(aes(color = summer_tnha.station, shape = cluster), size = 2) +
#     scale_shape_manual(values = c(16, 17, 18)) +  # Set shapes manually
#     theme_minimal() +
#     labs(title = "Biplot of Eigenvectors and PC Clusters",
#          x = "PC1",
#          y = "PC2") 
#   p + geom_segment(data = as.data.frame(eigenvectors), aes(x = 0, y = 0, xend = PC1*4, yend = #   PC2*4),
#                    arrow = arrow(type = "closed", length = unit(0.1, "inches")),
#                    linewidth = 0.5, color = "black") +
#     geom_text(data = as.data.frame(eigenvectors), aes(x = PC1*2.5, y = PC2*5+0.1, label = #   rownames(eigenvectors)))
#   
#   
```

