---
title: "Principal Component Analysis"
author: "Elizabeth Van Metre"
date: "2023-09-07"
output: html_document
---

```{r}
filepath = "~/Desktop/Arctic/X-Y_plots/data/processed/summer/"

gtfile = "grndtmp_long.csv"
airfile = "airtemp_long.csv"
vwcfile = "mst_long.csv"
solfile = "solar_long.csv"
windfile = "windspeed_summer.csv"


#30 cm depth
#15-30 VWC
#average sites
```


##Loading packages
```{r}
library(dplyr)
library(lubridate)
library(tidyverse)
```


Load Data
```{r}
grndtmp = read.csv(paste0(filepath, gtfile))
airtmp = read.csv(paste0(filepath, airfile))
vwc = read.csv(paste0(filepath, vwcfile))
solar = read.csv(paste0(filepath, solfile))
wind = read.csv(paste0(filepath, windfile))
```


Temporal Range: Summer
Vertical Spatial Range: 30 cm
Horizontal Spatial Range: by site
  --> Average Total Site
  --> North vs South


# PCA - TNHA Site

## Picking

```{r}
tnha_ground = grndtmp %>% filter(station == "TNHA") %>% filter(depth == 10)
tnha_air = airtmp %>% filter(station == "TNHA")
tnha_wind = wind %>% filter(station == "21198259")
tnha_solar = solar %>% filter(station == "TNHA")
tnha = full_join(tnha_ground, tnha_air, by="Date") %>% select(c(3:4,10))
colnames(tnha)[2:3] = c("groundtemp", "airtemp")
tnha = full_join(tnha, tnha_wind, by="Date") %>% select(c(1:3,6))
colnames(tnha)[4] = "windspeed"
tnha = full_join(tnha, tnha_solar, by="Date") %>% select(c(1:4,7))
colnames(tnha)[5] = "solar"
```

## Hourly Averages

```{r}

tnhahourly <- tnha %>% mutate(day = substr(Date, 1, 10)) %>% mutate(hour = substr(Date, 12, 13))

tnhahourlyavg <- tnhahourly %>% group_by(day, hour) %>% 
  summarize(groundavg = mean(groundtemp, na.rm=F), airavg = mean(airtemp, na.rm=F), windavg = mean(windspeed, na.rm=F), solaravg = mean(solar, na.rm=F))
tnhahourlyavg <- na.omit(tnhahourlyavg)
```

```{r}
tnhasd = data.frame(((tnhahourlyavg$groundavg - mean(tnhahourlyavg$groundavg))/sd(tnhahourlyavg$groundavg)),
                     ((tnhahourlyavg$airavg - mean(tnhahourlyavg$airavg))/sd(tnhahourlyavg$airavg)),
                     ((tnhahourlyavg$windavg - mean(tnhahourlyavg$windavg))/sd(tnhahourlyavg$windavg)),
                     ((tnhahourlyavg$solaravg - mean(tnhahourlyavg$solaravg))/sd(tnhahourlyavg$solaravg)))
colnames(tnhasd) = c("groundtemp", "airtemp", "windspeed", "solar")
```


## PCA
```{r}
tnha_pca <- prcomp(tnhasd, scale.=FALSE)

#take out variables
sd <- tnha_pca$sdev
loadings <- tnha_pca$rotation
rownames(loadings) <- colnames(tnhasd)
scores <- tnha_pca$x

var <- sd^2
varPercent <- var/sum(var) * 100
dev.new() #not sure what this does
```

### Show Results
```{r}
barplot(varPercent, xlab="PC", ylab="Percent Variance", names.arg=1:length(varPercent), las=1, ylim=c(0, max(varPercent)), col="gray")
abline(h=1/ncol(tnhasd)*100, col="red")
```

### Loadings
```{r}
loadings
sqrt(1/ncol(tnhasd)) # cutoff for "important" loadings
```


### Bi-plots
```{r}
dev.new(height=7, width=7)
biplot(scores[, 1:2], loadings[, 1:2], cex=0.7)
```



# PCA - TNHA North vs South
## TNHA-A vs TNHA-SC


## Picking

```{r}
tnha_ground = grndtmp %>% filter(station == "TNHA") %>% filter(depth == 10) %>% 
  filter(sensor=="TNHA-SA" | sensor=="TNHA-SC")

tnha_air = airtmp %>% filter(station == "TNHA") %>% filter(sensor=="TNHA-SA" | sensor=="TNHA-SC")

tnha_wind = wind %>% filter(station == "21198259") %>% filter(sensor=="21398709" | sensor=="21206909")
tnha_wind$sensor[tnha_wind$sensor=="21398709"] = "TNHA-SA"
tnha_wind$sensor[tnha_wind$sensor=="21206909"] = "TNHA-SC"

#tnha_solar = solar %>% filter(station == "TNHA") %>% filter(sensor=="21398620" | sensor=="TNHA-SC")
#tnha_solar$sensor[tnha_solar$sensor=="21398620"] = "TNHA-SA"

tnha = full_join(tnha_ground, tnha_air, by=c("Date","sensor")) %>% select(c(3,6,4,10))
colnames(tnha)[3:4] = c("groundtemp", "airtemp")
tnha = full_join(tnha, tnha_wind, by=c("Date","sensor")) %>% select(c(1:4,7))
colnames(tnha)[5] = "windspeed"
#tnha = full_join(tnha, tnha_solar, by=c("Date","sensor")) %>% select(c(1:5,8))
#colnames(tnha)[6] = "solar"
```

## Hourly Averages

```{r}

tnhahourly <- tnha %>% mutate(day = substr(Date, 1, 10)) %>% mutate(hour = substr(Date, 12, 13))

tnhahourlyavg <- tnhahourly %>% group_by(sensor, day, hour) %>% 
  summarize(groundavg = mean(groundtemp, na.rm=F), airavg = mean(airtemp, na.rm=F), windavg = mean(windspeed, na.rm=F))
tnhahourlyavg <- na.omit(tnhahourlyavg)
```





```{r}
tnhasd = data.frame(((tnhahourlyavg$groundavg - mean(tnhahourlyavg$groundavg))/sd(tnhahourlyavg$groundavg)),
                     ((tnhahourlyavg$airavg - mean(tnhahourlyavg$airavg))/sd(tnhahourlyavg$airavg)),
                     ((tnhahourlyavg$windavg - mean(tnhahourlyavg$windavg))/sd(tnhahourlyavg$windavg)))
colnames(tnhasd) = c("groundtemp", "airtemp", "windspeed")
```




## PCA
```{r}
tnha_pca <- prcomp(tnhasd, scale.=FALSE)

#take out variables
sd <- tnha_pca$sdev
loadings <- tnha_pca$rotation
rownames(loadings) <- colnames(tnhasd)
scores <- tnha_pca$x

var <- sd^2
varPercent <- var/sum(var) * 100
dev.new() #not sure what this does
```

### Show Results
```{r}
barplot(varPercent, xlab="PC", ylab="Percent Variance", names.arg=1:length(varPercent), las=1, ylim=c(0, max(varPercent)), col="gray")
abline(h=1/ncol(tnhasd)*100, col="red")
```

### Loadings
```{r}
loadings
sqrt(1/ncol(tnhasd)) # cutoff for "important" loadings
```


### Bi-plots
```{r}
dev.new(height=7, width=7)
TNHA_SA <- tnhahourlyavg$sensor == "TNHA-SA"
TNHA_SC <- tnhahourlyavg$sensor == "TNHA-SC"
 
plot(scores[, 1], scores[, 2], xlab="PCA 1", ylab="PCA 2", type="n", asp=1, las=1)

scaling <- 5
textNudge <- 1.05
 
points(scores[TNHA_SA, 1], scores[TNHA_SA, 2], pch=16, cex=0.7, col="lightblue")
 
points(scores[TNHA_SC, 1], scores[TNHA_SC, 2], pch=16, cex=0.7, col="pink")
 

arrows(0, 0, loadings[, 1]* scaling, loadings[, 2]* scaling, length=0.1, angle=20, col="darkred")
 
text(loadings[, 1]*scaling*textNudge, loadings[, 2]*scaling*textNudge, rownames(loadings), col="darkred", cex=0.7)

 

text(1, 3, "TNHA-SA \n(south)", col="darkblue")
text(-4, 2, "TNHA-SC \n(north)", col="darkred")
```

# PCA - SSMH

Average
```{r}

```


North vs South
SSMH-D vs SSMH-SB
```{r}

```


# PCA - BEO

JUST ONE SENSOR
BEO-6
```{r}

```




