---
title: "PCA Daily Seasons"
author: "Elizabeth Van Metre"
date: "2023-09-21"
output: html_document
---

Meteorological Fall: September 1 - November 30

```{r}
#github raw filepath
#https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/elizabeth%20seasonal%20analysis/sourcedata/airtemp_daily.csv

filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/elizabeth%20seasonal%20analysis/sourcedata/"

gtfile = "grndtemp_daily.csv"
airfile = "airtemp_daily.csv"
vwcfile = "vwc_daily.csv"
solfile = "solar_daily.csv"
windspeed = "windspeed_daily.csv"
winddir = "winddir_daily.csv"


#30 cm depth
#15-30 VWC
#average sites
```


##Loading packages
```{r}
library(dplyr)
library(lubridate)
library(tidyverse)
```


Load Data
```{r}
grndtmp = read.csv(paste0(filepath, gtfile))
airtmp = read.csv(paste0(filepath, airfile))
vwc = read.csv(paste0(filepath, vwcfile))
solar = read.csv(paste0(filepath, solfile))
wind = read.csv(paste0(filepath, windspeed))
```


Temporal Range: Summer
Vertical Spatial Range: 30 cm
Horizontal Spatial Range: by site
  --> Average Total Site
  --> North vs South


# PCA - TNHA Site

## Picking

```{r}
tnha_ground = grndtmp %>% filter(station == "TNHA") %>% filter(depth == 10)
tnha_air = airtmp %>% filter(station == "TNHA")
tnha_wind = wind %>% filter(station == "TNHA")
tnha_solar = solar %>% filter(station == "TNHA")
tnha_vwc = vwc %>% filter(station=="TNHA") %>% filter(depth == 2)
tnha = full_join(tnha_ground, tnha_air, by=c("date", "station", "sensor")) %>% select(c(2:4,6,9))
colnames(tnha)[2:5] = c("station","sensor","groundtemp", "airtemp")
tnha = full_join(tnha, tnha_wind, by=c("date", "station", "sensor")) %>% select(c(1:5,8))
colnames(tnha)[c(2:3,6)] = c("station","sensor","windspeed")
tnha = full_join(tnha, tnha_solar, by=c("date", "station", "sensor")) %>% select(c(1:6,9))
colnames(tnha)[c(2:3,7)] = c("station","sensor","solar")
tnha = full_join(tnha, tnha_vwc, by=c("date", "station", "sensor")) %>% select(c(1:7,10))
colnames(tnha)[c(2:3,8)] = c("station","sensor","vwc")
```

Get Rid of Rows with NAs
```{r}
tnha = na.omit(tnha)
```


## PCA
```{r}
tnha_pca <- prcomp(tnha[,4:8], center=TRUE, scale.=TRUE)

#take out variables
sd <- tnha_pca$sdev
loadings <- tnha_pca$rotation
rownames(loadings) <- colnames(tnha[4:8])
scores <- tnha_pca$x

var <- sd^2
varPercent <- var/sum(var) * 100
dev.new() #not sure what this does
```

### Show Results
```{r}
barplot(varPercent, xlab="PC", ylab="Percent Variance", names.arg=1:length(varPercent), las=1, ylim=c(0, max(varPercent)), col="gray")
abline(h=1/ncol(tnha[4:8])*100, col="red")
```

### Loadings
```{r}
loadings
sqrt(1/ncol(tnha[4:8])) # cutoff for "important" loadings
```
Eigenvectors and Eigenvalues
```{r}
eigenvectors <- tnha_pca$rotation
print("Eigenvectors (Loadings):")
print(eigenvectors)

# Access the eigenvalues (variances of the principal components)
eigenvalues <- (tnha_pca$sdev)^2
print("Eigenvalues:")
print(eigenvalues)
```



### Bi-plots
```{r}
dev.new(height=7, width=7)
biplot(scores[, 1:2], loadings[, 1:2], cex=0.7)
```

Select a value on here - what are its eigenvalues and eigenvector? What station? Sensor/location?


Clustering
```{r}
pc_data <- tnha_pca$x

# Select the first two principal components
pc_to_use <- pc_data[, 1:2]

# Perform k-means clustering
kmeans_result <- kmeans(pc_to_use, centers = 3)  # Change 'centers' based on your specific case
```

```{r}
# Run k-means for different numbers of clusters
wcss <- numeric()
for (k in 1:10) {
  kmeans_result <- kmeans(pc_to_use, centers = k)
  wcss[k] <- kmeans_result$tot.withinss
}

# Plot the results
plot(1:10, wcss, type = "b", xlab = "Number of Clusters", ylab = "Within-Cluster Sum of Squares")
```
We can use 2-3 clusters (where the change in slope drops off)


```{r}
kmeans_result <- kmeans(pc_to_use, centers = 3)  # Change 'centers' based on your specific case

# Add cluster assignments to the data
clustered_data <- data.frame(pc_to_use, cluster = kmeans_result$cluster)

# Plot the clusters
ggplot(clustered_data, aes(x = PC1, y = PC2, color = factor(cluster))) +
  geom_point() +
  theme_minimal() +
  labs(title = "Clusters in Principal Component Space")
```

Find Each Data Point
```{r}
#Want to look at clustered data

cluster_sensors = data.frame(clustered_data, tnha$sensor)

cluster1 = filter(cluster_sensors, cluster == 1)
cluster2 = filter(cluster_sensors, cluster == 2)
cluster3 = filter(cluster_sensors, cluster == 3)

cluster_sensors$cluster = as.factor(cluster_sensors$cluster)

ggplot(cluster_sensors, aes(x = tnha.sensor, fill = cluster)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(title = "Number of Instances in Each Group",
       x = "Category",
       y = "Count")

```

```{r}
p = ggplot(cluster_sensors, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = tnha.sensor, shape = cluster), size = 2) +
  scale_shape_manual(values = c(16, 17, 18)) +  # Set shapes manually
  theme_minimal() +
  labs(title = "Biplot of Eigenvectors and PC Clusters",
       x = "PC1",
       y = "PC2") 
p + geom_segment(data = as.data.frame(eigenvectors), aes(x = 0, y = 0, xend = PC1*4, yend = PC2*4),
                 arrow = arrow(type = "closed", length = unit(0.1, "inches")),
                 linewidth = 0.5, color = "black") +
  geom_text(data = as.data.frame(eigenvectors), aes(x = PC1*2.5, y = PC2*5+0.1, label = rownames(eigenvectors)))


```



# PCA - TNHA North vs South
## TNHA-A vs TNHA-SC


## Picking

```{r}
tnha_ground = grndtmp %>% filter(station == "TNHA") %>% filter(depth == 10) %>% 
  filter(sensor=="TNHA-SA" | sensor=="TNHA-SC")

tnha_air = airtmp %>% filter(station == "TNHA") %>% filter(sensor=="TNHA-SA" | sensor=="TNHA-SC")

tnha_wind = wind %>% filter(station == "TNHA") %>% filter(sensor=="TNHA-SA" | sensor=="TNHA-SC")

tnha_solar = solar %>% filter(station == "TNHA") %>% filter(sensor=="TNHA-SA" | sensor=="TNHA-SC")
#tnha_solar$sensor[tnha_solar$sensor=="21398620"] = "TNHA-SA"

tnha = full_join(tnha_ground, tnha_air, by=c("date","station","sensor")) %>% select(c(2:4,6,9))
colnames(tnha)[4:5] = c("groundtemp", "airtemp")
tnha = full_join(tnha, tnha_wind, by=c("date","station","sensor")) %>% select(c(1:5,8))
colnames(tnha)[6] = "windspeed"
tnha = full_join(tnha, tnha_solar, by=c("date","station","sensor")) %>% select(c(1:6,9))
colnames(tnha)[7] = "solar"
```



```{r}
tnha = na.omit(tnha)
```




## PCA
```{r}
tnha_pca <- prcomp(tnha[,4:7], scale.=TRUE)

#take out variables
sd <- tnha_pca$sdev
loadings <- tnha_pca$rotation
rownames(loadings) <- colnames(sd)
scores <- tnha_pca$x

var <- sd^2
varPercent <- var/sum(var) * 100
dev.new() #not sure what this does
```

### Show Results
```{r}
barplot(varPercent, xlab="PC", ylab="Percent Variance", names.arg=1:length(varPercent), las=1, ylim=c(0, max(varPercent)), col="gray")
abline(h=1/ncol(sd)*100, col="red")
```

### Loadings
```{r}
loadings
sqrt(1/ncol(sd)) # cutoff for "important" loadings
```


### Bi-plots
```{r}
dev.new(height=7, width=7)
TNHA_SA <- tnha$sensor == "TNHA-SA"
TNHA_SC <- tnha$sensor == "TNHA-SC"
 
plot(scores[, 1], scores[, 2], xlab="PCA 1", ylab="PCA 2", type="n", asp=1, las=1)

scaling <- 4
textNudge <- 1.05
 
points(scores[TNHA_SA, 1], scores[TNHA_SA, 2], pch=16, cex=0.7, col="lightblue")
 
points(scores[TNHA_SC, 1], scores[TNHA_SC, 2], pch=16, cex=0.7, col="pink")
 

arrows(0, 0, loadings[, 1]* scaling, loadings[, 2]* scaling, length=0.1, angle=20, col="darkred")
 
text(loadings[, 1]*scaling*textNudge, loadings[, 2]*scaling*textNudge, rownames(loadings), col="darkred", cex=0.7)

 

text(1, 2, "TNHA-SA \n(south)", col="darkblue")
text(-4, 2, "TNHA-SC \n(north)", col="darkred")
```

# PCA - SSMH

Average
```{r}

```


North vs South
SSMH-D vs SSMH-SB
```{r}

```


# PCA - BEO

JUST ONE SENSOR
BEO-6
```{r}

```




