---
title: "Freeze_Thaw_Gap_Fill"
author: "Mirella Shaban"
date: "2024-11-08"
output: html_document
---

based off of the original freeze_thaw_dates.rmd that used raw data

This script gives dates of above or below 0.5°C for stations to determine dates of freeze or thaw



Load in data and aggregate by station
------
Load packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape) # basically like text to columns
library(naniar)
library(RColorBrewer)
library(caret)
```

Load gap filled daily average data from GIT
```{r}
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/Meteorological_Seasons_Data/"

gapfilldata = "all_sites_daily_gap_filled.csv"

allvargapfill =read.csv(paste0(filepath, gapfilldata))
```
If no wifi, load through downloads folder:
```{r}
#allvargapfill <- all_sites_daily_gap_filled
```


Separate each variable into its own dataframe
```{r}
grndtmp = allvargapfill %>% subset() %>% select(day, fullname, site, station, grounddepth, groundtemp) %>% filter( !grounddepth == "-888.8") 

grndtmp$grounddepth <- as.factor(as.character(grndtmp$grounddepth))
grndtmp$grounddepth <- revalue(grndtmp$grounddepth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", "10" = "30cm", "11" = "40cm", "12" = "50cm", "13" = "55cm", "14" = "65cm", "15" = "75cm", "16" = "85cm", "17" = "90cm"))

vwc = allvargapfill %>% subset() %>% select(day, fullname, site, station, grounddepth, vwc) %>% filter( !vwc == "-888.8")

```


AVERAGE BY STATION (can compare sensors at the same site, such as N & S) by day
```{r}
grndtmp <- grndtmp %>% filter(!day <= "2022-06-19")

averaged_by_station <- grndtmp %>% filter(fullname == "TNHA-BASE") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station$fullname <- c("TNHA-BASE")

averaged_by_station1 <- grndtmp %>% filter(fullname == "TNHA-SA") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station1$fullname <- c("TNHA-SA")

averaged_by_station2 <- grndtmp %>% filter(fullname == "TNHA-SB") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station2$fullname <- c("TNHA-SB")

averaged_by_station3 <- grndtmp %>% filter(fullname == "TNHA-SC") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station3$fullname <- c("TNHA-SC")


averaged_by_station4 <- grndtmp %>% filter(fullname == "TNHA-SD") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station4$fullname <- c("TNHA-SD")

averaged_by_station5 <- grndtmp %>% filter(fullname == "SSMH-BASE") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station5$fullname <- c("SSMH-BASE")

averaged_by_station6 <- grndtmp %>% filter(fullname == "SSMH-SA") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station6$fullname <- c("SSMH-SA")

averaged_by_station7 <- grndtmp %>% filter(fullname == "SSMH-SB") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station7$fullname <- c("SSMH-SB")


averaged_by_station8 <- grndtmp %>% filter(fullname == "SSMH-SD") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station8$fullname <- c("SSMH-SD")


averaged_by_station9 <- grndtmp %>% filter(fullname == "SSMH-SE") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station9$fullname <- c("SSMH-SE")

averaged_by_station10 <- grndtmp %>% filter(fullname == "SSMH-SF") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station10$fullname <- c("SSMH-SF")

averaged_by_station11 <- grndtmp %>% filter(fullname == "SSMH-SG") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station11$fullname <- c("SSMH-SG")

averaged_by_station12 <- grndtmp %>% filter(fullname == "SSMH-SH") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station12$fullname <- c("SSMH-SH")

averaged_by_station13 <- grndtmp %>% filter(fullname == "SSMH-SI") %>% aggregate(groundtemp~day+grounddepth, mean)
averaged_by_station13$fullname <- c("SSMH-SI")

averaged_by_station14 <- grndtmp %>% filter(fullname == "BEO-B05" | fullname == "BEO-B06" | fullname == "BEO-B07" | fullname == "BEO-B08") %>% aggregate(groundtemp~day+grounddepth, mean)
 averaged_by_station14$fullname <- c("BEO")




Daily_grndtmp_avgs<- rbind(averaged_by_station, averaged_by_station1, averaged_by_station2, averaged_by_station3,averaged_by_station4,averaged_by_station5,averaged_by_station6,averaged_by_station7,averaged_by_station9,averaged_by_station8,averaged_by_station11,averaged_by_station13,averaged_by_station14)

Daily_grndtmp_avgs$day <- as.Date(Daily_grndtmp_avgs$day, format= "%Y-%m-%d",tz="UTC")
```





Freeze  dates
```{r}

#Constrain by season

freeze_dates <- Daily_grndtmp_avgs %>% filter(fullname == "SSMH-BASE" & groundtemp < 1 & grounddepth =="90cm" & day >= "2022-06-01" & day <= "2022-10-01")
list(freeze_dates)

#next, compare to thawed dates for summer to see. the table generated from this may end up being too many inputs in the cells
```

Thaw dates
```{r}
DF_1 <- data.frame(list(freeze_dates$day))
list(DF_1)

thaw_dates <- Daily_grndtmp_avgs %>% filter( fullname == "SSMH-SA" & groundtemp > -1 & grounddepth == "90cm" & day >= "2022-06-19" & day <= "2022-10-01") 

list(thaw_dates)

```



FIGURES OF ALL STATIONS DURING YEAR = X *VARIATION*
```{r fig1, fig.height = 30, fig.width = 26}
VARIATION <- thaw_dates %>% filter( groundtemp > -1 & groundtemp < 1)

ggplot(VARIATION, aes(x= day, y=groundtemp)) +
  geom_point() +
  theme_bw() +
  labs(title= "SUMMER 2024 THAW VARIATION", x = "Date", y="Ground Temperature (°C)")  +
  facet_wrap(grounddepth~fullname,ncol=5,drop=FALSE)
  
```

```{r fig2, fig.height = 30, fig.width = 20}
VARIATION <- SUMM24_ALLDEPTHS %>% filter(fullname == "TNHA-BASE", value > -1 & value < 1)

ggplot(VARIATION, aes(x= date, y=value)) +
  geom_point() +
  theme_bw() +
  labs(title= "SUMMER 2024 THAW VARIATION - TNHA-BASE", x = "Date", y="Ground Temperature (°C)")  +
  facet_wrap(depth~fullname,nrow=5,drop=FALSE)
  
```
