---
title: "Gap_Fill_Transpose_and_Join"
author: "Mirella Shaban"
date: "2026-01-08"
output: html_document
---

This code combines 2022-2025 data for three sites (TNHA, SSMH, BEO) and then gathers daily averages for each variable, maps depths for VWC to Grounddepth, and joins all variables (averaged daily) into one master dataframe

Switch out the file names in the first chunk, and in the subsequent chunks change the CSV name in the write.csv() line 
```{r}
library(dplyr)
library(tidyr)
library(splitstackshape)

path <- "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/"

#switch these out as needed
moistfile <- "SSMH_moisture_Data_CLEAN_OLDplusNEW_merged_2022-06-08_to_2025-01-18.csv"
solarfile <- "TNHA_solar_Data_merged_2022-05-17_to_2025-09-04_KNN_SAVGOL.csv"
windspfile <- "TNHA_wind_sp_Data_merged_2022-05-17_to_2025-09-04_CLEANED_KNN_SAVGOL.csv"
winddirfile <- "TNHA_wind_dir_Data_merged_2022-05-17_to_2025-09-04_CLEANED_KNN_SAVGOL.csv"

# filenames
  # TNHA_moisture_Data_merged_2022-05-17_to_2025-09-04_CLEANED_KNN_SAVGOL.csv
  # SSMH_moisture_Data_CLEAN_OLDplusNEW_merged_2022-06-08_to_2025-01-18.csv
  # BEO_moisture_Data_CLEANED_OLDplusNEW_merged_2022-06-13_to_2025-07-08_KNN_KNN_SAVGOL.csv
  
  # BEO_solar_Data_CLEANED_OLDplusNEW_merged_2022-06-13_to_2025-09-20_KNN_KNN_SAVGOL.csv
  # TNHA_solar_Data_merged_2022-05-17_to_2025-09-04_KNN_SAVGOL.csv 
  # SSMH_solar_Data_FINAL_CLEAN_KNN_Nov2025.csv

  # BEO_temp_Data_CLEANED_OLDplusNEW_merged_2022-06-13_to_2025-09-20_KNN_KNN_SAVGOL.csv
  # TNHA_temp_Data_CLEAN_OLDplusNEW_merged_from_OLDstart_to_2025-09-04_KNN_SAVGOL.csv  
  # SSMH_temp_Data_CLEAN_OLDplusNEW_merged_2022-06-08_to_2025-10-16.csv  
  
  # BEO_wind_dir_Data_CLEANED_OLDplusNEW_merged_2022-06-13_to_2025-09-20_KNN_KNN_SAVGOL.csv  
  # SSMH_wind_dir_Data_CLEAN_OLDplusNEW_merged_2022-06-08_to_2025-10-16.csv  
  # TNHA_wind_dir_Data_merged_2022-05-17_to_2025-09-04_CLEANED_KNN_SAVGOL.csv  
  
  # BEO_wind_sp_Data_CLEANED_OLDplusNEW_merged_2022-06-13_to_2025-09-20_KNN_KNN_SAVGOL.csv  
  # SSMH_wind_sp_Data_CLEAN_OLDplusNEW_merged_2022-06-08_to_2025-10-16.csv  
  # TNHA_wind_sp_Data_merged_2022-05-17_to_2025-09-04_CLEANED_KNN_SAVGOL.csv



moist <- read.csv(paste0(path, moistfile))
windsp <- read.csv(paste0(path, windspfile))
solar <- read.csv(paste0(path, solarfile))
winddir <- read.csv(paste0(path, winddirfile))

```


VWC
```{r}
# Your initial reshaping
grndtmp <- gather(moist, variable, response, 2:ncol(moist)) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

grndtmp <- cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(grndtmp)[2:4] <- c("value", "sensor", "depth")
grndtmp$sensor <- gsub('X', '', grndtmp$sensor)
grndtmp$Date <- as.POSIXct(grndtmp$Date,format="%Y-%m-%d %H:%M:%S",tz="UTC")

# Define sensor mappings
vwc_sensors <- c('21393048'='BEO-B06','21398591'='BEO-B05','21393045'='BEO-B07','21398588'='SSMH-BASE','21398599'='SSMH-SA','21393044'='SSMH-SB','21393049'='SSMH-SD','21398594'='SSMH-SE','21393043'='SSMH-SF','21398586'='SSMH-SG','21398580'='SSMH-SH','21398581'='SSMH-SI','21206939'='TNHA-BASE','21398593'='TNHA-SA','21398587'='TNHA-SB','21393047'='TNHA-SC','21398601'='TNHA-SD','21398577'='TNHA-SE','21398576'='BEO-B08','21393046'='TNHA-SC','21166008'='SSMH-SC')


# Rename depth values
grndtmp$depth <- as.character(grndtmp$depth)
grndtmp$depth <- revalue(grndtmp$depth, c("7" = "3.5cm", "8" = "10cm", "9" = "20cm", 
                                           "10" = "30cm", "11" = "40cm", "12" = "50cm", 
                                           "13" = "55cm", "14" = "65cm", "15" = "75cm", 
                                           "16" = "85cm", "17" = "90cm"))

# Depth mapping
depth_mapping <- c("7"="3.5cm","8"="10cm","9"="20cm","10"="30cm","11"="40cm",
                   "12"="50cm","13"="55cm","14"="65cm","15"="75cm","16"="85cm","17"="90cm")

# VWC (with depth)
  df_vwc <- grndtmp %>%
    filter(sensor %in% names(vwc_sensors)) %>%
    mutate(
      sensor_name = vwc_sensors[sensor],
      depth = revalue(as.character(depth), depth_mapping)
    ) %>%
    filter(sensor_name != "REMOVE") %>%
    group_by(Date, sensor_name, depth) %>%
    summarise(vwc = mean(value, na.rm = TRUE), .groups = "drop")
  

#write.csv(df_vwc, "~/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/SSMH_VWC.csv")

```



SOLAR
```{r}
grndtmp <- gather(solar, variable, response, 2:ncol(solar)) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

grndtmp <- cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(grndtmp)[2:4] <- c("value", "sensor", "depth")
grndtmp$sensor <- gsub('X', '', grndtmp$sensor)
grndtmp$Date <- as.POSIXct(grndtmp$Date,format="%Y-%m-%d %H:%M:%S",tz="UTC")

solar_sensors <- c('21390415'='BEO-BASE','21390413'='SSMH-BASE','21398622'='SSMH-SA','21362319'='SSMH-SB','21398619'='SSMH-SI','21362318'='TNHA-SB','21176526'='TNHA-BASE','21398620'='TNHA-SA','21398616'='TNHA-SB','21362317'='TNHA-SC','21362315'='TNHA-SD')

grndtmp <- grndtmp %>% filter(!value < 0)

# SOLAR
  df_solar <- grndtmp %>%
    filter(sensor %in% names(solar_sensors)) %>%
    mutate(sensor_name = solar_sensors[sensor]) %>%
    filter(sensor_name != "REMOVE") %>%
    group_by(Date, sensor_name) %>%
    summarise(solar = mean(value, na.rm = TRUE), .groups = "drop")
#write.csv(df_solar, "~/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/TNHA_Solar.csv")
```



WINDSP
```{r}
grndtmp <- gather(windsp, variable, response, 2:ncol(windsp)) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

grndtmp <- cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(grndtmp)[2:4] <- c("value", "sensor", "depth")
grndtmp$sensor <- gsub('X', '', grndtmp$sensor)
grndtmp$Date <- as.POSIXct(grndtmp$Date,format="%Y-%m-%d %H:%M:%S",tz="UTC")

wind_sensors <- c('21350915'='BEO-BASE','21350901'='SSMH-BASE','21350910'='SSMH-BASE','21206911'='SSMH-SA','21206912'='SSMH-SB','21398711'='SSMH-SI','21176861'='TNHA-BASE','21398709'='TNHA-SA','21181033'='TNHA-SB','21398715'='TNHA-SB','21206909'='TNHA-SC','21398712'='TNHA-SD','21350855'='BEO-BASE','21350820'='SSMH-BASE','21167037'='TNHA-BASE','21398716'='TNHA-SD')

winddir_ids <- c('21350855','21350894','21398590','21398729','21398724','21398660','21398719','21350820','21206911','21206912','21398711','21181033','21167037','21398709','21398715','21206909','21398712','21398716')

windspeed_ids <- c('21350915','21350894','21398590','21398729','21398724','21398660','21398719','21350901','21350910','21206911','21206912','21398711','21176861','21398709','21181033','21398715','21206909','21398712','21398716')

  # WIND SPEED
  df_windspeed <- grndtmp %>%
    filter(sensor %in% windspeed_ids & sensor %in% names(wind_sensors)) %>%
    mutate(sensor_name = wind_sensors[sensor]) %>%
    filter(sensor_name != "REMOVE") %>%
    group_by(Date, sensor_name) %>%
    summarise(wind_speed = mean(value, na.rm = TRUE), .groups = "drop")
#write.csv(df_windspeed, "~/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/TNHA_Windsp.csv")
```


WINDDIR
```{r}
grndtmp <- gather(winddir, variable, response, 2:ncol(winddir)) %>%
  filter(!is.na(response)) %>%
  mutate(variable = fct_inorder(variable))

grndtmp <- cSplit(grndtmp, 'variable', sep=".", direction = "wide") 
colnames(grndtmp)[2:4] <- c("value", "sensor", "depth")
grndtmp$sensor <- gsub('X', '', grndtmp$sensor)
grndtmp$Date <- as.POSIXct(grndtmp$Date,format="%Y-%m-%d %H:%M:%S",tz="UTC")

wind_sensors <- c('21350915'='BEO-BASE','21350901'='SSMH-BASE','21350910'='SSMH-BASE','21206911'='SSMH-SA','21206912'='SSMH-SB','21398711'='SSMH-SI','21176861'='TNHA-BASE','21398709'='TNHA-SA','21181033'='TNHA-SB','21398715'='TNHA-SB','21206909'='TNHA-SC','21398712'='TNHA-SD','21350855'='BEO-BASE','21350820'='SSMH-BASE','21167037'='TNHA-BASE','21398716'='TNHA-SD')

winddir_ids <- c('21350855','21350894','21398590','21398729','21398724','21398660','21398719','21350820','21206911','21206912','21398711','21181033','21167037','21398709','21398715','21206909','21398712','21398716')

windspeed_ids <- c('21350915','21350894','21398590','21398729','21398724','21398660','21398719','21350901','21350910','21206911','21206912','21398711','21176861','21398709','21181033','21398715','21206909','21398712','21398716')
  
  # WIND DIRECTION
  df_winddir <- grndtmp %>%
    filter(sensor %in% winddir_ids & sensor %in% names(wind_sensors)) %>%
    mutate(sensor_name = wind_sensors[sensor]) %>%
    filter(sensor_name != "REMOVE") %>%
    group_by(Date, sensor_name) %>%
    summarise(wind_dir = mean(value, na.rm = TRUE), .groups = "drop")

#write.csv(df_winddir, "~/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/TNHA_Winddir.csv")

```

Load all sites variable dataframes into R
```{r}
#Combining files of same type into 1 large file per variable

path_1 <- "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/"


a1 <- "BEO_Air.csv"
d1 <- "BEO_Ground.csv"
g1 <- "BEO_Solar.csv"
i1 <- "BEO_VWC.csv"
l1 <- "BEO_Winddir.csv"
o1 <- "BEO_Windsp.csv"
b1 <- "SSMH_Air_1.csv"
e1 <- "SSMH_Ground_1.csv"
x1 <- "SSMH_solar.csv"
j1 <- "SSMH_VWC.csv"
m1 <- "SSMH_Winddir.csv"
p1 <- "SSMH_Windsp.csv"
c1 <- "TNHA_Air.csv"
f1 <- "TNHA_Ground.csv"
h1 <- "TNHA_Solar.csv"
k1 <- "TNHA_VWC.csv"
n1 <- "TNHA_Winddir.csv"
q1 <- "TNHA_Windsp.csv"

a <- read.csv(paste0(path_1, a1))
b <- read.csv(paste0(path_1, b1))
c <- read.csv(paste0(path_1, c1))
d <- read.csv(paste0(path_1, d1))
e <- read.csv(paste0(path_1, e1))
f <- read.csv(paste0(path_1, f1))
g <- read.csv(paste0(path_1, g1))
h <- read.csv(paste0(path_1, h1))
i <- read.csv(paste0(path_1, i1))
j <- read.csv(paste0(path_1, j1))
k <- read.csv(paste0(path_1, k1))
l <- read.csv(paste0(path_1, l1))
m <- read.csv(paste0(path_1, m1))
n <- read.csv(paste0(path_1, n1))
o <- read.csv(paste0(path_1, o1))
p <- read.csv(paste0(path_1, p1))
q <- read.csv(paste0(path_1, q1))
x <- read.csv(paste0(path_1 ,x1))

gap_fill_air <- bind_rows(a,b,c)
gap_fill_ground <- bind_rows(d,e,f)
gap_fill_vwc <- bind_rows(i,j,k)
gap_fill_solar <- bind_rows(g,h,x)
gap_fill_wind_dir <- bind_rows(l,m,n)
gap_fill_wind_sp <- bind_rows(o,p,q)

#write.csv(gap_fill_solar, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/solar_22_25.csv")
#write.csv(gap_fill_air, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/air_22_25.csv")
#write.csv(gap_fill_ground, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/ground_22_25.csv")
#write.csv(gap_fill_vwc, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/vwc_22_25.csv")
#write.csv(gap_fill_wind_dir, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/winddir_22_25.csv")
#write.csv(gap_fill_wind_sp, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/windsp_22_25.csv")

```

load csv's created earlier in script (jump straight here if you want to just run a daily averaging and combination of all the scripts)
```{r}
ground <- read.csv("/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/ground_22_25_1.csv")
air <- read.csv("/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/air_22_25.csv")
vwc <- read.csv("/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/vwc_22_25.csv")
solar <- read.csv("/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/solar_22_25.csv")
windsp <- read.csv("/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/windsp_22_25.csv")
winddir <- read.csv("/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/winddir_22_25.csv")
```



Daily averages 
```{r}
#Daily combination?? taken from daily_Avg_All_Seasons_2025.... script

#need to map the BEO-BASE to BEO-B05, B06, B07, B08...

gt_daily <- ground %>%
  mutate(Date = as.Date(Date)) %>%
  aggregate( ground_temp~ Date + sensor_name + depth, na.rm = TRUE, mean)  ## THIS WORKS -- ADJUST THE REST tO LOOK LIKE THIS

air_daily <- air %>%   mutate(Date = as.Date(Date)) %>%
  aggregate( air_temp~ Date + sensor_name , na.rm = TRUE, mean)

solar_daily <- solar %>%   mutate(Date = as.Date(Date)) %>%
  aggregate( solar~ Date + sensor_name, na.rm = TRUE, mean)

vwc_daily <- vwc %>%   mutate(Date = as.Date(Date)) %>%
  aggregate( vwc~ Date + sensor_name + depth, na.rm = TRUE, mean)

ws_daily <- windsp %>%   mutate(Date = as.Date(Date)) %>%
  aggregate( wind_speed~ Date + sensor_name , na.rm = TRUE, mean)


wd_daily <- winddir %>%   mutate(Date = as.Date(Date)) %>%
  aggregate( wind_dir~ Date + sensor_name , na.rm = TRUE, mean)



#DAILY AVERAGED FOR EACH VARIABLE

# write.csv(solar_daily, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/solar_22_25_DAILY.csv")
# write.csv(air_daily, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/air_22_25_DAILY.csv")
# write.csv(gt_daily, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/ground_22_25_DAILY.csv")
# write.csv(vwc_daily, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/vwc_22_25_DAILY.csv")
# write.csv(wd_daily, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/winddir_22_25_DAILY.csv")
# write.csv(ws_daily, "/Users/shabangin/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/windsp_22_25_DAILY.csv")
```

Recode VWC depths, match to ground depths, join all variables together into one dataframe
Newest workflow
```{r}
# Step 1: Recode vwcdepth in vwc_daily
vwc_daily$vwcdepth <- recode(as.character(vwc_daily$depth), 
                              "1" = "0-15cm", 
                              "2" = "15-30cm", 
                              "3" = "30-45cm",
                              "4" = "45-60cm", 
                              "5" = "60-75cm", 
                              "6" = "75-90cm")

# Step 2: Apply vwcdepth mapping to gt_daily
colnames(gt_daily)[3] = "grounddepth"
gt_daily <- gt_daily %>% 
  mutate(vwcdepth = case_when(
    grounddepth == "3.5cm" ~ "0-15cm",
    grounddepth == "10cm" ~ "0-15cm",
    grounddepth == "20cm" ~ "15-30cm",
    grounddepth == "30cm" ~ "30-45cm",
    grounddepth == "40cm" ~ "30-45cm",
    grounddepth == "50cm" ~ "45-60cm",
    grounddepth == "55cm" ~ "45-60cm",
    grounddepth == "65cm" ~ "60-75cm",
    grounddepth == "75cm" ~ "75-90cm",
    grounddepth == "85cm" ~ "75-90cm",
    grounddepth == "90cm" ~ "75-90cm"
  ))

# Step 3: Expand BEO-BASE for air, solar, wind ONLY
beo_ground_sensors <- c('BEO-B05', 'BEO-B06', 'BEO-B07', 'BEO-B08')

expand_beo_base <- function(df) {
  beo_base_rows <- df %>% filter(sensor_name == "BEO-BASE")
  non_beo_rows <- df %>% filter(sensor_name != "BEO-BASE")
  
  beo_expanded <- beo_base_rows %>%
    crossing(new_sensor = beo_ground_sensors) %>%
    mutate(sensor_name = new_sensor) %>%
    select(-new_sensor)
  
  bind_rows(non_beo_rows, beo_expanded)
}

# Apply ONLY to datasets with BEO-BASE (NOT gt_daily or vwc_daily)
air_daily <- air_daily %>% expand_beo_base()
ws_daily <- ws_daily %>% expand_beo_base()
wd_daily <- wd_daily %>% expand_beo_base()
solar_daily <- solar_daily %>% expand_beo_base()

# Step 4: Verify sensor names
gt_daily %>% filter(grepl("BEO", sensor_name)) %>% distinct(sensor_name)
vwc_daily %>% filter(grepl("BEO", sensor_name)) %>% distinct(sensor_name)
air_daily %>% filter(grepl("BEO", sensor_name)) %>% distinct(sensor_name)

# # Step 5: Join
# join <- full_join(gt_daily, vwc_daily, by = c("Date", "sensor_name", "vwcdepth"), relationship = "many-to-many") %>%
#   full_join(air_daily, by = c("Date", "sensor_name"), relationship = "many-to-many") %>%
#   full_join(solar_daily, by = c("Date", "sensor_name"), relationship = "many-to-many") %>%
#   full_join(ws_daily, by = c("Date", "sensor_name"), relationship = "many-to-many") %>%
#   full_join(wd_daily, by = c("Date", "sensor_name"), relationship = "many-to-many") %>%
#   select(Date, sensor_name, grounddepth, vwcdepth, groundtemp, vwc, airtemp, solar, windspeed, winddirection) %>%
#   distinct() 


# Try without the intermediate unique() calls
join_test <- full_join(gt_daily, vwc_daily, by = c("Date", "sensor_name", "vwcdepth"), relationship = "many-to-many") %>%
  full_join(air_daily, by = c("Date", "sensor_name"), relationship = "many-to-many") %>%
  full_join(solar_daily, by = c("Date", "sensor_name"), relationship = "many-to-many") %>%
  full_join(ws_daily, by = c("Date", "sensor_name"), relationship = "many-to-many") %>%
  full_join(wd_daily, by = c("Date", "sensor_name"), relationship = "many-to-many") %>%
  select(Date, sensor_name, grounddepth, vwcdepth, ground_temp, vwc, air_temp, solar, wind_speed, wind_dir) %>%
  distinct()  # Use distinct() instead of unique()

#write.csv(join_test, "~/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/all_var_combined_22_25_BEO_COMP_updated.csv")

# Check results
sum(!is.na(join_test$airtemp))
head(join_test)


```

