---
title: "VWC_Water_Depth"
author: "Mirella Shaban"
date: "2026-03-13"
output: html_document
---


Load in data and aggregate by station
------
Load packages
```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(lubridate)
library(tidyr)
library(gridExtra)
library(forcats)
library(splitstackshape) # basically like text to columns
library(naniar)
library(RColorBrewer)
library(caret)
```

Load gap filled daily average data from DESKTOP (VERSIONING GIT - NEED TO UPDATE (02/25/26))
```{r}
filepath = "~/Desktop/UVA/RESEARCH/Barrow/GAP_FILLED_CSV_DOWNLOADS/Sep/"

gapfilldata = "all_var_combined_22_25_BEO_COMP_updated.csv" #is there no SSMH-BASE?? --> FOR SOME REASON IT DIDNT LOAD IN DURING GAP_FILL_TRANSPOSE_AND_JOIN

allvargapfill =read.csv(paste0(filepath, gapfilldata))

allvargapfill$vwc <- round(allvargapfill$vwc, 2)

allvargapfill$Date <- as.Date(allvargapfill$Date, format= "%Y-%m-%d",tz="UTC")
```


Separate each variable into its own dataframe
```{r}
vwc = allvargapfill %>% subset() %>% select(Date, sensor_name, vwcdepth, vwc) %>% filter( !vwc == "-888.8") 

```

This is a much more concise way of doing what the chunk below is doing......... sigh.
```{r}
averaged_by_station <- vwc %>% aggregate(vwc~sensor_name+Date+vwcdepth, mean)
```

```{r}

vwc <- averaged_by_station %>%
filter(sensor_name =="BEO-B05" | sensor_name =="BEO-B06" | sensor_name =="BEO-B07" | sensor_name =="BEO-B08" |sensor_name =="BEO1" |sensor_name =="BEO2" |sensor_name == "TNHA-BASE" | sensor_name == "TNHA-SA" |sensor_name =="TNHA-SB"|sensor_name =="TNHA-SC" |sensor_name == "TNHA-SD" |sensor_name == "SSMH-BASE" |sensor_name =="SSMH-SA"|sensor_name =="SSMH-SB" |sensor_name =="SSMH-SD")


#RENAME ALL STATIONS FOR PLOTTING
# Create a named vector for renaming stations
station_names <- c(
  "BEO-B05" = "Tundra-1",
  "BEO-B06" = "Tundra-2",
  "BEO-B07" = "Tundra-3",
  "BEO-B08" = "Tundra-4",
  "BEO1" = "Tundra-OMIT1",
  "BEO2" = "Tundra-OMIT2",
  "TNHA-BASE" = "RESIDENTIAL-S1",
  "TNHA-SA" = "RESIDENTIAL-S2",
  "TNHA-SB" = "RESIDENTIAL-W",
  "TNHA-SC" = "RESIDENTIAL-N",
  "TNHA-SD" = "RESIDENTIAL-E",
  "SSMH-BASE" = "HOSPITAL-S1",
  "SSMH-SA" = "HOSPITAL-S2",
  "SSMH-SB" = "HOSPITAL-N",
  "SSMH-SD" = "HOSPITAL-E"
)

# Rename stations for display
vwc$sensor_name <- recode(vwc$sensor_name, !!!station_names)

vwc <- vwc %>% mutate(across(c("vwc"), round, 2))
```


```{r}
#date constraint
season <- vwc %>% filter(Date >= "2022-06-09" & Date <= "2022-09-01")



# WATER DEPTH
#take the average of the water across all depths at each station, get the average for the season. Aggregate by sensor
waterdepth <- season %>%
  group_by(sensor_name) %>%
  aggregate(vwc~sensor_name, mean)

waterdepth <- waterdepth %>% mutate(across(c("vwc"), round, 2))
waterdepth$vwc_water <- waterdepth$vwc * 90

# SET THE FACTOR LEVELS TO CONTROL LEGEND ORDER
waterdepth$sensor_name <- factor(waterdepth$sensor_name, levels = c(
  "Tundra-1",
  "Tundra-2",
  "RESIDENTIAL-S1",
  "RESIDENTIAL-S2",
  "RESIDENTIAL-W",
  "RESIDENTIAL-N",
  "RESIDENTIAL-E",
  "HOSPITAL-N",
  "HOSPITAL-E",
  "HOSPITAL-S1",
  "HOSPITAL-S2"
))

water_plot <- waterdepth %>% ggplot() +
  geom_point(aes(x= sensor_name, y= vwc_water, color = "vwc_depth")) +
  theme_bw()+
  theme(legend.title = element_text(size = 8),legend.text = element_text(size = 10)) +
  labs(title = "Water Depths, averaged across depths + season ", y="Water Depth", x="Station",color="Depth") +
   theme(axis.text.y =element_text(face="bold",size = 8) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 8), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10))

#Next : calculate the porosity of the soil??

```
Water Depths across the season (daily averages facet wrapped )
```{r fig, fig.height = 10, fig.width = 15}
# WATER DEPTH
#take the average of the water across all depths at each station, get the average for each day Aggregate by sensor
waterdepth_day <- season %>%
  group_by(sensor_name) %>%
  aggregate(vwc~sensor_name + Date, mean)

#xyz$day <- as.Date(xyz$day, format = "%Y-%m-%d")

waterdepth_day %>%
  ggplot(aes(x=Date, y=vwc, color = sensor_name)) +
# geom_point() +
  geom_line(size =1) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw()+ 
  theme(legend.title = element_text(size = 15),legend.text = element_text(size = 10)) +
  labs(title = "Water Depth per station, Summer 2022", y="Depth (cm)", x="Week", color="Station") +
  scale_x_date(date_breaks = "1 weeks",
               date_labels = "%m/%d") +
 # scale_y_continuous(breaks=c(-5,-2.5,-1, -0.5, 0,0.5, 1, 2.5), limits=c(-5, 5)) +
  theme(axis.text.y =element_text(face="bold",size = 20) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 20), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 30),
        axis.title.x = element_text(size = 25),
        axis.title.y = element_text(size = 25))
```
Now, facet wrapped
```{r fig, fig.height = 20, fig.width = 10}
waterdepth_day %>%
  ggplot(aes(x=Date, y=vwc, color = sensor_name)) +
  geom_line(size =1) +
  facet_wrap(sensor_name ~ ., scales = "free_y", ncol=1) +
  guides(color = guide_legend(override.aes = list(size=10)))+
  theme_bw()+ 
  theme(legend.title = element_text(size = 15),legend.text = element_text(size = 10)) +
  labs(title = "Water Depth per station, Summer 2022", y="Depth (cm)", x="Week", color="Station") +
  scale_x_date(date_breaks = "1 weeks",
               date_labels = "%m/%d") +
 # scale_y_continuous(breaks=c(-5,-2.5,-1, -0.5, 0,0.5, 1, 2.5), limits=c(-5, 5)) +
  theme(axis.text.y =element_text(face="bold",size = 10) ,
        axis.text.x=element_text(angle = 45, hjust = 1.0, face = "bold", size = 10), 
        axis.line = element_line(colour = "black", size =0.5, linetype = "solid")) +
  theme(plot.title = element_text(hjust = 0.5, size = 20),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))
```
Now, how do i do this to where i have a column for each year?
```{r fig, fig.height = 18, fig.width = 22}
season1 <- vwc %>% filter(Date >= "2022-06-09" & Date <= "2022-09-01" |Date >= "2023-06-09" & Date <= "2023-09-01" | Date >= "2024-06-09" & Date <= "2024-09-01" |Date >= "2025-06-09" & Date <= "2025-09-01")

waterdepth_day1 <- season1 %>%
  group_by(sensor_name) %>%
  aggregate(vwc~sensor_name + Date, mean)


waterdepth_day1 %>%
  mutate(
    year = lubridate::year(Date),
    # Create a common date for comparison (using 2000 as base year)
    comparison_date = as.Date(paste0("2022-", 
                                      lubridate::month(Date), "-", 
                                      lubridate::day(Date)))
  ) %>%
  ggplot(aes(x = comparison_date, y = vwc, color = sensor_name)) +
  geom_line(size = 1) +
  facet_grid(sensor_name ~ year, scales = "free_y") +
  guides(color = guide_legend(override.aes = list(size = 10))) +
  theme_bw() + 
  theme(
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 10),
    axis.text.y = element_text(face = "bold", size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1.0, face = "bold", size = 10), 
    axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"),
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    strip.text = element_text(size = 12, face = "bold"),  # Facet label styling
    panel.spacing = unit(1, "lines")  # Space between panels
  ) +
  labs(
    title = "Water Depth per Station by Year", 
    y = "Depth (cm)", 
    x = "Date (Month/Day)", 
    color = "Station"
  ) +
  scale_x_date(
    date_breaks = "1 week",
    date_labels = "%m/%d"
  )
```

