---
title: "Generalized Linear Mixed Models"
author: "Elizabeth Van Metre"
date: "2024-10-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Generalized Linear Mixed Models

Created: October 1, 2024

**Fixed Effects**: variables that we expect to have an effect on the
response variable (i.e. independent or explanatory variables)

**Random Effects**: [grouping factors]{.underline} that may cause variation
that we want to control; ALWAYS CATEGORICAL

The data for our random effects are just a sample of all the
possibilities - our data are limited by time and resources, and
therefore are only a representation of part of the population.

```{r}
library(lme4)
```

## Load Data

```{r}
filepath <- "/Users/emvanmetre/Desktop/Arctic/Gap_Filled_Data/Transposed/"
df <- read.csv(paste0(filepath, "all_sites_hourly_gap_filled.csv"))
df$copy <- df$fullname
df <- cSplit(df, 'copy', sep="-", direction = "wide")
colnames(df)[12:13] <- c("site", "station")
```


We need to look at the distribution of our data.

First we fit a linear mixed model.

```{r}
mixed.lmer <- lmer(groundtemp ~ airtemp + (1|site), data = df)
summary(mixed.lmer)
```

Check Assumptions
```{r}
plot(mixed.lmer)
```


Sources:
https://ourcodingclub.github.io/tutorials/mixed-models/#what




START
1. North vs South at one site - try to figure out that there is a difference
2. South vs South at two different sites
Eventually... do we see a difference between years in our North vs South GLMMs


Seasonal
Summer 2022, Fall 2022, Spring 2023 - separate GLMMs
Site by Site
Also aspect by aspect (n vs s)
Depth: start with 10cm; 30cm; 55cm; 65cm; 90cm

change this order:
- aspects
- locations
- depth
- season