---
title: "Generalized Linear Mixed Models"
author: "Elizabeth Van Metre"
date: "2024-10-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Generalized Linear Mixed Models

**Created: October 1, 2024**

[Changes]{.underline}

-   10/1/24: added lme4 package and began analysis of LMM
-   10/30/24: updated with new averages from gap-filled data
-   1/21/24: added glmm code

## Linear Mixed Models

Source:
<https://stats.oarc.ucla.edu/other/mult-pkg/introduction-to-linear-mixed-models/>

Linear mixed models are an extension of simple linear models to allow
for both *fixed effects* and *random effects*. These are used when there
is a **lack of independence** between variables, and it creates
**groups** or **clusters** in the data.

**Fixed Effects**: variables that we expect to have an effect on the
response variable (i.e. independent or explanatory variables)

**Random Effects**: [grouping factors]{.underline} that may cause
variation that we want to control; *ALWAYS CATEGORICAL*

The data for our random effects are just a sample of all the
possibilities - our data are limited by time and resources, and
therefore are only a representation of part of the population.

## Generalized Linear Mixed Models

Source:
<https://stats.oarc.ucla.edu/other/mult-pkg/introduction-to-generalized-linear-mixed-models/>

GLMMs are a broader category of LMMs that account for data whose
**distributions are non-uniform** (non-normal, or not gaussian). This
includes binary responses. We need to add a **link function** (example:
a log link)

I believe that using lmer with an argument for family is equivalent to
running a GLMM. The [docs
site](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/family)
lists these as the possible args:

[family(object, …)]{.underline}

binomial(link = "logit")

gaussian(link = "identity") Gamma(link = "inverse")

inverse.gaussian(link = "1/mu\^2")

poisson(link = "log")

quasi(link = "identity", variance = "constant")

quasibinomial(link = "logit")

quasipoisson(link = "log")

For different distributions, we use different link functions and
families:

-   Binary outcome: logistic link function + probability density
    function for the logistic
-   Count outcome: log link function + probability mass function for the
    poisson
-   Continuous outcome w/ normal distribution: identity link function

[Link Functions]{.underline}

identity -\> f(x) = x

inverse -\> f(x) = 1/x

[Family (Default Link Function)]{.underline}

gaussian (link = "identity")

Gamma (link = "inverse") - right skew inverse.gaussian (link =
"1/mu\^2")

quasi (link = "identity", variance = "constant") - assumes constant
variance

Binary binomial (link = "logit") - log

quasibinomial (link = "logit") - log with higher variance

Count poisson (link = "log")

quasipoisson (link = "log")

Source: <https://www.datacamp.com/doc/r/glm>



```{r}
library(lme4)
library(dplyr)
library(ggplot2)
```

## Load Data

```{r}
# File from github
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/Analysis_Ready_Data/"
df <- read.csv(paste0(filepath, "daily_2022_2024.csv"))
df <- df %>% select(-X) %>% select(-X.1)
df$Date <- as.POSIXct(df$date, format="%Y-%m-%d")
```

We need to look at the distribution of our data.

First we fit a linear mixed model.

```{r}
# filter out values from before data collection
df_10cm <- df %>% filter(grounddepth == 8) %>% filter(date >= "2022-06-19") %>% filter(date < "2022-09-01")
#df_10cm$groundtemp <- sqrt(df_10cm$groundtemp)
lmm1 <- lmer(groundtemp ~ airtemp + vwc + (1|site), data = df_10cm)
summary(lmm1)
coef(lmm1)
```
t value is the ratio between estimate and standard error
we don't have p value for significance
larger absolute value of t-value is better, but look at estimate and standard error as opposed to using an arbitrary cutoff
standard error tells us how certain we are of our estimate

median should be close to zero
1Q and 3Q should match up abs value
min max ideally should also match

site vs residual - variance between sites vs variance within sites
it's fine that they are different and shows that we should probably be doing lmm here (because there is a difference) - if variance/standard deviation is close to zero then there isn't much variability between the sites
look at standard deviation because that's actually in our units (so don't scale it to make interpretation easier)

**Random effects** tells you how much variance you find among levels of
your grouping factor(s) as well as the residual variance. The residual
variance is the variation explained by the fixed effects.

Site grouping may have different climate qualities
Mixed effects maybe not to see if there is actual difference between the sites
ANOVA could be good for checking different sites or aspects (check out repeated measures ANOVA?)
NVS - random effect for site but use dummy variables for north vs south (make the variable a factor)

**Fixed effects** are similar to a linear model output with slope and
intercept as well as their respective errors.

## Check Assumptions


## Transform our Variable

```{r}
# Transform data
df_transformed <- df_10cm
df_transformed$groundtemp <- sqrt(df_transformed$groundtemp)
lmm2 <- lmer(groundtemp ~ airtemp + (1|site), data = df_transformed)

# Scale (normalize) data
df_scaled <- df_transformed
df_scaled$groundtemp <- scale(df_scaled$groundtemp)
lmm3 <- lmer(groundtemp ~ airtemp + (1|site), data = df_scaled)
```


```{r}
groundtemp <- df_10cm$groundtemp
groundtemp_sqrt <- df_transformed$groundtemp
groundtemp_scaled <- df_transformed$groundtemp
```


```{r}
plot(lmm1)
hist(groundtemp, freq=FALSE)
ggplot(data = data.frame(sample = groundtemp), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(col = "blue") +
  theme_minimal() +
  ggtitle("Q-Q Plot of Untransformed Data")
```

identify some of the points that the model isn't working as well for
this is good enough


```{r}
plot(lmm2)
hist(groundtemp_sqrt, freq=FALSE)
ggplot(data = data.frame(sample = groundtemp_sqrt), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(col = "blue") +
  theme_minimal() +
  ggtitle("Q-Q Plot of Transformed Data - sqrt(groundtemp)")
```

```{r}
plot(lmm3)
hist(groundtemp_scaled, freq=FALSE)
ggplot(data = data.frame(sample = groundtemp_scaled), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(col = "blue") +
  theme_minimal() +
  ggtitle("Q-Q Plot of Scaled Data - scale(sqrt(groundtemp))")
```

It is good practice to scale our explanatory variables so that we can
effectively compare the effects. Scaling keeps the distribution of the
variable the same shape but it makes the mean value zero and the
standard deviation 1.

**NO because it uses an estimated standard deviation**

```{r}
# df_scaled <- df_10cm
# df_scaled$vwc <- scale(df_scaled$vwc, center = TRUE, scale = TRUE)[,1]
# df_scaled$airtemp <- scale(df_scaled$airtemp, center = TRUE, scale = TRUE)[,1]
# df_scaled$solar <- scale(df_scaled$solar, center = TRUE, scale = TRUE)[,1]
# df_scaled$windspeed <- scale(df_scaled$windspeed, center = TRUE, scale = TRUE)[,1]
# df_scaled$winddir <- scale(df_scaled$winddir, center = TRUE, scale = TRUE)[,1]
```

Sources: <https://ourcodingclub.github.io/tutorials/mixed-models/#what>

START 1. North vs South at one site - try to figure out that there is a
difference 2. South vs South at two different sites Eventually... do we
see a difference between years in our North vs South GLMMs

# GLMM

By season (separate season and year) Then maybe look at one month or one
day within that season Eventually annual comparison

Summer: 6/1-9/1 Fall: 9/1-12/1 Winter: 12/1-3/1 Spring: 3/1-6/1

For now, ignore winter because it was mostly gap filled data.

Run analyses on depth separately within each season.

Depths: 10cm 30cm 55cm 65cm 90cm

## North vs South

TNHA North: TNHA-SC South: TNHA-SA

SSMH North: SSMH-SB South: SSMH-SA

BEO CONTROL: BEO-BASE \| BEO-B05 \| BEO-B06 \| BEO-B07 \| BEO-BO8

```{r}
beo <- df_scaled %>% filter(site == "BEO") %>% group_by(date, Date, grounddepth, vwcdepth, fullname, site, station) %>% summarize(groundtemp = mean(groundtemp), vwc = mean(vwc), airtemp = mean(airtemp), solar = mean(solar), windspeed = mean(windspeed), winddir = mean(winddir)) %>% ungroup() %>% mutate(fullname = "BEO-BASE", site = "BEO", station = "BASE") %>% select(colnames(df))

nvs <- df_scaled %>% filter(fullname == "TNHA-SA" | fullname == "TNHA-SC" |
                       fullname == "SSMH-SA" | fullname == "SSMH-SB") %>% rbind(beo)
```

Note: we cannot compare different aspects at BEO because the control
sensors are located in the tundra in close proximity.

### TNHA

```{r}
tnha_nvs <- nvs %>% filter(site == "TNHA")
mixed.glmer <- lme4::glmer(groundtemp ~ airtemp + (1|station), family=inverse.gaussian(link="1/mu^2"), data = tnha_nvs)
summary(mixed.glmer)
```

Compare with LMER?

```{r}
tnha_nvs <- nvs %>% filter(site == "TNHA")
mixed.lmer <- lmer(groundtemp ~ airtemp + (1|station), data = tnha_nvs)
summary(mixed.lmer)
```

The variance for station is low (0.2252) especially compared to the
variance of the residual (7.9030). This means that the aspect of our
sensor does not account for much variation in ground temp and instead
airtemp accounts for the majority of the variation.

We can calculate the percentage of the random effect's contribution to
variance:

```{r}
0.2252/(0.2252+7.9030)
```

Therefore the random effects (station, or rather aspect) contribute to
approximately 2.77% of the variance in ground temperature at TNHA.

### SSMH

```{r}
ssmh_nvs <- nvs %>% filter(site == "SSMH")
mixed.lmer <- lmer(groundtemp ~ airtemp + (1|station), data = ssmh_nvs)
summary(mixed.lmer)
```

## South vs South

### BEO vs TNHA

### BEO vs SSMH

Seasonal Summer 2022, Fall 2022, Spring 2023 - separate GLMMs Site by
Site Also aspect by aspect (n vs s) Depth: start with 10cm; 30cm; 55cm;
65cm; 90cm

change this order: - aspects - locations - depth - season
