---
title: "Generalized Linear Mixed Models"
author: "Elizabeth Van Metre"
date: "2024-10-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Generalized Linear Mixed Models

**Created: October 1, 2024**

[Changes]{.underline}

-   10/1/24: added lme4 package and began analysis of LMM
-   10/30/24: updated with new averages from gap-filled data

**Fixed Effects**: variables that we expect to have an effect on the
response variable (i.e. independent or explanatory variables)

**Random Effects**: [grouping factors]{.underline} that may cause
variation that we want to control; ALWAYS CATEGORICAL

The data for our random effects are just a sample of all the
possibilities - our data are limited by time and resources, and
therefore are only a representation of part of the population.

```{r}
library(lme4)
```

## Load Data

```{r}
# File from github
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/Analysis_Ready_Data/"
df <- read.csv(paste0(filepath, "daily_2022_2024.csv"))
df <- df %>% select(-X) %>% select(-X.1)
df$Date <- as.POSIXct(df$date, format="%Y-%m-%d")
```

We need to look at the distribution of our data.

First we fit a linear mixed model.

```{r}
df_10cm <- df %>% filter(grounddepth == 8)
mixed.lmer <- lmer(groundtemp ~ airtemp + (1|site), data = df_10cm)
summary(mixed.lmer)
```

**Random effects** tells you how much variance you find among levels of
your grouping factor(s) as well as the residual variance. The residual
variance is the variation explained by the fixed effects.

**Fixed effects** are similar to a linear model output with slope and
intercept as well as their respective errors.

## Check Assumptions

```{r}
plot(mixed.lmer)
```

Response Variable (outcome): groundtemp

We need to make sure that our response variable is normally distributed.
If it is not, then we will have to standardize groundtemp.

```{r}
groundtemp <- df_10cm$groundtemp
hist(groundtemp)
```

It is good practice to scale our explanatory variables so that we can
effectively compare the effects. Scaling keeps the distribution of the
variable the same shape but it makes the mean value zero and the
standard deviation 1.

```{r}
df_scaled <- df_10cm
df_scaled$vwc <- scale(df_scaled$vwc, center = TRUE, scale = TRUE)[,1]
df_scaled$airtemp <- scale(df_scaled$airtemp, center = TRUE, scale = TRUE)[,1]
df_scaled$solar <- scale(df_scaled$solar, center = TRUE, scale = TRUE)[,1]
df_scaled$windspeed <- scale(df_scaled$windspeed, center = TRUE, scale = TRUE)[,1]
df_scaled$winddir <- scale(df_scaled$winddir, center = TRUE, scale = TRUE)[,1]
```

Sources: <https://ourcodingclub.github.io/tutorials/mixed-models/#what>

START 1. North vs South at one site - try to figure out that there is a
difference 2. South vs South at two different sites Eventually... do we
see a difference between years in our North vs South GLMMs

# GLMM

By season (separate season and year)
Then maybe look at one month or one day within that season
Eventually annual comparison

Summer: 6/1-9/1
Fall: 9/1-12/1
Winter: 12/1-3/1
Spring: 3/1-6/1

For now, ignore winter because it was mostly gap filled data.

Run analyses on depth separately within each season.

Depths:
10cm
30cm
55cm
65cm
90cm

## North vs South

TNHA
North: TNHA-SC
South: TNHA-SA

SSMH
North: SSMH-SB
South: SSMH-SA

BEO CONTROL: BEO-BASE \| BEO-B05 \| BEO-B06 \| BEO-B07 \| BEO-BO8

```{r}
beo <- df_scaled %>% filter(site == "BEO") %>% group_by(date, Date, grounddepth, vwcdepth, fullname, site, station) %>% summarize(groundtemp = mean(groundtemp), vwc = mean(vwc), airtemp = mean(airtemp), solar = mean(solar), windspeed = mean(windspeed), winddir = mean(winddir)) %>% ungroup() %>% mutate(fullname = "BEO-BASE", site = "BEO", station = "BASE") %>% select(colnames(df))

nvs <- df_scaled %>% filter(fullname == "TNHA-SA" | fullname == "TNHA-SC" |
                       fullname == "SSMH-SA" | fullname == "SSMH-SB") %>% rbind(beo)
```

Note: we cannot compare different aspects at BEO because the control sensors
are located in the tundra in close proximity.

### TNHA

```{r}
tnha_nvs <- nvs %>% filter(site == "TNHA")
mixed.lmer <- lmer(groundtemp ~ airtemp + (1|station), data = tnha_nvs)
summary(mixed.lmer)
```

The variance for station is low (0.2252) especially compared to the variance of
the residual (7.9030). This means that the aspect of our sensor does not
account for much variation in ground temp and instead airtemp accounts
for the majority of the variation.

We can calculate the percentage of the random effect's contribution to
variance:

```{r}
0.2252/(0.2252+7.9030)
```
Therefore the random effects (station, or rather aspect) contribute to approximately 2.77% of
the variance in ground temperature at TNHA.

### SSMH

```{r}
ssmh_nvs <- nvs %>% filter(site == "SSMH")
mixed.lmer <- lmer(groundtemp ~ airtemp + (1|station), data = ssmh_nvs)
summary(mixed.lmer)
```

## South vs South

### BEO vs TNHA

### BEO vs SSMH

Seasonal Summer 2022, Fall 2022, Spring 2023 - separate GLMMs Site by
Site Also aspect by aspect (n vs s) Depth: start with 10cm; 30cm; 55cm;
65cm; 90cm

change this order: - aspects - locations - depth - season
