---
title: "LMM_NvS"
output:
  pdf_document: default
  html_document: default
---

**Created January 28, 2025**

[Changes]{.underline}

-   1/28/25: loading data


```{r}
library(lme4)
library(dplyr)
library(ggplot2)
```

## Load Data

```{r}
# File from github
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/main/Analysis_Ready_Data/"
df <- read.csv(paste0(filepath, "daily_2022_2024.csv"))
df <- df %>% select(-X) %>% select(-X.1)
df$Date <- as.POSIXct(df$date, format="%Y-%m-%d")
df <- df %>% filter(windspeed != -888.88) %>% filter(winddir != -888.88)
```

## Select and Transform Data

North vs South

TNHA:

North = TNHA-SC

South = TNHA-SA

SSMH:

North = SSMH-SB

South = SSMH-SA

BEO (Control): does not have different aspects

```{r}
nvs <- df %>% filter(fullname == "TNHA-SA" | fullname == "TNHA-SC" | fullname == "SSMH-SB" | fullname == "SSMH-SA") %>% select(-c(winddir, date)) %>% mutate(aspect = case_when(fullname == "TNHA-SC" | fullname == "SSMH-SB" ~ "North", .default = "South")) %>% na.omit() %>% filter(windspeed >= 0)
```


```{r}
# filter out data from before data collection
# filter to get only depth of 10cm for now
df_10cm <- nvs %>% filter(grounddepth == 8) %>% filter(Date >= "2022-06-19")
```


# Fit LMM

```{r}
lmm0 <- lmer(groundtemp ~ airtemp + vwc + solar + windspeed + aspect + (1|site), data = df_10cm)
summary(lmm0)
```

```{r}
lmm1 <- lmer(groundtemp ~ airtemp + vwc + solar + aspect + (1|site), data = df_10cm)
summary(lmm1)
```

```{r}
coef(lmm1)
```
It seems like there is not too much of a difference between the two sites, but there is still presence of a difference as shown by the two different intercepts for SSMH vs TNHA.

Graphing linear models:
Random effects have different intercepts, and the fixed effects have different slopes

```{r}
coeffs = coef(lmm1)$site
```

```{r}

```


