---
title: "LMM_NvS"
output:
  pdf_document: default
  html_document: default
---

**Created January 28, 2025**

[Changes]{.underline}

-   1/28/25: loading data


```{r}
library(lme4)
library(dplyr)
library(ggplot2)
```

## Load Data

```{r}
# File from github
filepath = "https://raw.githubusercontent.com/shabanm2/Utqiagvik/pca-25/Analysis_Ready_Data/"
df <- read.csv(paste0(filepath, "daily_2022_2024.csv"))
df <- df %>% select(-X) %>% select(-X.1)
df$Date <- as.POSIXct(df$date, format="%Y-%m-%d")
df <- df %>% filter(windspeed != -888.88) %>% filter(winddir != -888.88)
```

## Select and Transform Data

North vs South

TNHA:

North = TNHA-SC

South = TNHA-SA

SSMH:

North = SSMH-SB

South = SSMH-SA

BEO (Control): does not have different aspects

```{r}
nvs <- df %>% filter(fullname == "TNHA-SA" | fullname == "TNHA-SC" | fullname == "SSMH-SB" | fullname == "SSMH-SA") %>% select(-c(winddir, date)) %>% mutate(aspect = case_when(fullname == "TNHA-SC" | fullname == "SSMH-SB" ~ "North", .default = "South")) %>% na.omit() %>% filter(windspeed >= 0)
```


```{r}
# filter out data from before data collection
# filter to get only depth of 10cm for now
df_10cm <- nvs %>% filter(grounddepth == 8) %>% filter(Date >= "2022-06-19") %>% filter(Date < "2022-09-01")
```


# Fit LMM

```{r}
lmm0 <- lmer(groundtemp ~ airtemp + vwc + solar + windspeed + aspect + (1|site), data = df_10cm)
summary(lmm0)
```

It seems like there is not too much of a difference between the two sites, but there is still presence of a difference as shown by the two different intercepts for SSMH vs TNHA.

Graphing linear models:
Random effects have different intercepts, and the fixed effects have different slopes

```{r}
coeffs = coef(lmm0)$site
coeffs
```

```{r}
library(ggeffects)  # install the package first if you haven't already, then load it

# Extract the prediction data frame
pred.mm <- ggpredict(lmm0, terms = c("airtemp"))  # this gives overall predictions for the model

# Plot the predictions 

(ggplot(pred.mm) + 
    geom_point(data = df_10cm,                      # adding the raw data (scaled values)
              aes(x = airtemp, y = groundtemp, colour = site)) + 
   geom_line(aes(x = x, y = predicted, color = group)) +          # slope
   #geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error), 
               #fill = "lightgrey", alpha = 0.5) +  # error band
   labs(x = "Air Temperature", y = "Ground Temperature", 
        title = "Relationship between air temperature and ground temperature") + 
   theme_minimal()
)
```



```{r}
# i don't like this one
#library(ggeffects)
#ggpredict(lmm1, terms = c("airtemp","aspect", "site"), type = "random") %>% 
#   plot() +
#   labs(x = "Body Length", y = "Test Score", title = "Effect of body size on intelligence in dragons") + 
#   theme_minimal()
```



# Table Output
```{r}
library(stargazer)
stargazer(lmm0, type = "text",
          digits = 3,
          star.cutoffs = c(0.05, 0.01, 0.001),
          digit.separator = "")
```


```{r}
library(effects)
est<-Effect("airtemp", partial.residuals=T, lmm0)
plot(est)

plot(lmm0)

#library(remef)
#y_partial <- remef(model, fix = "x2", ran = "all")
```

```{r}
ggplot(df_10cm, aes(x=airtemp, y=solar)) + geom_point()
```


```{r}
var_df = df_10cm %>% select(groundtemp, airtemp, vwc, solar, windspeed)
round(cor(var_df),
  digits = 2 # rounded to 2 decimals
)
```

